### 2020-04-19 template.tex
add tex template

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{listings}
% Define slightly more reasonable Listings defaults
\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    prebreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
    frame=lines,
    showtabs=false,
    showspaces=false,
    showstringspaces=false,
    keywordstyle=\color[gray]{0.4}\bfseries,
    commentstyle=\color[gray]{0.65}\itshape,
    numbers=left,
    captionpos=b,
}
\usepackage{color}
\usepackage{fancyvrb}
\usepackage{fvextra}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, breaklines=true}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Devito for large scale elastic modelling and anisotropic inversion.},
            colorlinks=true,
            citecolor=black,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{longtable,booktabs}
\usepackage{caption}
\usepackage{subfig}
\captionsetup[subfloat]{margin=1em}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\bibliographystyle{IEEEtran}

\begin{document}

\title{Devito for large scale elastic modelling and anisotropic inversion}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Louboutin Mathias}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Antlanta, GA \\
mlouboutin3@gatech.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
\IEEEauthorblockA{\textit{Devito Codes}}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Philipp Witte}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{5\textsuperscript{th} Rhodri Nelson}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{6\textsuperscript{th} George Bisbas}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{6\textsuperscript{th} Felix J. Herrmann}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{6\textsuperscript{th} Geragd J. Gorman}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
\IEEEauthorblockA{\textit{TU-Delft}}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes,
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins,
column widths, line spaces, and text fonts are prescribed; please do not
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement
and others are deliberate, using specifications that anticipate your paper
as one part of the entire proceedings, and not as an independent document.
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a
separate text file. Complete all content and organizational editing before
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been
formatted and styled. Do not number text heads---{\LaTeX} will do that
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text,
even after they have been defined in the abstract. Abbreviations such as
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your
equations more compact, you may use the solidus (~/~), the exp function, or
appropriate exponents. Italicize Roman symbols for quantities and variables,
but not Greek symbols. Use a long dash rather than a hyphen for a minus
sign. Punctuate equations with commas or periods when they are part of a
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the
symbols in your equation have been defined before or immediately following
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A
minimum of one author is required for all conference articles. Author names
should be listed starting from left to right and then moving down to the
next line. This is the author sequence that will be used in future citations
and by indexing services. Names should not be listed in columns nor group by
affiliation. Please keep your affiliations as succinct as possible (for
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not
topically subordinate to each other. Examples include Acknowledgments and
References and, for these, the correct style to use is ``Heading 5''. Use
``figure caption'' for your Figure captions, and ``table head'' for your
table title. Run-in heads, such as ``Abstract'', will require you to apply a
style (in this case, italic) in addition to the style provided by the drop
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For
example, the paper title is the primary text head because all subsequent
material relates and elaborates on this one topic. If there are two or more
sub-topics, the next level head (uppercase Roman numerals) should be used
and, conversely, if there are not at least two sub-topics, then no subheads
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and
bottom of columns. Avoid placing them in the middle of columns. Large
figures and tables may span across both columns. Figure captions should be
below the figures; table heads should appear above the tables. Insert
figures and tables after they are cited in the text. Use the abbreviation
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4}
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure*}[htbp]
\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure*}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words
rather than symbols or abbreviations when writing Figure axis labels to
avoid confusing the reader. As an example, write the quantity
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including
units in the label, present them within parentheses. Do not label axes only
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of
quantities and units. For example, write ``Temperature (K)'', not
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B.
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at
the bottom of the column in which it was cited. Do not put footnotes in the
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use
``et al.''. Papers that have not been published, even if they have been
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers
that have been accepted for publication should be cited as ``in press'' \cite{b5}.
Capitalize only the first word in a paper title, except for proper nouns and
element symbols.

For papers published in translation journals, please give the English
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2020-04-20 paper.tex
finish pass + add tex file

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{listings}
% Define slightly more reasonable Listings defaults
\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    prebreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
    frame=lines,
    showtabs=false,
    showspaces=false,
    showstringspaces=false,
    keywordstyle=\color[gray]{0.4}\bfseries,
    commentstyle=\color[gray]{0.65}\itshape,
    numbers=left,
    captionpos=b,
}
\usepackage{color}
\usepackage{fancyvrb}
\usepackage{fvextra}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, breaklines=true}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Devito for large scale elastic modelling and anisotropic inversion.},
            colorlinks=true,
            citecolor=black,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{longtable,booktabs}
\usepackage{caption}
\usepackage{subfig}
\captionsetup[subfloat]{margin=1em}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\bibliographystyle{IEEEtran}

\begin{document}

\title{Devito for large scale elastic modelling and anisotropic inversion}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Louboutin Mathias}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Antlanta, GA \\
mlouboutin3@gatech.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
\IEEEauthorblockA{\textit{Devito Codes}}
\and
\IEEEauthorblockN{3\textsuperscript{th} Rhodri Nelson}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{4\textsuperscript{rd} Philipp Witte}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
\IEEEauthorblockA{\textit{TU-Delft}}
\and
\IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{8\textsuperscript{th} Geragd J. Gorman}
\IEEEauthorblockA{\textit{Imperial College London}}
}

\maketitle

\begin{abstract}\label{abstract}

\href{https://github.com/devitocodes/devito}{Devito} is an open-source
Python project based on domain-specific language and compiler
technology. Driven by the requirements of rapid HPC applications
development in exploration seismology, the language and compiler have
evolved significantly since inception. Sophisticated boundary
conditions, tensor contractions, sparse operations and features such as
staggered grids and sub-domains are all supported; operators of
essentially arbitrary complexity can be generated. To accommodate this
flexibility whilst ensuring performance, data dependency analysis is
utilized to schedule loops and detect computational-properties such as
parallelism. In this article, the generation and simulation of
MPI-parallel propagators (along with their adjoints) for the
pseudo-acoustic wave-equation in tilted transverse isotropic media and
the elastic wave-equation are presented. Simulations are carried out on
industry scale synthetic models in a HPC Cloud system and reach a
performance of 28TFLOP/s, hence demonstrating Devito's suitability for
production-grade seismic inversion problems.
\end{abstract}

\section{Introduction}\label{introduction}

Seismic imaging methods such Reverse Time Migration (RTM) and
Full-waveform inversion (FWI) rely on the numerical solution of an
underlying system of partial differential equations (PDEs), most
commonly some manifestation of the wave-equation. In the context of FWI,
the finite-difference (FDM) and the spectral-element (SEM) methods are
most frequently used to solve the wave-equation, with FDM methods
dominating within the seismic exploration community \cite{lyu2020}.
Various forms of the wave-equation and modelling strategies for FWI are
detailed in \cite{fichtner2011}.

Despite the theory of FWI dating back to the 1980s
\cite{tarantola198, tarantola1987}, among the first successful
expositions on real 3D data was presented in \cite{sirgue2009}. Other
studies utilizing FDM within the FWI workflow include
\cite{ratcliffe2011, petersson2013}. The aforementioned studies
approximate the underlying physics via the acoustic wave-equation;
higher fidelity models solving the non-isotropic elastic wave-equation
have been developed in, e.g.,
\cite{osti_1468379, osti_1561580, osti_1561581, sava1, sava2}. Owing to
the flexibility of the mathematical discretizations that can be
utilized, along with the ability to describe problems on complex meshes,
there has also been a great deal of interest in utilizing SEM to solve
inversion problems \cite{peter2011, krebsdg}. The recent study
\cite{trinh2019} presents an efficient SEM based inversion scheme using
a viscoelastic formulation of the wave-equation.

It is generally accepted that the PDE solver utilized within an
inversion workflow must satisfy the following criteria
\cite{virieuxmodelling}: - Efficient for multiple-source modeling - The
memory requirement of the modeling - The ability of a parallel algorithm
to use an increasing number of processors - Ability of the method to
process models of arbitrary levels of heterogeneity - Reduce the
nonlinearity of FWI - Feasibility of the extension of the modeling
approach to more realistic physical descriptions of the media.

It is with these specifications in mind that Devito, a symbolic domain
specific language (DSL) and compiler for the automatic generation of
finite-difference stencils, has been developed. Originally deigned to
accelerate research and development in exploration geophysics, the
high-level interface, previously described in detail in
\cite{devito-api}, is built on top of \texttt{SymPy} \cite{sympy} and
is inspired by the underlying mathematical formulations and other
high-level DSLs such as FEniCS \cite{fenics} and Firedrake
\cite{firedrake}. This interface allows the user to formulate
wave-equations, and more generally time-dependent PDEs in a simple and
mathematically coherent way. The symbolic API then automatically
generates finite-difference stencils from these mathematical
expressions. One of the main advantages of
\href{https://github.com/devitocodes/devito}{Devito} over other
finite-difference DSLs is that generic expressions such as sparse
operations (i.e.~point source injection or localized measurements) are
fully supported and expressible in a high-level fashion. The second
component of \href{https://github.com/devitocodes/devito}{Devito} is its
compiler (c.f \cite{devito-compiler}) that generates highly optimized C
code. Generated code is then compiled at runtime for the hardware at
hand.

Previous work focused on the DSL and compiler to highlight the potential
application and use cases of Devito. Here, we present a series of
extensions and applications to large-scale three-dimensional problem
sizes as encountered in exploration geophysics. These experiments are
carried out in Cloud-based HPC systems and include elastic forward
modelling using distributed-memory parallelism and imaging based on the
tilted transverse isotropic (TTI) wave-equation
(\cite{virieux, thomsen, zhang-tti, duveneck, louboutin2018segeow}).
These proof of concepts highlight two critical features: First, the
ability of the symbolic interface and the compiler to translate to
large-scale adjoint-based inversion problems that require massive
compute (since thousands of PDEs are solved) as well as large amounts of
memory (since the adjoint state method requires the forward model to be
saved in memory). Secondly, through the elastic modelling example, we
demonstrate that \href{https://github.com/devitocodes/devito}{Devito}
now fully supports and automates vectorial and second order tensorial
staggered-grid finite-differences with the same high-level API
previously presented for a scalar fields defined on cartesian grids.

This paper is organized as follows: First, we provide a brief overview
of \href{https://github.com/devitocodes/devito}{Devito} and its symbolic
API and present the distributed memory implementation that allows
large-scale modeling and inversion by means of domain decomposition. We
then provide a brief comparison with a state of the art hand-coded wave
propagator to validate the performance previously benchmarked with the
roofline model
(\cite{patterson, devito-compiler, devito-api, louboutin2016ppf}).
Before concluding, results from the Cloud-based experiments discussed
above are presented, highlighting the vectorial and tensorial
capabilities of Devito.

\section{Overview of Devito}\label{overview-of-devito}

Devito \cite{devito-api} provides a functional language built on top of
\texttt{SymPy} \cite{sympy} to symbolically express PDEs at a
mathematical level and implements automatic discretization with the
finite-difference method. The language is by design flexible enough to
express other types of non finite-diffrence operators, such as
interpolation and tensor contractions, that are inherent to
measurement-based inverse problems. Several additional features are
supported, among which staggered grids, sub-domains, and stencils with
custom coefficients. The last major building block of a good PDE solver
are the boundary conditions, that for finite-difference methods are
notoriously various and often complicated. The system is, however,
sufficiently flexible to express them through composition of core
mechanisms. For example, free surface and perfectly-matched layers
(PMLs) boundary conditions can be expressed as equations -- just like
any other PDE equations -- over a suitable sub-domain.

It is up to the \href{https://github.com/devitocodes/devito}{Devito}
compiler to translate the symbolic specification into C code. The
lowering of the input language down to C consists of several compilation
passes, some of which introducing performance optimizations that are the
key to fast code. Next to classic stencil optimizations (e.g., cache
blocking, alignment, SIMD and OpenMP parallelism),
\href{https://github.com/devitocodes/devito}{Devito} applies a series of
FLOP-reducing transformations as well as aggressive loop fusion. For a
complete treatment, the interested reader should refer to
\cite{devito-compiler}.

\subsection{Symbolic language and
compilation}\label{symbolic-language-and-compilation}

In this section we illustrate the
\href{https://github.com/devitocodes/devito}{Devito} language showing
how to implement the acoustic wave-equation in isotropic media
%
\begin{equation}
\begin{cases}
 m \frac{d^2 u(t, x)}{dt^2} - \Delta u(t, x) = \delta(xs) q(t) \\
 u(0, .) = \frac{d u(t, x)}{dt}(0, .) = 0 \\
 d(t, .) = u(t, xr).
 \end{cases}
\label{acou}
\end{equation}
%
 The core of the \href{https://github.com/devitocodes/devito}{Devito}
symbolic API consists of three classes:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \texttt{Grid}, a representation of the discretized model.
\item
  \texttt{(Time)Function}, a representation of spatially (and time-)
  varying variables defined on a \texttt{Grid} object.
\item
  \texttt{Sparse(Time)Function} a represention of (time-varying) point
  objects on a \texttt{Grid} object, generally unaligned with respect to
  the grid points, hence called ``sparse''.
\end{itemize}

A \texttt{Grid} represents a discretized finite n-dimensional space and
is created as follows

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{Grid}
\NormalTok{grid = Grid(shape=(nx, ny, nz), extent=(ext_x, ext_y, ext_z), origin=(o_x, o_y, o_z))}
\end{Highlighting}
\end{Shaded}

where \texttt{(nx,\phantom{\ }ny,\phantom{\ }nz)} are the number of grid
points in each direction,
\texttt{(ext\_x,\phantom{\ }ext\_y,\phantom{\ }ext\_z)} is the physical
extent of the domain in physical units (i.e \texttt{m}) and
\texttt{(o\_x,\phantom{\ }o\_y,\phantom{\ }o\_z)} is the origin of the
domain in the same physical units. The object \texttt{grid} contains all
the information related to the discretization such as the grid spacing.
We use \texttt{grid} to create the symbolic objects that will be used to
express the wave-equation. First, we define a spatially varying model
parameter \texttt{m} and a time-space varying field \texttt{u}

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{Function, TimeFunction}
\NormalTok{m = Function(name=}\StringTok{"m"}\NormalTok{, grid=grid, space_order=so)}
\NormalTok{u = TimeFunction(name=}\StringTok{"u"}\NormalTok{, grid=grid, space_order=so, time_order=to)}
\end{Highlighting}
\end{Shaded}

where \texttt{so} is the spatial discretization order and \texttt{to}
the time discretization order that is used for the generation of the
finite-difference stencil. Second, we define point-wise objects such as
point sources \texttt{src}, located at physical coordinates $x_s$, and
receiver (measurement) objects \texttt{rec}, with sensors located at the
physical locations $x_r$

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{Function, TimeFunction}
\NormalTok{src = SparseFunction(name=}\StringTok{"src"}\NormalTok{, grid=grid, npoint=}\DecValTok{1}\NormalTok{, coordinates=x_s)}
\NormalTok{d = SparseTimeFunction(name=}\StringTok{"d"}\NormalTok{, grid=grid, npoint=}\DecValTok{1}\NormalTok{, nt=nt, coordinates=x_r)}
\end{Highlighting}
\end{Shaded}

The source term is handled separately from the PDE as a pointwise
operation called \texttt{injection}, while the measurement is handled
with an interpolation. By default,
\href{https://github.com/devitocodes/devito}{Devito} initializes all
\texttt{Function} data to 0, and thus automatically satisfies the zero
Dirichlet condition at \texttt{t=0}. The isotropic acoustic
wave-equation can then be implemented in
\href{https://github.com/devitocodes/devito}{Devito} as below

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{solve, Eq, Operator}
\NormalTok{eq = m * u.dt2 - u.laplace}
\NormalTok{stencil = [Eq(u.forward, solve(eq, u.forward))]}
\NormalTok{src_eqns = s1.inject(u.forward, expr=s1 * dt**}\DecValTok{2} \NormalTok{/ m)}
\NormalTok{d_eqns = d.interpolate(u)}
\end{Highlighting}
\end{Shaded}

To trigger compilation one needs to pass the constructed equations to an
\texttt{Operator}.

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{Operator}
\NormalTok{op = Operator(stencil + src_eqns + d_eqns)}
\end{Highlighting}
\end{Shaded}

The first compilation pass processes the equations individually. The
equations are lowered to an enriched representation, while the
finite-difference constructs (e.g., derivatives) are translated into
actual arithmetic operations. Subsequently, data dependency analysis is
used to compute a performance-optimized topological ordering of the
input equations (e.g., to maximize the likelihood of loop fusion) and to
group them into so called ``clusters''. Basically, a cluster will
eventually be a loop nest in the generated code, and consecutive
clusters may share some outer loops. The ordered sequence of clusters
undergoes several optimization passes, including cache blocking and
FLOP-reducing transformations. It is then further lowered into an
abstract syntax tree, and it is on such representation that parallelisms
is introduced (SIMD, shared-memory, MPI). Finally, all remaining
low-level aspects of code generation are handled, among which the most
relevant one is the data management (e.g., definition of variables,
transfers between host and device).

The output of the \href{https://github.com/devitocodes/devito}{Devito}
compiler for the running example used in this section is available at
\href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
in \texttt{acou-so8.c}.

\subsection{Distributed-memory
parallelism}\label{distributed-memory-parallelism}

We here provide a succinct description of distributed-memory parallelism
in Devito; the interested reader should refer to the mpi tutorial at
\href{https://github.com/devitocodes/devito/blob/master/examples/mpi/overview.ipynb}{mpi-notebook}
for thorough explanations and practical examples.

Devito implements distributed-memory parallelism on top of MPI. The
design is such that users can almost entirely abstract away from it and
reuse non-distributed code as is. Given any
\href{https://github.com/devitocodes/devito}{Devito} code, just running
it as

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DEVITO_MPI=}\DecValTok{1} \NormalTok{mpirun -n X python ...}
\end{Highlighting}
\end{Shaded}

triggers the generation of code with routines for halo exchanges. The
routines are scheduled at a suitable depth in the various loop nests
thanks to data dependency analysis. The following optimizations are
automatically applied:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  redundant halo exchanges are detected and dropped;
\item
  computation/communication overlap, with prodding of the asynchronous
  progress engine by a designated thread through repeated calls to
  \texttt{MPI\_Test};
\item
  a halo exchange is placed as far as possible from where it is needed
  to maximize computation/communication overlap;
\item
  data packing and unpacking is threaded.
\end{itemize}

The domain decomposition occurs in Python upon creation of a
\texttt{Grid} object. Exploiting the MPI Cartesian topology abstraction,
\href{https://github.com/devitocodes/devito}{Devito} logically splits a
grid based on the number of available MPI processes (users are given an
``escape hatch'' to override Devito's default decomposition strategy).
\texttt{Function} and \texttt{TimeFunction} objects inherit the
\texttt{Grid} decomposition. For \texttt{SparseFunction} objects the
approach is different. Since a \texttt{SparseFunction} represents a
sparse set of points,
\href{https://github.com/devitocodes/devito}{Devito} looks at the
physical coordinates of each point and, based on the \texttt{Grid}
decomposition, schedules the logical ownership to an MPI rank. If a
sparse point lies along the boundary of two or more MPI ranks, then it
is duplicated in each of these ranks to be accessible by all neighboring
processes. Eventually, a duplicated point may be redundantly computed by
multiple processes, but any redundant increments will be discarded.

When accessing or manipulating data in a
\href{https://github.com/devitocodes/devito}{Devito} code, users have
the illusion to be working with classic NumPy arrays, while underneath
they are actually distributed. All manner of NumPy indexing schemes
(basic, slicing, etc.) are supported. In the implementation, proper
global-to-local and local-to-global index conversion routines are used
to propagate a read/write access to the impacted subset of ranks. For
example, consider the array

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{A = [[ }\DecValTok{1}\NormalTok{,  }\DecValTok{2}\NormalTok{,  }\DecValTok{3}\NormalTok{,  }\DecValTok{4}\NormalTok{],}
     \NormalTok{[ }\DecValTok{5}\NormalTok{,  }\DecValTok{6}\NormalTok{,  }\DecValTok{7}\NormalTok{,  }\DecValTok{8}\NormalTok{],}
     \NormalTok{[ }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{11}\NormalTok{, }\DecValTok{12}\NormalTok{],}
     \NormalTok{[}\DecValTok{13}\NormalTok{, }\DecValTok{14}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{16}\NormalTok{]])}
\end{Highlighting}
\end{Shaded}

which is distributed across 4 ranks such that \texttt{rank\phantom{\ }0}
contains the elements reading
\texttt{1,\phantom{\ }2,\phantom{\ }5,\phantom{\ }6},
\texttt{rank\phantom{\ }1} the elements
\texttt{3,\phantom{\ }4,\phantom{\ }7,\phantom{\ }8},
\texttt{rank\phantom{\ }2} the elements
\texttt{9,\phantom{\ }10,\phantom{\ }13,\phantom{\ }14} and
\texttt{rank\phantom{\ }3} the elements
\texttt{11,\phantom{\ }12,\phantom{\ }15,\phantom{\ }16}. The slicing
operation \texttt{A{[}::-1,\phantom{\ }::-1{]}} will then return

\begin{Shaded}
\begin{Highlighting}[]
    \NormalTok{[[ }\DecValTok{16}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{14}\NormalTok{, }\DecValTok{13}\NormalTok{],}
     \NormalTok{[ }\DecValTok{12}\NormalTok{, }\DecValTok{11}\NormalTok{, }\DecValTok{10}\NormalTok{,  }\DecValTok{9}\NormalTok{],}
     \NormalTok{[  }\DecValTok{8}\NormalTok{,  }\DecValTok{7}\NormalTok{,  }\DecValTok{6}\NormalTok{,  }\DecValTok{5}\NormalTok{],}
     \NormalTok{[  }\DecValTok{4}\NormalTok{,  }\DecValTok{3}\NormalTok{,  }\DecValTok{2}\NormalTok{,  }\DecValTok{1}\NormalTok{]])}
\end{Highlighting}
\end{Shaded}

such that now \texttt{rank\phantom{\ }0} contains the elements
\texttt{16,\phantom{\ }15,\phantom{\ }12,\phantom{\ }11} and so forth.

Finally, we remark that while providing abstractions for distributed
data manipulation, \href{https://github.com/devitocodes/devito}{Devito}
does not support natively any mechanisms for parallel I/O. The
distributed NumPy arrays, however, provide a generic and flexible
infrastructure for the implementation of parallel I/O for any sort of
file format (e.g., see \cite{witte2018alf}).

\section{Industry-scale 3D seismic imaging in anisotropic
media}\label{industry-scale-3d-seismic-imaging-in-anisotropic-media}

One of the main applications of seismic finite-difference modeling in
exploration geophysics is reverse-time migration (RTM), a wave-equation
based seismic imaging technique. Real-world seismic imaging presents a
number of challenges that make applying this method to industry-scale
problem sizes difficult. First of all, RTM requires an accurate
representation of the physics through sophisticated wave-equations such
as the tilted-transverse isotropic (TTI) wave-equation, for which both
forward and adjoint implementations have to be provided. Secondly,
wave-equations must be solved for a large number of independent
experiments, where each individual PDE solve in itself is expensive in
terms of FLOPs and memory usage. For certain workloads, limited domain
decomposition, that balances the domain size and the number of
independent experiment, and checkpointing techniques have to be applied.
In the following sections, we describe an industry-scale seismic imaging
problem presenting all of the aforementioned challenges, its
implementation with Devito, and the results of an experimentation
carried out on the Azure Cloud using a synthetic data set.

\subsection{Anisotropic wave-equation}\label{anisotropic-wave-equation}

In our seismic imaging case study, we use an anisotropic representation
of the physics called Tilted Transverse Isotropic (TTI) modeling
\cite{thomsen1986}. This representation for wave motion is one of the
most widely used in exploration geophysics since it captures the leading
order kinematics and dynamics of acoustic wave motion in highly
heterogeneous elastic media where the medium properties vary more
rapidly in the direction perpendicular to sedimentary strata
\cite{alkhalifah2000, baysal1983, bubetti2012, bubetti2014, bubesatti2016, chu2011, duveneck, fletcher, fowlertti2010, louboutin2018segeow, whitmore1983, witte2016segpve, xu2014, zhang2005, zhang2011, zhan2013}.
The TTI wave-equation is an acoustic, low dimensional (4 parameters, 2
wavefields) simplification of the 21 parameter and 12 wavefields
tensorial equations of motions \cite{hooke}. This simplified
representation is parametrized by the Thomsen parameters
$\epsilon(x), \delta(x)$ that relate to the global (many wavelength
propagation) difference in propagation speed in the vertical and
horizontal directions, and the tilt and azimuth angles
$\theta(x), \phi(x)$ that define the rotation of the vertical and
horizontal axis around the cartesian directions. However, unlike the
scalar isotropic acoustic wave-equation itself, the TTI wave-equation is
extremely computationally costly to solve and it is also not
self-adjoint as shown in \cite{louboutin2018segeow}.

The main complexity of the TTI wave-equation is due to rotation of the
symmetry axis of the physics that leads to rotated second-order
finite-difference stencils. In order to ensure numerical stability,
these rotated finite-difference operators are designed to be
self-adjoint (c.f. \cite{zhang2011, duveneck}). For example, we define
the rotated second order derivative with respect to $x$ as:
%
\begin{equation}
\begin{aligned}
  G_{\bar{x}\bar{x}} &= D_{\bar{x}}^T D_{\bar{x}} \\
  D_{\bar{x}} &= \cos(\mathbf{\theta})\cos(\mathbf{\phi})\frac{\mathrm{d}}{\mathrm{d}x} + \cos(\mathbf{\theta})\sin(\mathbf{\phi})\frac{\mathrm{d}}{\mathrm{d}y} - \sin(\mathbf{\theta})\frac{\mathrm{d}}{\mathrm{d}z}.
\end{aligned}
\label{rot}
\end{equation}
%
 We enable the simple expression of these complex stencils in
\href{https://github.com/devitocodes/devito}{Devito} as
finite-difference shortcuts, such as \texttt{u.dx} where \texttt{u} is a
\texttt{Function}, are enabled not only for the basic types but for
generic expressions, for example
\texttt{(u\phantom{\ }+\phantom{\ }v.dx).dy}. As a consequence, the
rotated derivative defined in~\ref{rot} is implemented with
\href{https://github.com/devitocodes/devito}{Devito} in two lines as:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dx_u = cos(theta) * cos(phi) * u.dx + cos(theta) * sin(phi) * u.dy - sin(theta) * u.dz}
\NormalTok{dxx_u = (cos(theta) * cos(phi) * dx_u).dx.T + (cos(theta) * sin(phi) * dx_u).dy.T - (sin(theta) * dx_u).dz.T}
\end{Highlighting}
\end{Shaded}

Note that while the adjoint of the finite-difference stencil is enabled
via the standard Python \texttt{.T} shortcut, the expression needs to be
reordered by hand as the tilt and azymuth angle are spatially dependent
and require to be inside the second pass of first-order derivative. We
can see from these simple two lines that the rotated stencil involves
all second-order derivatives (\texttt{.dx.dx}, \texttt{.dy.dy} and
\texttt{.dz.dz}) and all second-order cross-derivatives (\texttt{dx.dy},
\texttt{.dx.dz} and \texttt{.dy.dz}) that leads to a denser stencil
support and higher computational complexity (c.f.
\cite{louboutin2016ppf}). For illustration purpose, the complete
generated code for tti modeling with and without MPI is made available
at
\href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
in \texttt{tti-so8.c} and \texttt{tti-so8-mpi.c}.

Because of the very high number of floating-point operations (FLOP)
needed per grid point for the weighted rotated Laplacian, this
anisotropic wave-equation is extremely challenging to implement. As we
show in Table~\ref{ttiFLOPs}, and previously analysed in
\cite{louboutin2016ppf}, the computational cost with high-order
finite-difference is in the order of thousands of FLOPs per grid point
without optimizations.

\begin{table}
\centering
\begin{tabular}{lll}
\toprule\addlinespace
spatial order & w/o optimizations & w/ optimizations\tabularnewline
\midrule
4 & 501 & 95\tabularnewline
8 & 539 & 102\tabularnewline
12 & 1613 & 160\tabularnewline
16 & 5489 & 276\tabularnewline
\bottomrule
\end{tabular}
\caption{Per-grid-point FLOPs of the finite-difference stencil for the
TTI wave-equation with different spatial discretization
orders.}\label{ttiFLOPs}
\end{table}

The version without FLOP-reducing optimizations is a direct translation
of the discretized operators into stencil expressions. The version with
optimizations employs transformations such as common sub-expressions
elimination, factorization, and cross-iteration redundancy elimination
-- the latter being key in removing the redundancies induced by mixed
derivatives. Implementing all of these techniques manually is inherently
difficult and laborious. Further, to obtain the desired performance
improvements it is necessary to orchestrate them with aggressive loop
fusion (for data locality), tiling (for data locality and tensor
temporaries), and potentially ad-hoc vectorization strategies (if
rotating registers are used). While an explanation of the optimization
strategy employed by
\href{https://github.com/devitocodes/devito}{Devito} is beyond the scope
of this paper (see \cite{devito-compiler} for details), what should be
appreciated here is that all this complexity is hidden away from the
users.

With such complex physics, mathematics, and engineering, it becomes
evident that the implementation of a solver for this wave-equation is
exceptionally time-consuming and can lead to thousands of lines of code
even for a single type of discretization. The verification of the result
is no less complicated, since any small error is effectively untrackable
and any change to the finite-difference scheme or to the time-stepper is
difficult to achieve without substantial re-coding. Another complication
stems from the fact that practitioners of seismic inversion are often
geoscientists and not computer scientists/programmers. Unfortunately,
this background often either results in poorly written low performant
codes or it leads to complications when research codes are handed off to
computer scientists who know how to write fast codes but who often miss
the necessary geophysical domain knowledge. Neither situation is
conducive to addressing the complexities that come with implementing
codes based on the latest geophysical insights and high-performance
computing. With \href{https://github.com/devitocodes/devito}{Devito} on
the other hand, both the forward and adjoint equations can be
implemented in a few lines of Python code as illustrated with the
rotated operator in \#rotxpy.

Simulation of wave motion is only one aspect of solving problems in
seismology. During wave-equation based imaging, we also need to compute
sensitivities (gradient) with respect to the quantities of interest.
This imposes additional constraints on the design and implementations of
our simulation codes as outlined in \cite{virieux}. Among several
factors, such as fast setup time, we focused on correct and testable
implementations for the adjoint wave-equation and the gradient (action
of the adjoint Jacobian)
\cite{louboutin2018segeow, louboutin2020THmfi}, which exactness is
mandatory requirements for gradient based iterative optimization
algorithms.

\subsection{3D Imaging example on
Azure}\label{d-imaging-example-on-azure}

We now demonstrate the scalability of
\href{https://github.com/devitocodes/devito}{Devito} to real-world
applications with the imaging of an industry-scale three-dimensional TTI
subsurface model. This imaging was carried out in the Cloud on Azure and
takes advantage of recent work to port conventional cluster code to the
Cloud using a serverless approach. The serverless implementation is
described in detail in \cite{witte2019TPDedas, witte2019SEGedw} and
describes the steps to run computationnaly and financially efficient HPC
workloads in the Cloud. This imaging project, in collaboration with
Azure, demonstrates the scalability and robustness of
\href{https://github.com/devitocodes/devito}{Devito} to large scale
wave-equation based inverse problems and its cost-effectiveness in
combination with a serverless implementation of seismic imaging in the
Cloud. In this example, we imaged a synthetic three-dimensional
anisotropic subsurface model that mimics a realistic industry size
problem with a realistic representation of the physics (TTI). The
physical size of the problem is \texttt{10kmx10kmx2.8km} discretized on
a \texttt{12.5m} grid with 40 points of absorbing layer on each sides
that leads to \texttt{881x881x371} computational grid points (300
Million grid points). The final image is the sum of 1500 single-source
images and we computed 100 single-source images in parallel on 200 nodes
available using two ndes per source experiment.

\textbf{\emph{Computational performance}}

We briefly describe the computational setup and the performance achieved
for this anisotropic imaging problem. Due to time constraints, and
because the resources we were given access to for this proof of concept
with Azure were somewhat limited, we did not have access to HPC virtual
machines (VM) nor Infiniband enabled ones. The nodes we ran this
experiment on are \texttt{Standard\_E64\_v3} and
\texttt{Standard\_E64s\_v3} that while not HPC VM are memory optimized
allowing to save the wavefield in memory for imaging (TTI adjoint state
gradient \cite{virieux, louboutin2018segeow}). These VMs are Intel®
Broadwell E5-2673 v4 2.3GH that are dual socket, 32 physical cores (and
hyperthreading enabled) and 432Gb of DRAM. The overall inversion
involved computing the image for 1500 source positions, i.e.~solving
1500 forward and 1500 adjoint TTI wave-equations. A single image
required 600Gb of memory in single precision and we used two VMs per
source with MPI with one rank per socket (4 MPI ranks per source) and
imaged 100 sources in parallel. The performance achieved, in single
precision, was as follow:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  140 GFLOP/s per VM;
\item
  280 GFLOP/s per source;
\item
  28 TFLOP/s for all 100 running sources;
\item
  110min runtime per source (forward + adjoint + image computation).
\end{itemize}

We also observe that if we had had access to more resources, and because
the imaging problem is embarassingly parallel over sources and can scale
to any computational ressource, we could have attempted imaging of all
of the 1500 sources in parallel, which theoretically leads to a
performance of 0.4PFLOP/s.

\textbf{\emph{How was the performance measured}}

The execution time is computed through Python-level timers prefixed by
an MPI barrier. The floating-point operations are counted once all of
the symbolic FLOP-reducing transformations have been performed during
compilation. \href{https://github.com/devitocodes/devito}{Devito} uses
an in-house estimate of cost, rather than \texttt{SymPy}'s estimate, to
take care of some low-level intricacies. For example, Devito's estimate
ignores the cost of integer arithmetic used for offset indexing into
multi-dimensional arrays. To calculate the total number of FLOPs
performed, \href{https://github.com/devitocodes/devito}{Devito}
multiplies the floating-point operations calculated at compile time by
the size of the iteration space, and it does that at the granularity of
individual expressions. Thanks to aggressive code motion, the amount of
innermost-loop-invariant sub-expressions in a
\href{https://github.com/devitocodes/devito}{Devito} Operator is
typically negligible, so the
\href{https://github.com/devitocodes/devito}{Devito} estimate does not
basically suffer from this issue, or at least not in a tangible way to
the best of our knowledge. The Devito-reported GFLOP/s were also checked
against those produced by Intel Advisor on several single-node
experiments: the differences -- typically
\href{https://github.com/devitocodes/devito}{Devito} underestimating the
achieved performance -- were always at most in the order of units, and
therefore negligible.

\textbf{\emph{Imaging result}}

The subsurface velocity model that was used in this study is an
artificial anisotropic model that is designed and built combining two
broadly known and used open-source SEG/EAGE acoustic velocity models
that come each with realistic geophysical imaging challenges such as
sub-salt imaging. The anisotropy parameters are derived from a smoothed
version of the velocity while the tilt angles were derived from a
combination of the smooth velocity models and vertical and horizontal
edge detection. The final seismic image of the subsurface model is
plotted in Figure~\ref{OverTTI} and highlights the fact that 3D seismic
imaging based on a serverless approach and automatic code generation is
feasible and provides good results on a realistic model.

\begin{figure*}
\centering
\captionsetup[subfigure]{labelformat=empty}
\subfloat[]{\includegraphics[width=0.500\hsize]{./Figures/OverTTI1.png}}
\subfloat[]{\includegraphics[width=0.500\hsize]{./Figures/OverTTI2.png}}
\\
\subfloat[]{\includegraphics[width=0.500\hsize]{./Figures/OverTTI3.png}}
\subfloat[]{\includegraphics[width=0.500\hsize]{./Figures/OverTTI4.png}}
\caption{3D TTI imaging on a custom made model.}\label{OverTTI}
\end{figure*}

\cite{witte2019TPDedas} fully describes the serverless implementation
of seismic inverse problems, including iterative algorithms for
least-square minimization problems (LSRTM). The 3D anisotropic imaging
results were presented as part of a keynote presentation at the EAGE HPC
workshop in October 2019 \cite{herrmann2019EAGEHPCaii} and at the Rice
O\&G HPC workshop \cite{witte2019RHPCssi} where we focused on the
serverless implementation of seismic inversion algorithms in the Cloud.
This work illustrates the flexibility and portability of Devito, as we
were able to easily port a code only tested and developed on local
hardware to the Cloud, with only minor adjustments. This portability
included the possibility to run MPI-based code for domain decomposition
in the Cloud, after developing it on a desktop computer. Our experiments
are reproducible using the instructions in a public repository
\href{https://github.com/slimgroup/Azure2019}{AzureTTI}, which contains,
among the other things, the Dockerfiles and Azure
\href{https://batch-shipyard.readthedocs.io}{batch-shipyard} setup. This
example can also easily ber un on a traditional HPC cluster environment
using for exaple \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or
Dask \cite{dask} for parallelization over sources.

\section{Elastic modelling}\label{elastic-modelling}

While the subsurface image in the previous Section is obtained with
anisotropic propagators capable of mimicing the real physics, in order
to model both the kinematics and the amplitudes correctly elastic
propagators are required. These propagators are for example extremely
important for global seismology, as the Shear surface waves (component
ignored in TTI) are the most hazardous ones. In this section, we exploit
the tensor algebra language introduced in
\href{https://github.com/devitocodes/devito}{Devito} v4.0 to express
elastic waves in a compact and elegant notation.

The elastic isotropic wave-equation, parametrized by the Lamé parameters
$\lambda, \mu$ and the density $\rho$ reads:
%
\begin{equation}
\begin{aligned}
&\frac{1}{\rho}\frac{dv}{dt} = \nabla . \tau \\
&\frac{d \tau}{dt} = \lambda \mathrm{tr}(\nabla v) \mathbf{I}  + \mu (\nabla v + (\nabla v)^T)
\end{aligned}
\label{elas1}
\end{equation}
%
 where $v$ is a vector valued function with one component per cartesian
direction:
%
\begin{equation}
\begin{split}
v =  \begin{bmatrix} v_x(t, x, y) \\ v_y(t, x, y)) \end{bmatrix}
\end{split}
\label{partvel}
\end{equation}
%
 and the stress $\tau$ is a symmetric second-order tensor-valued
function:
%
\begin{equation}
\begin{aligned}
    \tau = \begin{bmatrix}\tau_{xx}(t, x, y) & \tau_{xy}(t, x, y)\\\tau_{xy}t, x, y) & \tau_{yy}(t, x, y)\end{bmatrix}.
\end{aligned}
\label{stress}
\end{equation}
%
 The discretization of such a set of coupled PDEs requires five
equations in two dimensions (two equations for the particle velocity and
three for stress) and nine equations in three dimensions (three particle
velocities and six stress equations). However, the mathematical
definition only requires two coupled vector/tensor-valued equations for
any number of dimensions.

\subsection{Tensor algebra language}\label{tensor-algebra-language}

We have augmented the
\href{https://github.com/devitocodes/devito}{Devito} language with
tensorial objects to enable a straightforward -- and at the same time
mathematically rigorous -- definition of high-dimensional PDEs, such as
the elastic wave-equation in Equation~\ref{elas1}. This effort was
inspired by \cite{ufl}, a functional language for finite element
methods.

The extended \href{https://github.com/devitocodes/devito}{Devito}
language introduces two new types, \texttt{VectorFunction} (and
\texttt{VectorTimeFunction}) for vectorial objects such as the particle
velocity, and \texttt{TensorFunction} (and \texttt{TensorTimeFunction})
for second-order tensor objects (matrices) such as the stress. These new
objects are constructed the exact same way as the scalar
\texttt{Function} objects. They also automatically implement staggered
grid and staggered finite-differences with the possibility of half-node
averaging. Each component of a tensorial object -- a (scalar)
\href{https://github.com/devitocodes/devito}{Devito} \texttt{Function}
-- is accessible via conventional vector notation (i.e.
\texttt{v{[}0{]},\phantom{\ }t{[}0,1{]},....}).

With this extended language, the elastic wave-equation can be defined in
only four lines as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{v = VectorTimeFunction(name=}\StringTok{'v'}\NormalTok{, grid=model.grid, space_order=so, time_order=}\DecValTok{1}\NormalTok{)}
\NormalTok{tau = TensorTimeFunction(name=}\StringTok{'t'}\NormalTok{, grid=model.grid, space_order=so, time_order=}\DecValTok{1}\NormalTok{)}

\NormalTok{u_v = Eq(v.forward, model.damp * (v + s/rho*div(tau)))}
\NormalTok{u_t = Eq(tau.forward,  model.damp *  (tau + s * (l * diag(div(v.forward)) + mu * (grad(v.forward) + grad(v.forward).T))))}
\end{Highlighting}
\end{Shaded}

The \texttt{SymPy} expressions created by these commands can be
displayed with \texttt{sympy.pprint} as shown in
Figure~\ref{PrettyElas}. This representation reflects perfectly the
mathematics while still providing computational portability and
efficiency through the
\href{https://github.com/devitocodes/devito}{Devito} compiler. The
complete generated code for the elastic wave-equation with and without
MPI is made available at
\href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
in \texttt{elastic-so8.c} and \texttt{elastic-so8-mpi.c}.

\begin{figure*}
\centering
\includegraphics[width=1.000\hsize]{./Figures/vel_symb.png}
\caption{Update stencil for the particle velocity. The stencil for
updating the stress component is left out for readability, as the
equation does not fit onto a single page. However, it can be found in
the \href{https://github.com/devitocodes/devito}{Devito} tutorial on
elastic modelling on github.}\label{PrettyElas}
\end{figure*}

\subsection{2D example}\label{d-example}

We show the elastic particle velocity and stress for a broadly
recognized 2D synthetic model, the elastic
Marmousi-ii\cite{versteeg927, marmouelas} model. The wavefields are
shown on Figure~\ref{ElasWf} and its corresponding elastic shot records
are displayed in Figure~\ref{ElasShot}. Thes two figures show that the
wavefield is as expected purely acoustic in the water layer
(\texttt{t\_\{xy\}=0}) and transition perfectly at the ocean bottom to
an elastic wavefield. We can also clearly see the Shear wave wavefront
in the subsurface (around 1km depth). Thes two plots show that we
effectively solve the elastic wave-equation with an high-level
mathematical definition of the PDE that abstracts away all the
computational technicalities such as staggered grid analysis. The shot
records in Figure~\ref{ElasShot} also match the original data generated
by the creator of the elastic model in \cite{marmouelas}.

\begin{figure*}
\centering
\includegraphics[width=1.000\hsize]{./Figures/marmou_snap.png}
\caption{Particle velocities and stress at time $t=3\text{s}$ for a
source at 10m depth and \texttt{x=5\textbackslash{}text\{km\}} in the
marmousi-ii model.}\label{ElasWf}
\end{figure*}

\begin{figure*}
\centering
\captionsetup[subfigure]{labelformat=empty}
\subfloat[]{\includegraphics[width=0.300\hsize]{./Figures/pressure_marmou.png}}
\subfloat[]{\includegraphics[width=0.300\hsize]{./Figures/vz_marmou.png}}
\subfloat[]{\includegraphics[width=0.300\hsize]{./Figures/vx_marmou.png}}
\caption{Seismic shot record for 5sec of modelling. \texttt{a} is the
pressure (trace of stress tensor) at the surface (5m depth), \texttt{b}
is the vertical particle velocity and \texttt{c} is the horizontal
particle velocity at the ocean bottom (450m depth).}\label{ElasShot}
\end{figure*}

\subsection{3D proof of concept}\label{d-proof-of-concept}

Finally, we modelled three dimensional elastic data in the Cloud to
demonstrate the scalability of
\href{https://github.com/devitocodes/devito}{Devito} to cluster-size
problems in the Cloud. The model we chose mimics the reference model in
geophysics known as the SEAM model \cite{fehler2011seam} that is a
three dimensional extreme-scale synthetic representation of the
subsurface. The physical dimension of the model are
\texttt{45kmx35kmx15km} then discretized with a grid spacing of
\texttt{20mx20mx10m} that led to a computational grid of
\texttt{2250x1750x1500} grid points (5.9 billion grid points). One of
the main challenges of elastic modelling is the extreme memory cost due
to the number of wavefield. For a three dimensional propagator, a
minimum of 21 fields need to be stored:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Three particle velocities with two time steps (\texttt{v.forward} and
  \texttt{v})
\item
  Six stress with two time steps (\texttt{tau.forward} and \texttt{tau})
\item
  Three model parameters \texttt{lamda}, \texttt{mu} and \texttt{rho}
\end{itemize}

These 21 fields, with the grid we just describe, leads to a minimum of
461Gb of memory for modelling only. For this experiment, we obtained
access to small HPC VM on azure called \texttt{Standard\_H16r} that are
16 cores Intel Xeon E5 2667 v3 with no hyperthreading and used 32 nodes
for a single source experiment (we solved a single wave-equation). We
used a 12th order discretization that leads to 2.8TFLOP/time-step to be
computed for this model and propagated the elastic wave for 16 seconds
(23000 time steps). The modelling ran in 16 hours which converts to
1.1TFLOP/s. While these number may seem low, the elastic kernel is
extremely memory bound, while the TTI kernel is nearly compute bound
(see rooflines in \cite{louboutin2016ppf, devito-api, devito-compiler})
making it more computationally efficient, in particular in combination
with MPI. Future work with involve working on the InfiniBand enabled and
true HPC VM on azure to achieve Cloud performance on par with state of
the art Cluster performance. From the performance obtained in this
experiement, and assuming a fairly standard setup of 5000 independent
source experiements, computing a synthetic dataset would require 322
EFLOP to be computed (23k time-steps x 2.8TFLOP/time-step x 500
sources), or using the full scalability and computing all sources in
parallel would lead to 5.5PFLOP/s from our single source experiement
result.

\section{Performance comparison with other
codes}\label{performance-comparison-with-other-codes}

Because earlier performance benchmark focuse mainly on roofline model
analysis, we also compared the runtime of
\href{https://github.com/devitocodes/devito}{Devito} with a reference
open source hand-coded propagator in collaboration with its author for
compeleteness. This propagator, described in \cite{thorbecke} is a
state of the art elastic kernel (Equation~\ref{elas1}). For a fair
comparison, we ensure, in collaboration with the author of
\href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} that the
physical and computational settings were identical. The setting were the
following:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  2001 by 1001 physical grid points.
\item
  200 grid points of dampening layer (absorbing layer \cite{cerjan}) on
  all four sides (total of 2401x1401 computational grid points).
\item
  10001 time steps.
\item
  Single point source, 2001 receivers.
\item
  Same compiler (gcc/icc) to compile
  \href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} and
  run Devito.
\item
  Intel(R) Xeon(R) CPU E3-1270 v6 @ 3.8GHz.
\item
  Single socket, four physical cores, four physical threads, thread
  pinning to cores and hyperthreading off.
\end{itemize}

The runtime obtained for this problem for both propagators were
identical with less than one percent of difference. This similar runtime
were obtained both with the Intel compiler and the GNU compiler and we
ran the experiment with a fourth order and a sixth order discretization.
The kernels were executed five time each to ensure consistency between
the results and we consistently obtained similar runtimes. This
comparison illustrates the performance achieved with
\href{https://github.com/devitocodes/devito}{Devito} is at least on par
with hand-coded propagators. Considering we do not take adantage of the
\href{https://github.com/devitocodes/devito}{Devito} compiler full
capabilities in the two dimensional case, we are confident that the
generated code will be at least on par with hand-coded version for three
dimensional probalems and this comparison will be part of future work.

\section{Conclusions}\label{conclusions}

The transition from academic size problems, such as the two-dimensional
acoustic wave-equation, to real-world application can be challenging in
particular when attempted as an afterthought. In this work, we showed
that thanks to design principles that aimed at this type of application
from the beginning, \href{https://github.com/devitocodes/devito}{Devito}
provides the scalability necessary. First, we showed that we provide a
high-level interface not only for simple scalar equations but for
coupled PDEs and allow the expression of non-trivial differential
operators in a simple and concise way. Second and most importantly, we
demonstrated that the compiler enables large-scale with state-of-the art
computational performance and programming paradigm. The single-node
performance is on par with hand-coded codes while providing the
flexibility from the symbolic interface, and multi-node parallelism is
integrated in the compiler and interface in a conservative way. Finally,
we demonstrated that our abstraction provides to performance portability
that enables both on-premise HPC and Cloud HPC.

\bibliography{sc20_paper}

\end{document}

### 2020-04-20 paper.tex
Fix references

@@
 Despite the theory of FWI dating back to the 1980s
-\cite{tarantola198, tarantola1987}, among the first successful
+\cite{tarantola}, among the first successful
 expositions on real 3D data was presented in \cite{sirgue2009}. Other
@@
 tilted transverse isotropic (TTI) wave-equation
-(\cite{virieux, thomsen, zhang-tti, duveneck, louboutin2018segeow}).
+(\cite{virieux, thomsen1986, zhang-tti, duveneck, louboutin2018segeow}).
 These proof of concepts highlight two critical features: First, the

### 2020-04-20 paper.tex
fixing layout ...

@@
 }
+
+% Default fixed font does not support bold face
+\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{9} % for bold
+\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{9}  % for normal
+
+% Custom colors
+\usepackage{color}
+\definecolor{deepblue}{rgb}{0,0,0.5}
+\definecolor{deepred}{rgb}{0.6,0,0}
+\definecolor{deepgreen}{rgb}{0,0.5,0}
+
+% Python style for highlighting
+\newcommand\pythonstyle{\lstset{
+    language=Python,
+    basicstyle=\ttm,
+    aboveskip=10pt,
+    belowskip=10pt,
+    otherkeywords={self},             % Add keywords here
+    keywordstyle=\ttb\color{deepblue},
+    emph={MyClass,__init__},          % Custom highlighting
+    emphstyle=\ttb\color{deepred},    % Custom highlighting style
+    stringstyle=\color{deepgreen},
+    frame=tb,                         % Any extra options here
+    showstringspaces=false            % 
+}}
+
+% Python environment
+\lstnewenvironment{python}[1][]
+{
+    \pythonstyle
+    \lstset{#1}
+}
+{}
+
 \usepackage{color}
@@
 
-It is with these specifications in mind that Devito, a symbolic domain
+It is with these specifications in mind that
+\href{https://github.com/devitocodes/devito}{Devito}, a symbolic domain
 specific language (DSL) and compiler for the automatic generation of
 finite-difference stencils, has been developed. Originally deigned to
-accelerate research and development in exploration geophysics, the
-high-level interface, previously described in detail in
-\cite{devito-api}, is built on top of \texttt{SymPy} \cite{sympy} and
-is inspired by the underlying mathematical formulations and other
-high-level DSLs such as FEniCS \cite{fenics} and Firedrake
-\cite{firedrake}. This interface allows the user to formulate
-wave-equations, and more generally time-dependent PDEs in a simple and
-mathematically coherent way. The symbolic API then automatically
-generates finite-difference stencils from these mathematical
-expressions. One of the main advantages of
+accelerate research and development in exploration geophysics, the high-level
+interface, previously described in detail in \cite{devito-api}, is built on top
+of \texttt{SymPy} \cite{sympy} and is inspired by the underlying mathematical
+formulations and other DSLs such as FEniCS \cite{fenics} and Firedrake
+\cite{firedrake}. This interface allows the user to formulate wave-equations,
+and more generally time-dependent PDEs in a simple and mathematically coherent
+way. The Devito compiler then automatically generates finite-difference
+stencils from these mathematical expressions. One of the main advantages of
 \href{https://github.com/devitocodes/devito}{Devito} over other
-finite-difference DSLs is that generic expressions such as sparse
-operations (i.e.~point source injection or localized measurements) are
-fully supported and expressible in a high-level fashion. The second
-component of \href{https://github.com/devitocodes/devito}{Devito} is its
-compiler (c.f \cite{devito-compiler}) that generates highly optimized C
-code. Generated code is then compiled at runtime for the hardware at
-hand.
+finite-difference DSLs is that generic expressions such as sparse operations
+(i.e.~point source injection or localized measurements) are fully supported and
+expressible in a high-level fashion. The second component of
+\href{https://github.com/devitocodes/devito}{Devito} is its compiler (c.f
+\cite{devito-compiler}) that generates highly optimized C code. The generated
+code is then compiled at runtime for the hardware at hand.
 
@@
 (\cite{virieux, thomsen1986, zhang-tti, duveneck, louboutin2018segeow}).
-These proof of concepts highlight two critical features: First, the
+These proof of concepts highlight two critical features: first, the
 ability of the symbolic interface and the compiler to translate to
@@
 
-This paper is organized as follows: First, we provide a brief overview
+This paper is organized as follows: first, we provide a brief overview
 of \href{https://github.com/devitocodes/devito}{Devito} and its symbolic
@@
 notoriously various and often complicated. The system is, however,
-sufficiently flexible to express them through composition of core
+sufficiently general to express them through composition of core
 mechanisms. For example, free surface and perfectly-matched layers
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{Grid}
-\NormalTok{grid = Grid(shape=(nx, ny, nz), extent=(ext_x, ext_y, ext_z), origin=(o_x, o_y, o_z))}
-\end{Highlighting}
-\end{Shaded}
+\begin{python}
+from devito import Grid
+grid = Grid(shape=(nx, ny, nz),
+            extent=(ext_x, ext_y, ext_z),
+            origin=(o_x, o_y, o_z))
+\end{python}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{Function, TimeFunction}
-\NormalTok{m = Function(name=}\StringTok{"m"}\NormalTok{, grid=grid, space_order=so)}
-\NormalTok{u = TimeFunction(name=}\StringTok{"u"}\NormalTok{, grid=grid, space_order=so, time_order=to)}
-\end{Highlighting}
-\end{Shaded}
+\begin{python}
+from devito import Function, TimeFunction
+m = Function(name="m", grid=grid,
+             space_order=so)
+u = TimeFunction(name="u", grid=grid,
+                 space_order=so,
+                 time_order=to)
+\end{python}
 
@@
 point sources \texttt{src}, located at physical coordinates $x_s$, and
-receiver (measurement) objects \texttt{rec}, with sensors located at the
+receiver (measurement) objects \texttt{d}, with sensors located at the
 physical locations $x_r$
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{Function, TimeFunction}
-\NormalTok{src = SparseFunction(name=}\StringTok{"src"}\NormalTok{, grid=grid, npoint=}\DecValTok{1}\NormalTok{, coordinates=x_s)}
-\NormalTok{d = SparseTimeFunction(name=}\StringTok{"d"}\NormalTok{, grid=grid, npoint=}\DecValTok{1}\NormalTok{, nt=nt, coordinates=x_r)}
-\end{Highlighting}
-\end{Shaded}
+\begin{python}
+from devito import SparseTimeFunction
+src = SparseTimeFunction(name="src", grid=grid,
+                         npoint=1,
+                         coordinates=x_s)
+d = SparseTimeFunction(name="d", grid=grid,
+                       npoint=1, nt=nt,
+                       coordinates=x_r)
+\end{python}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{solve, Eq, Operator}
-\NormalTok{eq = m * u.dt2 - u.laplace}
-\NormalTok{stencil = [Eq(u.forward, solve(eq, u.forward))]}
-\NormalTok{src_eqns = s1.inject(u.forward, expr=s1 * dt**}\DecValTok{2} \NormalTok{/ m)}
-\NormalTok{d_eqns = d.interpolate(u)}
-\end{Highlighting}
-\end{Shaded}
+\begin{python}
+from devito import solve, Eq, Operator
+eq = m * u.dt2 - u.laplace
+update = Eq(u.forward, solve(eq, u.forward))
+src_eqns = s1.inject(u.forward, expr=s1*dt**2/m)
+d_eqns = d.interpolate(u)
+\end{python}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{devito }\CharTok{import} \NormalTok{Operator}
-\NormalTok{op = Operator(stencil + src_eqns + d_eqns)}
-\end{Highlighting}
-\end{Shaded}
+\begin{python}
+from devito import Operator
+op = Operator(update + src_eqns + d_eqns)
+\end{python}
 

### 2020-04-20 paper.tex
typos-george

@@
 % Default fixed font does not support bold face
-\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{9} % for bold
-\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{9}  % for normal
+\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8.5} % for bold
+\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8.5}  % for normal
 
@@
     stringstyle=\color{deepgreen},
+    numbers=none,
     frame=tb,                         % Any extra options here
@@
 \and
-\IEEEauthorblockN{8\textsuperscript{th} Geragd J. Gorman}
+\IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
 \IEEEauthorblockA{\textit{Imperial College London}}
@@
 process models of arbitrary levels of heterogeneity - Reduce the
-nonlinearity of FWI - Feasibility of the extension of the modeling
+non-linearity of FWI - Feasibility of the extension of the modeling
 approach to more realistic physical descriptions of the media.
@@
 staggered-grid finite-differences with the same high-level API
-previously presented for a scalar fields defined on cartesian grids.
+previously presented for scalar fields defined on cartesian grids.
 
@@
 from devito import SparseTimeFunction
-src = SparseTimeFunction(name="src", grid=grid,
-                         npoint=1,
+s = SparseTimeFunction(name="src",
+                       grid=grid, npoint=1,
                        coordinates=x_s)
@@
 update = Eq(u.forward, solve(eq, u.forward))
-src_eqns = s1.inject(u.forward, expr=s1*dt**2/m)
+src_eqns = s.inject(u.forward, expr=s*dt**2/m)
 d_eqns = d.interpolate(u)
@@
 We here provide a succinct description of distributed-memory parallelism
-in Devito; the interested reader should refer to the mpi tutorial at
+in Devito; the interested reader should refer to the MPI tutorial at
 \href{https://github.com/devitocodes/devito/blob/master/examples/mpi/overview.ipynb}{mpi-notebook}
@@
 decomposition, that balances the domain size and the number of
-independent experiment, and checkpointing techniques have to be applied.
+independent experiments, and checkpointing techniques have to be applied.
 In the following sections, we describe an industry-scale seismic imaging
@@
 The TTI wave-equation is an acoustic, low dimensional (4 parameters, 2
-wavefields) simplification of the 21 parameter and 12 wavefields
+wavefields) simplification of the 21 parameters and 12 wavefields
 tensorial equations of motions \cite{hooke}. This simplified
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{dx_u = cos(theta) * cos(phi) * u.dx + cos(theta) * sin(phi) * u.dy - sin(theta) * u.dz}
-\NormalTok{dxx_u = (cos(theta) * cos(phi) * dx_u).dx.T + (cos(theta) * sin(phi) * dx_u).dy.T - (sin(theta) * dx_u).dz.T}
-\end{Highlighting}
-\end{Shaded}
+\begin{python}
+dx_u = cos(theta) * cos(phi) * u.dx +
+       cos(theta) * sin(phi) * u.dy -
+       sin(theta) * u.dz
+dxx_u = (cos(theta) * cos(phi) * dx_u).dx.T +
+        (cos(theta) * sin(phi) * dx_u).dy.T -
+        (sin(theta) * dx_u).dz.T
+\end{python}
 
@@
 \centering
-\begin{tabular}{lll}
+\begin{tabular}{ccc}
 \toprule\addlinespace
@@
 implemented in a few lines of Python code as illustrated with the
-rotated operator in \#rotxpy.
+rotational operator in \#rotxpy.
 
@@
 
-\subsection{3D Imaging example on
-Azure}\label{d-imaging-example-on-azure}
+\subsection{3D Imaging example on Azure}\label{d-imaging-example-on-azure}
 
@@
 images and we computed 100 single-source images in parallel on 200 nodes
-available using two ndes per source experiment.
+available using two nodes per source experiment.
 
@@
 the imaging problem is embarassingly parallel over sources and can scale
-to any computational ressource, we could have attempted imaging of all
+to any computational resources, we could have attempted imaging of all
 of the 1500 sources in parallel, which theoretically leads to a
@@
 example can also easily ber un on a traditional HPC cluster environment
-using for exaple \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or
+using for example \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or
 Dask \cite{dask} for parallelization over sources.
@@
 While the subsurface image in the previous Section is obtained with
-anisotropic propagators capable of mimicing the real physics, in order
+anisotropic propagators capable of mimicking the real physics, in order
 to model both the kinematics and the amplitudes correctly elastic
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{v = VectorTimeFunction(name=}\StringTok{'v'}\NormalTok{, grid=model.grid, space_order=so, time_order=}\DecValTok{1}\NormalTok{)}
-\NormalTok{tau = TensorTimeFunction(name=}\StringTok{'t'}\NormalTok{, grid=model.grid, space_order=so, time_order=}\DecValTok{1}\NormalTok{)}
-
-\NormalTok{u_v = Eq(v.forward, model.damp * (v + s/rho*div(tau)))}
-\NormalTok{u_t = Eq(tau.forward,  model.damp *  (tau + s * (l * diag(div(v.forward)) + mu * (grad(v.forward) + grad(v.forward).T))))}
-\end{Highlighting}
-\end{Shaded}
+\begin{python}
+from devito import VectorTimeFunction,
+                   TensorTimeFunction
+v = VectorTimeFunction(name="v", grid=grid,
+                       space_order=so,
+                       time_order=1)
+tau = TensorTimeFunction(name="t", grid=grid,
+                         space_order=so,
+                         time_order=1)
+u_v = Eq(v.forward,
+         damp * (v + s/rho*div(tau)))
+u_t = Eq(tau.forward,
+         damp * (tau + s * (l * diag(div(v.forward)) + mu * (grad(v.forward) + grad(v.forward).T))))
+\end{python}
 
@@
 in the subsurface (around 1km depth). Thes two plots show that we
-effectively solve the elastic wave-equation with an high-level
+effectively solve the elastic wave-equation with a high-level
 mathematical definition of the PDE that abstracts away all the
@@
 the art Cluster performance. From the performance obtained in this
-experiement, and assuming a fairly standard setup of 5000 independent
-source experiements, computing a synthetic dataset would require 322
+experiment, and assuming a fairly standard setup of 5000 independent
+source experiments, computing a synthetic dataset would require 322
 EFLOP to be computed (23k time-steps x 2.8TFLOP/time-step x 500
 sources), or using the full scalability and computing all sources in
-parallel would lead to 5.5PFLOP/s from our single source experiement
+parallel would lead to 5.5PFLOP/s from our single source experiment
 result.
@@
 
-Because earlier performance benchmark focuse mainly on roofline model
+Because earlier performance benchmarks focus mainly on roofline model
 analysis, we also compared the runtime of
@@
 open source hand-coded propagator in collaboration with its author for
-compeleteness. This propagator, described in \cite{thorbecke} is a
+completeness. This propagator, described in \cite{thorbecke} is a
 state of the art elastic kernel (Equation~\ref{elas1}). For a fair
@@
 \item
-  Same compiler (gcc/icc) to compile
+  Same compiler (GCC/ICC) to compile
   \href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} and

### 2020-04-20 paper.tex
typos-george

@@
 % Default fixed font does not support bold face
-\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8.5} % for bold
-\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8.5}  % for normal
+\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{9} % for bold
+\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{9}  % for normal
 
@@
     stringstyle=\color{deepgreen},
-    numbers=none,
     frame=tb,                         % Any extra options here
@@
 from devito import SparseTimeFunction
-s = SparseTimeFunction(name="src",
-                       grid=grid, npoint=1,
+src = SparseTimeFunction(name="src", grid=grid,
+                         npoint=1,
                          coordinates=x_s)
@@
 update = Eq(u.forward, solve(eq, u.forward))
-src_eqns = s.inject(u.forward, expr=s*dt**2/m)
+src_eqns = s1.inject(u.forward, expr=s1*dt**2/m)
 d_eqns = d.interpolate(u)
@@
 
-\begin{python}
-dx_u = cos(theta) * cos(phi) * u.dx +
-       cos(theta) * sin(phi) * u.dy -
-       sin(theta) * u.dz
-dxx_u = (cos(theta) * cos(phi) * dx_u).dx.T +
-        (cos(theta) * sin(phi) * dx_u).dy.T -
-        (sin(theta) * dx_u).dz.T
-\end{python}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{dx_u = cos(theta) * cos(phi) * u.dx + cos(theta) * sin(phi) * u.dy - sin(theta) * u.dz}
+\NormalTok{dxx_u = (cos(theta) * cos(phi) * dx_u).dx.T + (cos(theta) * sin(phi) * dx_u).dy.T - (sin(theta) * dx_u).dz.T}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \centering
-\begin{tabular}{ccc}
+\begin{tabular}{lll}
 \toprule\addlinespace
@@
 
-\subsection{3D Imaging example on Azure}\label{d-imaging-example-on-azure}
+\subsection{3D Imaging example on
+Azure}\label{d-imaging-example-on-azure}
 
@@
 
-\begin{python}
-from devito import VectorTimeFunction,
-                   TensorTimeFunction
-v = VectorTimeFunction(name="v", grid=grid,
-                       space_order=so,
-                       time_order=1)
-tau = TensorTimeFunction(name="t", grid=grid,
-                         space_order=so,
-                         time_order=1)
-u_v = Eq(v.forward,
-         damp * (v + s/rho*div(tau)))
-u_t = Eq(tau.forward,
-         damp * (tau + s * (l * diag(div(v.forward)) + mu * (grad(v.forward) + grad(v.forward).T))))
-\end{python}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{v = VectorTimeFunction(name=}\StringTok{'v'}\NormalTok{, grid=model.grid, space_order=so, time_order=}\DecValTok{1}\NormalTok{)}
+\NormalTok{tau = TensorTimeFunction(name=}\StringTok{'t'}\NormalTok{, grid=model.grid, space_order=so, time_order=}\DecValTok{1}\NormalTok{)}
+
+\NormalTok{u_v = Eq(v.forward, model.damp * (v + s/rho*div(tau)))}
+\NormalTok{u_t = Eq(tau.forward,  model.damp *  (tau + s * (l * diag(div(v.forward)) + mu * (grad(v.forward) + grad(v.forward).T))))}
+\end{Highlighting}
+\end{Shaded}
 

### 2020-04-20 paper.tex
Update paper.tex

--- 2020-04-20 paper.tex
@@
 process models of arbitrary levels of heterogeneity - Reduce the
-non-linearity of FWI - Feasibility of the extension of the modeling
+nonlinearity of FWI - Feasibility of the extension of the modeling
 approach to more realistic physical descriptions of the media.

### 2020-04-20 paper.tex
fixing layout

--- 2020-04-20 paper.tex
@@
 \and
-\IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
+\IEEEauthorblockN{8\textsuperscript{th} Geragd J. Gorman}
 \IEEEauthorblockA{\textit{Imperial College London}}
@@
 staggered-grid finite-differences with the same high-level API
-previously presented for scalar fields defined on cartesian grids.
+previously presented for a scalar fields defined on cartesian grids.
 
@@
 We here provide a succinct description of distributed-memory parallelism
-in Devito; the interested reader should refer to the MPI tutorial at
+in Devito; the interested reader should refer to the mpi tutorial at
 \href{https://github.com/devitocodes/devito/blob/master/examples/mpi/overview.ipynb}{mpi-notebook}
@@
 decomposition, that balances the domain size and the number of
-independent experiments, and checkpointing techniques have to be applied.
+independent experiment, and checkpointing techniques have to be applied.
 In the following sections, we describe an industry-scale seismic imaging
@@
 The TTI wave-equation is an acoustic, low dimensional (4 parameters, 2
-wavefields) simplification of the 21 parameters and 12 wavefields
+wavefields) simplification of the 21 parameter and 12 wavefields
 tensorial equations of motions \cite{hooke}. This simplified
@@
 implemented in a few lines of Python code as illustrated with the
-rotational operator in \#rotxpy.
+rotated operator in \#rotxpy.
 
@@
 images and we computed 100 single-source images in parallel on 200 nodes
-available using two nodes per source experiment.
+available using two ndes per source experiment.
 
@@
 the imaging problem is embarassingly parallel over sources and can scale
-to any computational resources, we could have attempted imaging of all
+to any computational ressource, we could have attempted imaging of all
 of the 1500 sources in parallel, which theoretically leads to a
@@
 example can also easily ber un on a traditional HPC cluster environment
-using for example \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or
+using for exaple \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or
 Dask \cite{dask} for parallelization over sources.
@@
 While the subsurface image in the previous Section is obtained with
-anisotropic propagators capable of mimicking the real physics, in order
+anisotropic propagators capable of mimicing the real physics, in order
 to model both the kinematics and the amplitudes correctly elastic
@@
 in the subsurface (around 1km depth). Thes two plots show that we
-effectively solve the elastic wave-equation with a high-level
+effectively solve the elastic wave-equation with an high-level
 mathematical definition of the PDE that abstracts away all the
@@
 the art Cluster performance. From the performance obtained in this
-experiment, and assuming a fairly standard setup of 5000 independent
-source experiments, computing a synthetic dataset would require 322
+experiement, and assuming a fairly standard setup of 5000 independent
+source experiements, computing a synthetic dataset would require 322
 EFLOP to be computed (23k time-steps x 2.8TFLOP/time-step x 500
 sources), or using the full scalability and computing all sources in
-parallel would lead to 5.5PFLOP/s from our single source experiment
+parallel would lead to 5.5PFLOP/s from our single source experiement
 result.
@@
 
-Because earlier performance benchmarks focus mainly on roofline model
+Because earlier performance benchmark focuse mainly on roofline model
 analysis, we also compared the runtime of
@@
 open source hand-coded propagator in collaboration with its author for
-completeness. This propagator, described in \cite{thorbecke} is a
+compeleteness. This propagator, described in \cite{thorbecke} is a
 state of the art elastic kernel (Equation~\ref{elas1}). For a fair
@@
 \item
-  Same compiler (GCC/ICC) to compile
+  Same compiler (gcc/icc) to compile
   \href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} and

### 2020-04-20 paper.tex
Update tex file.

--- 2020-04-20 paper.tex
@@
 One of the main applications of seismic finite-difference modeling in
-exploration geophysics is reverse-time migration (RTM), a wave-equation
-based seismic imaging technique. Real-world seismic imaging presents a
-number of challenges that make applying this method to industry-scale
-problem sizes difficult. First of all, RTM requires an accurate
-representation of the physics through sophisticated wave-equations such
-as the tilted-transverse isotropic (TTI) wave-equation, for which both
-forward and adjoint implementations have to be provided. Secondly,
-wave-equations must be solved for a large number of independent
-experiments, where each individual PDE solve in itself is expensive in
+exploration geophysics is reverse-time migration (RTM), a wave-equation based
+seismic imaging technique. Real-world seismic imaging presents a number of
+challenges that make applying this method to industry-scale problem sizes
+difficult. Firstly, RTM requires an accurate representation of the underlying
+physics via sophisticated wave-equations such as the tilted-transverse isotropic
+(TTI) wave-equation, for which both forward and adjoint implementations must to
+be provided. Secondly, wave-equations must be solved for a large number of
+independent experiments, where each individual PDE solve is itself expensive in
 terms of FLOPs and memory usage. For certain workloads, limited domain
-decomposition, that balances the domain size and the number of
-independent experiments, and checkpointing techniques have to be applied.
-In the following sections, we describe an industry-scale seismic imaging
-problem presenting all of the aforementioned challenges, its
-implementation with Devito, and the results of an experimentation
-carried out on the Azure Cloud using a synthetic data set.
+decomposition, that balances the domain size and the number of independent
+experiments, and checkpointing techniques must to be adopted. In the following
+sections, we describe an industry-scale seismic imaging problem that poses all
+the aforementioned challenges, its implementation with Devito, and the results
+of an experiment carried out on the Azure Cloud using a synthetic data set.
 
@@
 
-In our seismic imaging case study, we use an anisotropic representation
-of the physics called Tilted Transverse Isotropic (TTI) modeling
-\cite{thomsen1986}. This representation for wave motion is one of the
-most widely used in exploration geophysics since it captures the leading
-order kinematics and dynamics of acoustic wave motion in highly
-heterogeneous elastic media where the medium properties vary more
-rapidly in the direction perpendicular to sedimentary strata
-\cite{alkhalifah2000, baysal1983, bubetti2012, bubetti2014, bubesatti2016, chu2011, duveneck, fletcher, fowlertti2010, louboutin2018segeow, whitmore1983, witte2016segpve, xu2014, zhang2005, zhang2011, zhan2013}.
-The TTI wave-equation is an acoustic, low dimensional (4 parameters, 2
-wavefields) simplification of the 21 parameters and 12 wavefields
-tensorial equations of motions \cite{hooke}. This simplified
-representation is parametrized by the Thomsen parameters
-$\epsilon(x), \delta(x)$ that relate to the global (many wavelength
-propagation) difference in propagation speed in the vertical and
-horizontal directions, and the tilt and azimuth angles
-$\theta(x), \phi(x)$ that define the rotation of the vertical and
-horizontal axis around the cartesian directions. However, unlike the
-scalar isotropic acoustic wave-equation itself, the TTI wave-equation is
-extremely computationally costly to solve and it is also not
-self-adjoint as shown in \cite{louboutin2018segeow}.
-
-The main complexity of the TTI wave-equation is due to rotation of the
-symmetry axis of the physics that leads to rotated second-order
-finite-difference stencils. In order to ensure numerical stability,
-these rotated finite-difference operators are designed to be
-self-adjoint (c.f. \cite{zhang2011, duveneck}). For example, we define
-the rotated second order derivative with respect to $x$ as:
+In our seismic imaging case study, we use an anisotropic representation of the
+physics called Tilted Transverse Isotropic (TTI) modeling \cite{thomsen1986}. This
+representation for wave motion is one of the most widely used in exploration
+geophysics since it captures the leading order kinematics and dynamics of
+acoustic wave motion in highly heterogeneous elastic media where the medium
+properties vary more rapidly in the direction perpendicular to sedimentary
+strata \cite{alkhalifah2000, baysal1983, bubetti2012, bubetti2014,
+bubesatti2016, chu2011, duveneck, fletcher, fowlertti2010,
+louboutin2018segeow, whitmore1983, witte2016segpve, xu2014, ,zhang2005,
+zhang2011, zhan2013}. The TTI wave-equation is an acoustic, low dimensional (4
+parameters, 2 wavefields) simplification of the 21 parameter and 12 wavefields
+tensorial equations of motions \cite{hooke}. This simplified representation is
+parametrized by the Thomsen parameters $\epsilon(x), \delta(x)$ that relate to
+the global (propagation over many wavelengths) difference in propagation speed
+in the vertical and horizontal directions, and the tilt and azimuth angles
+$\theta(x), \phi(x)$ that define the rotation of the vertical and horizontal
+axis around the cartesian directions. However, unlike the scalar isotropic
+acoustic wave-equation itself, the TTI wave-equation is extremely
+computationally costly to solve and it is also not self-adjoint as shown in
+\cite{louboutin2018segeow}.
+
+The main complexity of the TTI wave-equation is that the rotation of the
+symmetry axis of the physics leads to rotated second-order finite-difference
+stencils. In order to ensure numerical stability, these rotated
+finite-difference operators are designed to be self-adjoint (c.f.
+\cite{zhang2011, duveneck}). For example, we define the rotated second order
+derivative with respect to $x$ as:
 %
@@
 %
- We enable the simple expression of these complex stencils in
+ We enable the simple expression of these complicated stencils in
 \href{https://github.com/devitocodes/devito}{Devito} as
-finite-difference shortcuts, such as \texttt{u.dx} where \texttt{u} is a
-\texttt{Function}, are enabled not only for the basic types but for
-generic expressions, for example
+finite-difference shortcuts such as \texttt{u.dx} where \texttt{u}
+is a \texttt{Function}. Such shortcuts are enabled not only for the
+basic types but for generic composite expressions, for example
 \texttt{(u\phantom{\ }+\phantom{\ }v.dx).dy}. As a consequence, the
@@
 
-Note that while the adjoint of the finite-difference stencil is enabled
-via the standard Python \texttt{.T} shortcut, the expression needs to be
-reordered by hand as the tilt and azymuth angle are spatially dependent
-and require to be inside the second pass of first-order derivative. We
-can see from these simple two lines that the rotated stencil involves
-all second-order derivatives (\texttt{.dx.dx}, \texttt{.dy.dy} and
-\texttt{.dz.dz}) and all second-order cross-derivatives (\texttt{dx.dy},
-\texttt{.dx.dz} and \texttt{.dy.dz}) that leads to a denser stencil
-support and higher computational complexity (c.f.
-\cite{louboutin2016ppf}). For illustration purpose, the complete
+Note that while the adjoint of the finite-difference stencil is enabled via the
+standard Python \texttt{.T} shortcut, the expression needs to be reordered by hand
+since the tilt and azymuth angles are spatially dependent and require to be
+inside the second pass of first-order derivative. We can see from these simple
+two lines that the rotated stencil involves all second-order derivatives
+(\texttt{.dx.dx}, \texttt{.dy.dy} and \texttt{.dz.dz}) and all second-order
+cross-derivatives (\texttt{dx.dy}, \texttt{.dx.dz} and \texttt{.dy.dz}) which
+leads to a denser stencil support and higher computational complexity
+(c.f. \cite{louboutin2016ppf}). For illustrative purposes, the complete
 generated code for tti modeling with and without MPI is made available
-at
-\href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
+at \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
 in \texttt{tti-so8.c} and \texttt{tti-so8-mpi.c}.
 
-Because of the very high number of floating-point operations (FLOP)
-needed per grid point for the weighted rotated Laplacian, this
-anisotropic wave-equation is extremely challenging to implement. As we
-show in Table~\ref{ttiFLOPs}, and previously analysed in
-\cite{louboutin2016ppf}, the computational cost with high-order
-finite-difference is in the order of thousands of FLOPs per grid point
-without optimizations.
+Owing to the very high number of floating-point operations (FLOPs) needed per
+grid point for the weighted rotated Laplacian, this anisotropic wave-equation is
+extremely challenging to implement. As we show in Table Table~\ref{ttiFLOPs},
+and previously analysed in \cite{louboutin2016ppf}, the computational cost with
+high-order finite-difference is in the order of thousands of FLOPs per grid
+point without optimizations. The version without FLOP-reducing optimizations is
+a direct translation of the discretized operators into stencil expressions. The
+version with optimizations employs transformations such as common
+sub-expressions elimination, factorization, and cross-iteration redundancy
+elimination -- the latter being key in removing redundancies introduced by mixed
+derivatives. Implementing all of these techniques manually is inherently
+difficult and laborious. Further, to obtain the desired performance improvements
+it is necessary to orchestrate them with aggressive loop fusion (for data
+locality), tiling (for data locality and tensor temporaries), and potentially
+ad-hoc vectorization strategies (if rotating registers are used). While an
+explanation of the optimization strategies employed by
+\href{https://github.com/devitocodes/devito}{Devito} is beyond the scope of this
+paper (see \cite{devito-compiler} for details), what is emphasized here is that
+users can easily take full advantage of these optimization without needed to
+concern themselves with the details.
 
@@
 \end{tabular}
-\caption{Per-grid-point FLOPs of the finite-difference stencil for the
-TTI wave-equation with different spatial discretization
-orders.}\label{ttiFLOPs}
+\caption{Per-grid-point FLOPs of the finite-difference TTI wave-equation stencil
+with different spatial discretization orders.}\label{ttiFLOPs}
 \end{table}
 
-The version without FLOP-reducing optimizations is a direct translation
-of the discretized operators into stencil expressions. The version with
-optimizations employs transformations such as common sub-expressions
-elimination, factorization, and cross-iteration redundancy elimination
--- the latter being key in removing the redundancies induced by mixed
-derivatives. Implementing all of these techniques manually is inherently
-difficult and laborious. Further, to obtain the desired performance
-improvements it is necessary to orchestrate them with aggressive loop
-fusion (for data locality), tiling (for data locality and tensor
-temporaries), and potentially ad-hoc vectorization strategies (if
-rotating registers are used). While an explanation of the optimization
-strategy employed by
-\href{https://github.com/devitocodes/devito}{Devito} is beyond the scope
-of this paper (see \cite{devito-compiler} for details), what should be
-appreciated here is that all this complexity is hidden away from the
-users.
-
-With such complex physics, mathematics, and engineering, it becomes
-evident that the implementation of a solver for this wave-equation is
-exceptionally time-consuming and can lead to thousands of lines of code
-even for a single type of discretization. The verification of the result
-is no less complicated, since any small error is effectively untrackable
-and any change to the finite-difference scheme or to the time-stepper is
-difficult to achieve without substantial re-coding. Another complication
-stems from the fact that practitioners of seismic inversion are often
-geoscientists and not computer scientists/programmers. Unfortunately,
-this background often either results in poorly written low performant
-codes or it leads to complications when research codes are handed off to
-computer scientists who know how to write fast codes but who often miss
-the necessary geophysical domain knowledge. Neither situation is
-conducive to addressing the complexities that come with implementing
-codes based on the latest geophysical insights and high-performance
-computing. With \href{https://github.com/devitocodes/devito}{Devito} on
-the other hand, both the forward and adjoint equations can be
-implemented in a few lines of Python code as illustrated with the
-rotational operator in \#rotxpy.
-
-Simulation of wave motion is only one aspect of solving problems in
-seismology. During wave-equation based imaging, we also need to compute
-sensitivities (gradient) with respect to the quantities of interest.
-This imposes additional constraints on the design and implementations of
-our simulation codes as outlined in \cite{virieux}. Among several
-factors, such as fast setup time, we focused on correct and testable
-implementations for the adjoint wave-equation and the gradient (action
-of the adjoint Jacobian)
-\cite{louboutin2018segeow, louboutin2020THmfi}, which exactness is
-mandatory requirements for gradient based iterative optimization
-algorithms.
+It is evident that developing an appropriate solver for the TTI wave-equation,
+an endeavor involving complicated physics, mathematics, and engineering, is
+exceptionally time-consuming and can lead to thousands of lines of code even for
+a single choice of discretization. Verification of the results is no less
+complicated, any minor error is effectively untrackable and any change to the
+finite-difference scheme or to the time-stepping algorithm is difficult to
+achieve without substantial re-coding. Another complication stems from the fact
+that practitioners of seismic inversion are often geoscientists, not computer
+scientists/programmers. Low level implementations from non-specialists can often
+often lead to poorly performing code. However, if research codes are passed to
+specialists in the domain of low level code optimization they often lack the
+necessary geophysical domain knowledge, resulting in code that may lack a key
+feature required by the geoscientist. Neither situation is conducive to
+addressing the complexities that come with implementing codes based on the
+latest geophysical insights in tandem with those from high-performance
+computing. With \href{https://github.com/devitocodes/devito}{Devito} on the
+other hand, both the forward and adjoint equations can be implemented in a few
+lines of Python code as illustrated with the rotated operator in \#rotxpy. The
+low level optimization element of the development is then taken care of under
+the hood by the \href{https://github.com/devitocodes/devito}{Devito} compiler.
+
+The simulation of wave motion is only one aspect of solving problems in
+seismology. During wave-equation based imaging, it is also required to compute
+sensitivities (gradient) with respect to the quantities of interest. This
+requirement imposes additional constraints on the design and implementation of
+model codes as outlined in \cite{virieux}. Along with several factors, such as
+fast setup time, we focused on correct and testable implementations for the
+adjoint wave-equation and the gradient (action of the adjoint Jacobian)
+\cite{louboutin2018segeow, louboutin2020THmfi}; exactness being a mandatory
+requirement of gradient based iterative optimization algorithms.
 
@@
 We now demonstrate the scalability of
-\href{https://github.com/devitocodes/devito}{Devito} to real-world
-applications with the imaging of an industry-scale three-dimensional TTI
-subsurface model. This imaging was carried out in the Cloud on Azure and
-takes advantage of recent work to port conventional cluster code to the
-Cloud using a serverless approach. The serverless implementation is
-described in detail in \cite{witte2019TPDedas, witte2019SEGedw} and
-describes the steps to run computationnaly and financially efficient HPC
-workloads in the Cloud. This imaging project, in collaboration with
-Azure, demonstrates the scalability and robustness of
+\href{https://github.com/devitocodes/devito}{Devito} to real-world applications
+by imaging an industry-scale three-dimensional TTI subsurface model. This
+imaging was carried out in the Cloud on Azure and takes advantage of recent work
+to port conventional cluster code to the Cloud using a serverless approach. The
+serverless implementation is detailed in \cite{witte2019TPDedas,
+witte2019SEGedw} where the steps to run computationally and financially
+efficient HPC workloads in the Cloud are described. This imaging project, in
+collaboration with Azure, demonstrates the scalability and robustness of
 \href{https://github.com/devitocodes/devito}{Devito} to large scale
-wave-equation based inverse problems and its cost-effectiveness in
-combination with a serverless implementation of seismic imaging in the
-Cloud. In this example, we imaged a synthetic three-dimensional
-anisotropic subsurface model that mimics a realistic industry size
-problem with a realistic representation of the physics (TTI). The
-physical size of the problem is \texttt{10kmx10kmx2.8km} discretized on
-a \texttt{12.5m} grid with 40 points of absorbing layer on each sides
-that leads to \texttt{881x881x371} computational grid points (300
-Million grid points). The final image is the sum of 1500 single-source
-images and we computed 100 single-source images in parallel on 200 nodes
-available using two nodes per source experiment.
+wave-equation based inverse problems in combination with a cost-effective
+serverless implementation of seismic imaging in the Cloud. In this example, we
+imaged a synthetic three-dimensional anisotropic subsurface model that mimics a
+realistic industry size problem with a realistic representation of the physics
+(TTI). The physical size of the problem is \texttt{10kmx10kmx2.8km} discretized
+on a \texttt{12.5m} grid with absorbing layers of width 40 grid points on each
+side leading to \texttt{881x881x371} computational grid points ($\approx300$
+Million grid points). The final image is the sum of 1500 single-source images:
+100 single-source images were computed in parallel on the 200 nodes available
+using two nodes per source experiment.
 
@@
 
-We briefly describe the computational setup and the performance achieved
-for this anisotropic imaging problem. Due to time constraints, and
-because the resources we were given access to for this proof of concept
-with Azure were somewhat limited, we did not have access to HPC virtual
-machines (VM) nor Infiniband enabled ones. The nodes we ran this
-experiment on are \texttt{Standard\_E64\_v3} and
-\texttt{Standard\_E64s\_v3} that while not HPC VM are memory optimized
-allowing to save the wavefield in memory for imaging (TTI adjoint state
-gradient \cite{virieux, louboutin2018segeow}). These VMs are Intel®
-Broadwell E5-2673 v4 2.3GH that are dual socket, 32 physical cores (and
-hyperthreading enabled) and 432Gb of DRAM. The overall inversion
-involved computing the image for 1500 source positions, i.e.~solving
-1500 forward and 1500 adjoint TTI wave-equations. A single image
-required 600Gb of memory in single precision and we used two VMs per
-source with MPI with one rank per socket (4 MPI ranks per source) and
-imaged 100 sources in parallel. The performance achieved, in single
-precision, was as follow:
+We briefly describe the computational setup and the performance achieved for
+this anisotropic imaging problem. Due to time constraints, and because the
+resources we were given access to for this proof of concept with Azure were
+somewhat limited, we did not have access to HPC virtual machines (VM) nor
+Infiniband enabled ones. This experiment was carried out on
+\texttt{Standard\_E64\_v3} and \texttt{Standard\_E64s\_v3} nodes which, while
+not HPC VM nodes, are memory optimized thus allowing to the wavefield to be
+stored in memory for imaging (TTI adjoint state gradient \cite{virieux,
+louboutin2018segeow}). These VMs are Intel® Broadwell E5-2673 v4 2.3GH that are
+dual socket, 32 physical cores (with hyperthreading enabled) and 432Gb of DRAM.
+The overall inversion involved computing the image for 1500 source positions,
+i.e.~solving 1500 forward and 1500 adjoint TTI wave-equations. A single image
+required, in single precision, 600Gb of memory. Two VMs were used per source and
+MPI set with one rank per socket (4 MPI ranks per source) and 100 sources were
+imaged in parallel. The performance achieved was as follows:
 
@@
 
-We also observe that if we had had access to more resources, and because
-the imaging problem is embarassingly parallel over sources and can scale
-to any computational resources, we could have attempted imaging of all
-of the 1500 sources in parallel, which theoretically leads to a
-performance of 0.4PFLOP/s.
-
-\textbf{\emph{How was the performance measured}}
-
-The execution time is computed through Python-level timers prefixed by
-an MPI barrier. The floating-point operations are counted once all of
-the symbolic FLOP-reducing transformations have been performed during
-compilation. \href{https://github.com/devitocodes/devito}{Devito} uses
-an in-house estimate of cost, rather than \texttt{SymPy}'s estimate, to
-take care of some low-level intricacies. For example, Devito's estimate
-ignores the cost of integer arithmetic used for offset indexing into
-multi-dimensional arrays. To calculate the total number of FLOPs
-performed, \href{https://github.com/devitocodes/devito}{Devito}
-multiplies the floating-point operations calculated at compile time by
-the size of the iteration space, and it does that at the granularity of
-individual expressions. Thanks to aggressive code motion, the amount of
-innermost-loop-invariant sub-expressions in a
-\href{https://github.com/devitocodes/devito}{Devito} Operator is
-typically negligible, so the
-\href{https://github.com/devitocodes/devito}{Devito} estimate does not
-basically suffer from this issue, or at least not in a tangible way to
-the best of our knowledge. The Devito-reported GFLOP/s were also checked
-against those produced by Intel Advisor on several single-node
-experiments: the differences -- typically
+We comment that if more resources were available, and because the imaging
+problem is embarrassingly parallel over sources and can scale arbitrarily,the
+imaging of all of the 1500 sources in parallel could have been attempted, which
+theoretically leads to a performance of 0.4PFLOP/s.
+
+\textbf{\emph{How performance was measured}}
+
+The execution time is computed through Python-level timers prefixed by an MPI
+barrier. The floating-point operations are counted once all of the symbolic
+FLOP-reducing transformations have been performed during compilation.
+\href{https://github.com/devitocodes/devito}{Devito} uses an in-house estimate
+of cost, rather than \texttt{SymPy}'s estimate, to take care of some low-level
+intricacies. For example, Devito's estimate ignores the cost of integer
+arithmetic used for offset indexing into multi-dimensional arrays. To calculate
+the total number of FLOPs performed,
+\href{https://github.com/devitocodes/devito}{Devito} multiplies the
+floating-point operations calculated at compile time by the size of the
+iteration space, and it does that at the granularity of individual expressions.
+Thanks to aggressive code motion, the amount of innermost-loop-invariant
+sub-expressions in a \href{https://github.com/devitocodes/devito}{Devito}
+Operator is typically negligible and therefore the
+\href{https://github.com/devitocodes/devito}{Devito} estimate does not suffer
+from this issue, or at least not, to the best of our knowledge, in a tangible
+way. The Devito-reported GFLOP/s were also checked against those produced by
+Intel Advisor on several single-node experiments: the differences -- typically
 \href{https://github.com/devitocodes/devito}{Devito} underestimating the
-achieved performance -- were always at most in the order of units, and
-therefore negligible.
+achieved performance -- were always at most in the order of units, and therefore
+negligible.
 
@@
 
-The subsurface velocity model that was used in this study is an
-artificial anisotropic model that is designed and built combining two
-broadly known and used open-source SEG/EAGE acoustic velocity models
-that come each with realistic geophysical imaging challenges such as
-sub-salt imaging. The anisotropy parameters are derived from a smoothed
-version of the velocity while the tilt angles were derived from a
-combination of the smooth velocity models and vertical and horizontal
-edge detection. The final seismic image of the subsurface model is
-plotted in Figure~\ref{OverTTI} and highlights the fact that 3D seismic
-imaging based on a serverless approach and automatic code generation is
-feasible and provides good results on a realistic model.
+The subsurface velocity model used in this study is an artificial anisotropic
+model that is designed and built combining two broadly known and used
+open-source SEG/EAGE acoustic velocity models that each come with realistic
+geophysical imaging challenges such as sub-salt imaging. The anisotropy
+parameters are derived from a smoothed version of the velocity while the tilt
+angles were derived from a combination of the smooth velocity models and
+vertical and horizontal edge detection. The final seismic image of the
+subsurface model is displayed in Figure~\ref{OverTTI} and highlights the fact
+that 3D seismic imaging based on a serverless approach and automatic code
+generation is feasible and provides good results.
 
@@
 
-\cite{witte2019TPDedas} fully describes the serverless implementation
-of seismic inverse problems, including iterative algorithms for
-least-square minimization problems (LSRTM). The 3D anisotropic imaging
-results were presented as part of a keynote presentation at the EAGE HPC
-workshop in October 2019 \cite{herrmann2019EAGEHPCaii} and at the Rice
-O\&G HPC workshop \cite{witte2019RHPCssi} where we focused on the
-serverless implementation of seismic inversion algorithms in the Cloud.
-This work illustrates the flexibility and portability of Devito, as we
-were able to easily port a code only tested and developed on local
-hardware to the Cloud, with only minor adjustments. This portability
-included the possibility to run MPI-based code for domain decomposition
-in the Cloud, after developing it on a desktop computer. Our experiments
-are reproducible using the instructions in a public repository
-\href{https://github.com/slimgroup/Azure2019}{AzureTTI}, which contains,
-among the other things, the Dockerfiles and Azure
-\href{https://batch-shipyard.readthedocs.io}{batch-shipyard} setup. This
-example can also easily ber un on a traditional HPC cluster environment
-using for example \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or
-Dask \cite{dask} for parallelization over sources.
+\cite{witte2019TPDedas} describe the serverless implementation of seismic
+inverse problems in detail, including iterative algorithms for least-square
+minimization problems (LSRTM). The 3D anisotropic imaging results were presented
+as part of a keynote presentation at the EAGE HPC workshop in October 2019
+\cite{herrmann2019EAGEHPCaii} and at the Rice O\&G HPC workshop
+\cite{witte2019RHPCssi} in which the focus was on the serverless implementation
+of seismic inversion algorithms in the Cloud. This work illustrates the
+flexibility and portability of Devito: we were able to easily port a code
+developed and tested on local hardware to the Cloud, with only minor
+adjustments. Further, note that this experiment included the porting of
+MPI-based code for domain decomposition developed on desktop computers to the
+Cloud. Our experiments are reproducible using the instructions in a public
+repository \href{https://github.com/slimgroup/Azure2019}{AzureTTI}, which
+contains, among the other things, the Dockerfiles and Azure
+\href{https://batch-shipyard.readthedocs.io}{batch-shipyard} setup. This example
+can also be easily run on a traditional HPC cluster environment using, for
+example, \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or Dask~\cite{dask}
+for parallelization over sources.
 

### 2020-04-20 paper.tex
Elastic section update.

@@
 
-While the subsurface image in the previous Section is obtained with
-anisotropic propagators capable of mimicking the real physics, in order
-to model both the kinematics and the amplitudes correctly elastic
-propagators are required. These propagators are for example extremely
-important for global seismology, as the Shear surface waves (component
-ignored in TTI) are the most hazardous ones. In this section, we exploit
+While the subsurface image obtained in section \ref{d-imaging-example-on-azure}
+utilized anisotropic propagators capable of mimicking intricate physics, in order
+to model both the wave kinematics and amplitudes correctly, elastic
+propagators are required. These propagators are, for example, extremely
+important in global seismology since shear surface waves (which are
+ignored in acoustic models) are the most hazardous. In this section, we exploit
 the tensor algebra language introduced in
 \href{https://github.com/devitocodes/devito}{Devito} v4.0 to express
-elastic waves in a compact and elegant notation.
-
-The elastic isotropic wave-equation, parametrized by the Lamé parameters
+an elastic model with compact and elegant notation.
+
+The isotropic elastic wave-equation, parametrized by the so-called Lamé parameters
 $\lambda, \mu$ and the density $\rho$ reads:
@@
 %
- The discretization of such a set of coupled PDEs requires five
+ The discretization of \eqref{elas1} and \eqref{stress} requires five
 equations in two dimensions (two equations for the particle velocity and
-three for stress) and nine equations in three dimensions (three particle
-velocities and six stress equations). However, the mathematical
-definition only requires two coupled vector/tensor-valued equations for
+three for the stress) and nine equations in three dimensions (three for velocity
+and six for stress). However, the mathematical
+definition only require two coupled vector/tensor-valued equations for
 any number of dimensions.
@@
 \href{https://github.com/devitocodes/devito}{Devito} language with
-tensorial objects to enable a straightforward -- and at the same time
-mathematically rigorous -- definition of high-dimensional PDEs, such as
-the elastic wave-equation in Equation~\ref{elas1}. This effort was
+tensorial objects to enable straightforward and
+mathematically rigorous definitions of high-dimensional PDEs, such as
+the elastic wave-equation defined in \eqref{elas1}. This implementation was
 inspired by \cite{ufl}, a functional language for finite element
@@
 The extended \href{https://github.com/devitocodes/devito}{Devito}
-language introduces two new types, \texttt{VectorFunction} (and
-\texttt{VectorTimeFunction}) for vectorial objects such as the particle
-velocity, and \texttt{TensorFunction} (and \texttt{TensorTimeFunction})
+language introduces two new types, \texttt{VectorFunction}/
+\texttt{VectorTimeFunction} for vectorial objects such as the particle
+velocity, and \texttt{TensorFunction}/\texttt{TensorTimeFunction}
 for second-order tensor objects (matrices) such as the stress. These new
-objects are constructed the exact same way as the scalar
+objects are constructed in the same manner as scalar
 \texttt{Function} objects. They also automatically implement staggered
@@
 
-With this extended language, the elastic wave-equation can be defined in
-only four lines as follows:
+With this extended language, the elastic wave-equation defined in
+\eqref{elas1} and \eqref{stress} can be expressed in only four lines
+of code:
 
@@
 Figure~\ref{PrettyElas}. This representation reflects perfectly the
-mathematics while still providing computational portability and
+mathematics while maintaining computational portability and
 efficiency through the
@@
 \includegraphics[width=1.000\hsize]{./Figures/vel_symb.png}
-\caption{Update stencil for the particle velocity. The stencil for
-updating the stress component is left out for readability, as the
-equation does not fit onto a single page. However, it can be found in
+\caption{Particle velocity update stencil. The stress update
+is omitted for readability (the
+equation does not fit into a single page). However, it can be found in
 the \href{https://github.com/devitocodes/devito}{Devito} tutorial on
@@
 
-We show the elastic particle velocity and stress for a broadly
-recognized 2D synthetic model, the elastic
+To demonstrate the efficacy of the elastic implementation outlined
+above we utilized a broadly recognized 2D synthetic model, the elastic
 Marmousi-ii\cite{versteeg927, marmouelas} model. The wavefields are
 shown on Figure~\ref{ElasWf} and its corresponding elastic shot records
-are displayed in Figure~\ref{ElasShot}. Thes two figures show that the
-wavefield is as expected purely acoustic in the water layer
-(\texttt{t\_\{xy\}=0}) and transition perfectly at the ocean bottom to
-an elastic wavefield. We can also clearly see the Shear wave wavefront
-in the subsurface (around 1km depth). Thes two plots show that we
-effectively solve the elastic wave-equation with a high-level
-mathematical definition of the PDE that abstracts away all the
-computational technicalities such as staggered grid analysis. The shot
-records in Figure~\ref{ElasShot} also match the original data generated
-by the creator of the elastic model in \cite{marmouelas}.
+are displayed in Figure~\ref{ElasShot}. These two figures show that the
+wavefield is, as expected, purely acoustic in the water layer
+(\texttt{t\_\{xy\}=0}) and transitions correctly at the ocean bottom to
+an elastic wavefield. We can also clearly see the shear wave-front
+in the subsurface (at a depth of approximately 1km). Figures \ref{ElasWf}
+and \ref{ElasShot} demonstrate that this high-level
+\href{https://github.com/devitocodes/devito}{Devito} implementation
+of the elastic wave-equation is effective and accurate.
+Importantly, in constructing this model within the
+\href{https://github.com/devitocodes/devito}{Devito} DSL framework,
+computational technicalities such as the staggered grid analysis are
+abstracted away. We note that the shot records displayed in Figure~\ref{ElasShot}
+match the original data generated by the creator of this elastic model
+(\cite{marmouelas}).
 
@@
 \caption{Particle velocities and stress at time $t=3\text{s}$ for a
-source at 10m depth and \texttt{x=5\textbackslash{}text\{km\}} in the
+10m deep source and \texttt{x=5\textbackslash{}text\{km\}} in the
 marmousi-ii model.}\label{ElasWf}
@@
 pressure (trace of stress tensor) at the surface (5m depth), \texttt{b}
-is the vertical particle velocity and \texttt{c} is the horizontal
+and \texttt{c} display, respectively, the vertical and horizontal
 particle velocity at the ocean bottom (450m depth).}\label{ElasShot}
@@
 
-Finally, we modelled three dimensional elastic data in the Cloud to
+Finally, three dimensional elastic data was modelled in the Cloud to
 demonstrate the scalability of
 \href{https://github.com/devitocodes/devito}{Devito} to cluster-size
-problems in the Cloud. The model we chose mimics the reference model in
-geophysics known as the SEAM model \cite{fehler2011seam} that is a
-three dimensional extreme-scale synthetic representation of the
-subsurface. The physical dimension of the model are
-\texttt{45kmx35kmx15km} then discretized with a grid spacing of
-\texttt{20mx20mx10m} that led to a computational grid of
+problems. The model used in this experiment mimics a reference model in
+geophysics known as the SEAM model \cite{fehler2011seam}, a
+three dimensional extreme-scale synthetic representation of a
+subsurface. The physical dimensions of the model are
+\texttt{45kmx35kmx15km} discretized with a grid spacing of
+\texttt{20mx20mx10m} leading to a computational grid of
 \texttt{2250x1750x1500} grid points (5.9 billion grid points). One of
-the main challenges of elastic modelling is the extreme memory cost due
-to the number of wavefield. For a three dimensional propagator, a
-minimum of 21 fields need to be stored:
+the main challenges of elastic modelling is the extreme memory cost owing
+to the number of wavefields (a minimum of 21 fields in a three dimensional
+propagator) that need to be stored:
 
@@
 
-These 21 fields, with the grid we just describe, leads to a minimum of
-461Gb of memory for modelling only. For this experiment, we obtained
-access to small HPC VM on azure called \texttt{Standard\_H16r} that are
-16 cores Intel Xeon E5 2667 v3 with no hyperthreading and used 32 nodes
-for a single source experiment (we solved a single wave-equation). We
-used a 12th order discretization that leads to 2.8TFLOP/time-step to be
-computed for this model and propagated the elastic wave for 16 seconds
-(23000 time steps). The modelling ran in 16 hours which converts to
-1.1TFLOP/s. While these number may seem low, the elastic kernel is
-extremely memory bound, while the TTI kernel is nearly compute bound
-(see rooflines in \cite{louboutin2016ppf, devito-api, devito-compiler})
-making it more computationally efficient, in particular in combination
-with MPI. Future work with involve working on the InfiniBand enabled and
-true HPC VM on azure to achieve Cloud performance on par with state of
-the art Cluster performance. From the performance obtained in this
+These 21 fields, for the 5.9 billion point grid defined above, lead to a
+minimum memory requirement of 461Gb for modelling alone. For this experiment,
+access was obtained for small HPC VMs (on azure) called \texttt{Standard\_H16r}.
+These VMs contain 16 core Intel Xeon E5 2667 v3 chips, with no hyperthreading,
+and 32 nodes were used for a single source experiment
+(i.e. a single wave-equation was solved). We
+used a 12th order discretization in space that led to 2.8TFLOP/time-step being
+computed by this model and the elastic wave was propagated for 16 seconds
+(23000 time steps). Completion of this modelling run took 16 hours, converting to
+1.1TFLOP/s. While these numbers may appear low, it should be noted that
+the elastic kernel is extremely memory bound, while the TTI kernel is nearly
+compute bound (see rooflines in \cite{louboutin2016ppf, devito-api, devito-compiler})
+making it more computationally efficient, particularly in combination
+with MPI. Future work will involve working on InfiniBand enabled and
+true HPC VMs on azure to achieve Cloud performance on par with that of state of
+the art HPC clusters. Extrapolating from the performance obtained in this
 experiment, and assuming a fairly standard setup of 5000 independent
-source experiments, computing a synthetic dataset would require 322
-EFLOP to be computed (23k time-steps x 2.8TFLOP/time-step x 500
-sources), or using the full scalability and computing all sources in
-parallel would lead to 5.5PFLOP/s from our single source experiment
-result.
+source experiments, computing an elastic synthetic dataset would require 322
+EFLOPs (23k time-steps x 2.8TFLOP/time-step x 500
+sources), or utilizing the full scalability and computing all sources in
+parallel this becomes 5.5PFLOP/s.
 

### 2020-04-21 paper.tex
minor polishing

@@
 finite-difference method. The language is by design flexible enough to
-express other types of non finite-diffrence operators, such as
+express other types of non finite-difference operators, such as
 interpolation and tensor contractions, that are inherent to
@@
 \item
-  \texttt{Sparse(Time)Function} a represention of (time-varying) point
+  \texttt{Sparse(Time)Function} a representation of (time-varying) point
   objects on a \texttt{Grid} object, generally unaligned with respect to
@@
 The source term is handled separately from the PDE as a pointwise
-operation called \texttt{injection}, while the measurement is handled
+operation called injection, while the measurement is handled
 with an interpolation. By default,

### 2020-04-21 paper.tex
Add tti-so8-unoptimized.c to the text

@@
 terms of FLOPs and memory usage. For certain workloads, limited domain
-decomposition, that balances the domain size and the number of independent
-experiments, and checkpointing techniques must to be adopted. In the following
+decomposition, which balances the domain size and the number of independent
+experiments, as well as checkpointing techniques must be adopted. In the following
 sections, we describe an industry-scale seismic imaging problem that poses all
@@
 at \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
-in \texttt{tti-so8.c} and \texttt{tti-so8-mpi.c}.
+in \texttt{tti-so8-unoptimized.c}, \texttt{tti-so8.c} and \texttt{tti-so8-mpi.c}.
 
 Owing to the very high number of floating-point operations (FLOPs) needed per
-grid point for the weighted rotated Laplacian, this anisotropic wave-equation is
-extremely challenging to implement. As we show in Table Table~\ref{ttiFLOPs},
-and previously analysed in \cite{louboutin2016ppf}, the computational cost with
+grid point for the weighted rotated Laplacian, this anisotropic wave-equation
+is extremely challenging to implement. As we show in Table~\ref{ttiFLOPs}, and
+previously analysed in \cite{louboutin2016ppf}, the computational cost with
 high-order finite-difference is in the order of thousands of FLOPs per grid
 point without optimizations. The version without FLOP-reducing optimizations is
-a direct translation of the discretized operators into stencil expressions. The
-version with optimizations employs transformations such as common
-sub-expressions elimination, factorization, and cross-iteration redundancy
-elimination -- the latter being key in removing redundancies introduced by mixed
-derivatives. Implementing all of these techniques manually is inherently
-difficult and laborious. Further, to obtain the desired performance improvements
-it is necessary to orchestrate them with aggressive loop fusion (for data
-locality), tiling (for data locality and tensor temporaries), and potentially
-ad-hoc vectorization strategies (if rotating registers are used). While an
-explanation of the optimization strategies employed by
-\href{https://github.com/devitocodes/devito}{Devito} is beyond the scope of this
-paper (see \cite{devito-compiler} for details), what is emphasized here is that
-users can easily take full advantage of these optimization without needed to
-concern themselves with the details.
+a direct translation of the discretized operators into stencil expressions (see
+\texttt{tti-so8-unoptimized.c}). The version with optimizations employs
+transformations such as common sub-expressions elimination, factorization, and
+cross-iteration redundancy elimination -- the latter being key in removing
+redundancies introduced by mixed derivatives. Implementing all of these
+techniques manually is inherently difficult and laborious. Further, to obtain
+the desired performance improvements it is necessary to orchestrate them with
+aggressive loop fusion (for data locality), tiling (for data locality and
+tensor temporaries), and potentially ad-hoc vectorization strategies (if
+rotating registers are used). While an explanation of the optimization
+strategies employed by \href{https://github.com/devitocodes/devito}{Devito} is
+beyond the scope of this paper (see \cite{devito-compiler} for details), what
+is emphasized here is that users can easily take full advantage of these
+optimization without needed to concern themselves with the details.
 

### 2020-04-21 paper.tex
Drop hyperlinks to Devito repo

@@
 
-It is with these specifications in mind that
-\href{https://github.com/devitocodes/devito}{Devito}, a symbolic domain
+It is with these specifications in mind that Devito, a symbolic domain
 specific language (DSL) and compiler for the automatic generation of
 finite-difference stencils, has been developed. Originally deigned to
-accelerate research and development in exploration geophysics, the high-level
-interface, previously described in detail in \cite{devito-api}, is built on top
-of \texttt{SymPy} \cite{sympy} and is inspired by the underlying mathematical
-formulations and other DSLs such as FEniCS \cite{fenics} and Firedrake
-\cite{firedrake}. This interface allows the user to formulate wave-equations,
-and more generally time-dependent PDEs in a simple and mathematically coherent
-way. The Devito compiler then automatically generates finite-difference
-stencils from these mathematical expressions. One of the main advantages of
-\href{https://github.com/devitocodes/devito}{Devito} over other
-finite-difference DSLs is that generic expressions such as sparse operations
-(i.e.~point source injection or localized measurements) are fully supported and
-expressible in a high-level fashion. The second component of
-\href{https://github.com/devitocodes/devito}{Devito} is its compiler (c.f
-\cite{devito-compiler}) that generates highly optimized C code. The generated
-code is then compiled at runtime for the hardware at hand.
+accelerate research and development in exploration geophysics, the
+high-level interface, previously described in detail in \cite{devito-api},
+is built on top of \texttt{SymPy} \cite{sympy} and is inspired by the
+underlying mathematical formulations and other DSLs such as FEniCS
+\cite{fenics} and Firedrake \cite{firedrake}. This interface allows the
+user to formulate wave-equations, and more generally time-dependent PDEs
+in a simple and mathematically coherent way. The Devito compiler then
+automatically generates finite-difference stencils from these mathematical
+expressions. One of the main advantages of Devito over other
+finite-difference DSLs is that generic expressions such as sparse
+operations (i.e.~point source injection or localized measurements) are
+fully supported and expressible in a high-level fashion. The second
+component of Devito is its compiler (c.f \cite{devito-compiler}) that
+generates highly optimized C code. The generated code is then compiled at
+runtime for the hardware at hand.
 
@@
 application and use cases of Devito. Here, we present a series of
-extensions and applications to large-scale three-dimensional problem
-sizes as encountered in exploration geophysics. These experiments are
-carried out in Cloud-based HPC systems and include elastic forward
-modelling using distributed-memory parallelism and imaging based on the
-tilted transverse isotropic (TTI) wave-equation
-(\cite{virieux, thomsen1986, zhang-tti, duveneck, louboutin2018segeow}).
-These proof of concepts highlight two critical features: first, the
-ability of the symbolic interface and the compiler to translate to
-large-scale adjoint-based inversion problems that require massive
-compute (since thousands of PDEs are solved) as well as large amounts of
-memory (since the adjoint state method requires the forward model to be
-saved in memory). Secondly, through the elastic modelling example, we
-demonstrate that \href{https://github.com/devitocodes/devito}{Devito}
-now fully supports and automates vectorial and second order tensorial
-staggered-grid finite-differences with the same high-level API
-previously presented for scalar fields defined on cartesian grids.
-
-This paper is organized as follows: first, we provide a brief overview
-of \href{https://github.com/devitocodes/devito}{Devito} and its symbolic
-API and present the distributed memory implementation that allows
-large-scale modeling and inversion by means of domain decomposition. We
-then provide a brief comparison with a state of the art hand-coded wave
-propagator to validate the performance previously benchmarked with the
-roofline model
-(\cite{patterson, devito-compiler, devito-api, louboutin2016ppf}).
-Before concluding, results from the Cloud-based experiments discussed
-above are presented, highlighting the vectorial and tensorial
-capabilities of Devito.
+extensions and applications to large-scale three-dimensional problem sizes
+as encountered in exploration geophysics. These experiments are carried
+out in Cloud-based HPC systems and include elastic forward modelling using
+distributed-memory parallelism and imaging based on the tilted transverse
+isotropic (TTI) wave-equation (\cite{virieux, thomsen1986, zhang-tti,
+duveneck, louboutin2018segeow}).  These proof of concepts highlight two
+critical features: first, the ability of the symbolic interface and the
+compiler to translate to large-scale adjoint-based inversion problems that
+require massive compute (since thousands of PDEs are solved) as well as
+large amounts of memory (since the adjoint state method requires the
+forward model to be saved in memory). Secondly, through the elastic
+modelling example, we demonstrate that Devito now fully supports and
+automates vectorial and second order tensorial staggered-grid
+finite-differences with the same high-level API previously presented for
+scalar fields defined on cartesian grids.
+
+This paper is organized as follows: first, we provide a brief overview of
+Devito and its symbolic API and present the distributed memory
+implementation that allows large-scale modeling and inversion by means of
+domain decomposition. We then provide a brief comparison with a state of
+the art hand-coded wave propagator to validate the performance previously
+benchmarked with the roofline model (\cite{patterson, devito-compiler,
+devito-api, louboutin2016ppf}).  Before concluding, results from the
+Cloud-based experiments discussed above are presented, highlighting the
+vectorial and tensorial capabilities of Devito.
 
@@
 
-It is up to the \href{https://github.com/devitocodes/devito}{Devito}
-compiler to translate the symbolic specification into C code. The
-lowering of the input language down to C consists of several compilation
-passes, some of which introducing performance optimizations that are the
-key to fast code. Next to classic stencil optimizations (e.g., cache
-blocking, alignment, SIMD and OpenMP parallelism),
-\href{https://github.com/devitocodes/devito}{Devito} applies a series of
-FLOP-reducing transformations as well as aggressive loop fusion. For a
-complete treatment, the interested reader should refer to
-\cite{devito-compiler}.
+It is up to the Devito compiler to translate the symbolic specification
+into C code. The lowering of the input language down to C consists of
+several compilation passes, some of which introducing performance
+optimizations that are the key to fast code. Next to classic stencil
+optimizations (e.g., cache blocking, alignment, SIMD and OpenMP
+parallelism), Devito applies a series of FLOP-reducing transformations as
+well as aggressive loop fusion. For a complete treatment, the interested
+reader should refer to \cite{devito-compiler}.
 
@@
 
-In this section we illustrate the
-\href{https://github.com/devitocodes/devito}{Devito} language showing
-how to implement the acoustic wave-equation in isotropic media
+In this section we illustrate the Devito language showing how to implement
+the acoustic wave-equation in isotropic media
 %
@@
 %
- The core of the \href{https://github.com/devitocodes/devito}{Devito}
+ The core of the Devito
 symbolic API consists of three classes:
@@
 The source term is handled separately from the PDE as a pointwise
-operation called injection, while the measurement is handled
-with an interpolation. By default,
-\href{https://github.com/devitocodes/devito}{Devito} initializes all
-\texttt{Function} data to 0, and thus automatically satisfies the zero
-Dirichlet condition at \texttt{t=0}. The isotropic acoustic
-wave-equation can then be implemented in
-\href{https://github.com/devitocodes/devito}{Devito} as below
+operation called injection, while the measurement is handled with an
+interpolation. By default, Devito initializes all \texttt{Function} data
+to 0, and thus automatically satisfies the zero Dirichlet condition at
+\texttt{t=0}. The isotropic acoustic wave-equation can then be implemented
+in Devito as below
 
@@
 
-The output of the \href{https://github.com/devitocodes/devito}{Devito}
-compiler for the running example used in this section is available at
+The output of the Devito compiler for the running example used in this
+section is available at
 \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
@@
 
-Devito implements distributed-memory parallelism on top of MPI. The
-design is such that users can almost entirely abstract away from it and
-reuse non-distributed code as is. Given any
-\href{https://github.com/devitocodes/devito}{Devito} code, just running
-it as
+Devito implements distributed-memory parallelism on top of MPI. The design
+is such that users can almost entirely abstract away from it and reuse
+non-distributed code as is. Given any Devito code, just running it as
 
@@
 
-The domain decomposition occurs in Python upon creation of a
-\texttt{Grid} object. Exploiting the MPI Cartesian topology abstraction,
-\href{https://github.com/devitocodes/devito}{Devito} logically splits a
-grid based on the number of available MPI processes (users are given an
-``escape hatch'' to override Devito's default decomposition strategy).
-\texttt{Function} and \texttt{TimeFunction} objects inherit the
-\texttt{Grid} decomposition. For \texttt{SparseFunction} objects the
-approach is different. Since a \texttt{SparseFunction} represents a
-sparse set of points,
-\href{https://github.com/devitocodes/devito}{Devito} looks at the
-physical coordinates of each point and, based on the \texttt{Grid}
-decomposition, schedules the logical ownership to an MPI rank. If a
-sparse point lies along the boundary of two or more MPI ranks, then it
-is duplicated in each of these ranks to be accessible by all neighboring
+The domain decomposition occurs in Python upon creation of a \texttt{Grid}
+object. Exploiting the MPI Cartesian topology abstraction, Devito
+logically splits a grid based on the number of available MPI processes
+(users are given an ``escape hatch'' to override Devito's default
+decomposition strategy).  \texttt{Function} and \texttt{TimeFunction}
+objects inherit the \texttt{Grid} decomposition. For
+\texttt{SparseFunction} objects the approach is different. Since a
+\texttt{SparseFunction} represents a sparse set of points, Devito looks at
+the physical coordinates of each point and, based on the \texttt{Grid}
+decomposition, schedules the logical ownership to an MPI rank. If a sparse
+point lies along the boundary of two or more MPI ranks, then it is
+duplicated in each of these ranks to be accessible by all neighboring
 processes. Eventually, a duplicated point may be redundantly computed by
@@
 
-When accessing or manipulating data in a
-\href{https://github.com/devitocodes/devito}{Devito} code, users have
-the illusion to be working with classic NumPy arrays, while underneath
-they are actually distributed. All manner of NumPy indexing schemes
-(basic, slicing, etc.) are supported. In the implementation, proper
-global-to-local and local-to-global index conversion routines are used
-to propagate a read/write access to the impacted subset of ranks. For
+When accessing or manipulating data in a Devito code, users have the
+illusion to be working with classic NumPy arrays, while underneath they
+are actually distributed. All manner of NumPy indexing schemes (basic,
+slicing, etc.) are supported. In the implementation, proper
+global-to-local and local-to-global index conversion routines are used to
+propagate a read/write access to the impacted subset of ranks. For
 example, consider the array
@@
 Finally, we remark that while providing abstractions for distributed
-data manipulation, \href{https://github.com/devitocodes/devito}{Devito}
+data manipulation, Devito
 does not support natively any mechanisms for parallel I/O. The
@@
 %
- We enable the simple expression of these complicated stencils in
-\href{https://github.com/devitocodes/devito}{Devito} as
-finite-difference shortcuts such as \texttt{u.dx} where \texttt{u}
-is a \texttt{Function}. Such shortcuts are enabled not only for the
-basic types but for generic composite expressions, for example
+
+We enable the simple expression of these complicated stencils in Devito
+as finite-difference shortcuts such as \texttt{u.dx} where \texttt{u} is
+a \texttt{Function}. Such shortcuts are enabled not only for the basic
+types but for generic composite expressions, for example
 \texttt{(u\phantom{\ }+\phantom{\ }v.dx).dy}. As a consequence, the
-rotated derivative defined in~\ref{rot} is implemented with
-\href{https://github.com/devitocodes/devito}{Devito} in two lines as:
+rotated derivative defined in~\ref{rot} is implemented with Devito in two
+lines as:
 
@@
 rotating registers are used). While an explanation of the optimization
-strategies employed by \href{https://github.com/devitocodes/devito}{Devito} is
-beyond the scope of this paper (see \cite{devito-compiler} for details), what
-is emphasized here is that users can easily take full advantage of these
-optimization without needed to concern themselves with the details.
+strategies employed by Devito is beyond the scope of this paper (see
+\cite{devito-compiler} for details), what is emphasized here is that users
+can easily take full advantage of these optimization without needed to
+concern themselves with the details.
 
@@
 latest geophysical insights in tandem with those from high-performance
-computing. With \href{https://github.com/devitocodes/devito}{Devito} on the
-other hand, both the forward and adjoint equations can be implemented in a few
-lines of Python code as illustrated with the rotated operator in \#rotxpy. The
-low level optimization element of the development is then taken care of under
-the hood by the \href{https://github.com/devitocodes/devito}{Devito} compiler.
+computing. With Devito on the other hand, both the forward and adjoint
+equations can be implemented in a few lines of Python code as illustrated
+with the rotated operator in \#rotxpy. The low level optimization element
+of the development is then taken care of under the hood by the Devito
+compiler.
 
@@
 
-We now demonstrate the scalability of
-\href{https://github.com/devitocodes/devito}{Devito} to real-world applications
-by imaging an industry-scale three-dimensional TTI subsurface model. This
-imaging was carried out in the Cloud on Azure and takes advantage of recent work
-to port conventional cluster code to the Cloud using a serverless approach. The
-serverless implementation is detailed in \cite{witte2019TPDedas,
-witte2019SEGedw} where the steps to run computationally and financially
-efficient HPC workloads in the Cloud are described. This imaging project, in
-collaboration with Azure, demonstrates the scalability and robustness of
-\href{https://github.com/devitocodes/devito}{Devito} to large scale
-wave-equation based inverse problems in combination with a cost-effective
-serverless implementation of seismic imaging in the Cloud. In this example, we
-imaged a synthetic three-dimensional anisotropic subsurface model that mimics a
-realistic industry size problem with a realistic representation of the physics
-(TTI). The physical size of the problem is \texttt{10kmx10kmx2.8km} discretized
-on a \texttt{12.5m} grid with absorbing layers of width 40 grid points on each
-side leading to \texttt{881x881x371} computational grid points ($\approx300$
-Million grid points). The final image is the sum of 1500 single-source images:
-100 single-source images were computed in parallel on the 200 nodes available
+We now demonstrate the scalability of Devito to real-world applications by
+imaging an industry-scale three-dimensional TTI subsurface model. This
+imaging was carried out in the Cloud on Azure and takes advantage of
+recent work to port conventional cluster code to the Cloud using a
+serverless approach. The serverless implementation is detailed in
+\cite{witte2019TPDedas, witte2019SEGedw} where the steps to run
+computationally and financially efficient HPC workloads in the Cloud are
+described. This imaging project, in collaboration with Azure, demonstrates
+the scalability and robustness of Devito to large scale wave-equation
+based inverse problems in combination with a cost-effective serverless
+implementation of seismic imaging in the Cloud. In this example, we imaged
+a synthetic three-dimensional anisotropic subsurface model that mimics a
+realistic industry size problem with a realistic representation of the
+physics (TTI). The physical size of the problem is
+\texttt{10kmx10kmx2.8km} discretized on a \texttt{12.5m} grid with
+absorbing layers of width 40 grid points on each side leading to
+\texttt{881x881x371} computational grid points ($\approx300$ Million grid
+points). The final image is the sum of 1500 single-source images: 100
+single-source images were computed in parallel on the 200 nodes available
 using two nodes per source experiment.
@@
 
-The execution time is computed through Python-level timers prefixed by an MPI
-barrier. The floating-point operations are counted once all of the symbolic
-FLOP-reducing transformations have been performed during compilation.
-\href{https://github.com/devitocodes/devito}{Devito} uses an in-house estimate
-of cost, rather than \texttt{SymPy}'s estimate, to take care of some low-level
-intricacies. For example, Devito's estimate ignores the cost of integer
-arithmetic used for offset indexing into multi-dimensional arrays. To calculate
-the total number of FLOPs performed,
-\href{https://github.com/devitocodes/devito}{Devito} multiplies the
-floating-point operations calculated at compile time by the size of the
-iteration space, and it does that at the granularity of individual expressions.
-Thanks to aggressive code motion, the amount of innermost-loop-invariant
-sub-expressions in a \href{https://github.com/devitocodes/devito}{Devito}
-Operator is typically negligible and therefore the
-\href{https://github.com/devitocodes/devito}{Devito} estimate does not suffer
-from this issue, or at least not, to the best of our knowledge, in a tangible
-way. The Devito-reported GFLOP/s were also checked against those produced by
-Intel Advisor on several single-node experiments: the differences -- typically
-\href{https://github.com/devitocodes/devito}{Devito} underestimating the
-achieved performance -- were always at most in the order of units, and therefore
-negligible.
+The execution time is computed through Python-level timers prefixed by an
+MPI barrier. The floating-point operations are counted once all of the
+symbolic FLOP-reducing transformations have been performed during
+compilation.  Devito uses an in-house estimate of cost, rather than
+\texttt{SymPy}'s estimate, to take care of some low-level intricacies. For
+example, Devito's estimate ignores the cost of integer arithmetic used for
+offset indexing into multi-dimensional arrays. To calculate the total
+number of FLOPs performed, Devito multiplies the floating-point operations
+calculated at compile time by the size of the iteration space, and it does
+that at the granularity of individual expressions.  Thanks to aggressive
+code motion, the amount of innermost-loop-invariant sub-expressions in an
+\texttt{Operator} is typically negligible and therefore the Devito estimate
+does not suffer from this issue, or at least not, to the best of our
+knowledge, in a tangible way. The Devito-reported GFLOP/s were also
+checked against those produced by Intel Advisor on several single-node
+experiments: the differences -- typically Devito underestimating the
+achieved performance -- were always at most in the order of units, and
+therefore negligible.
 
@@
 important in global seismology since shear surface waves (which are
-ignored in acoustic models) are the most hazardous. In this section, we exploit
-the tensor algebra language introduced in
-\href{https://github.com/devitocodes/devito}{Devito} v4.0 to express
+ignored in acoustic models) are the most hazardous. In this section, we
+exploit the tensor algebra language introduced in Devito v4.0 to express
 an elastic model with compact and elegant notation.
@@
 
-We have augmented the
-\href{https://github.com/devitocodes/devito}{Devito} language with
-tensorial objects to enable straightforward and
-mathematically rigorous definitions of high-dimensional PDEs, such as
-the elastic wave-equation defined in \eqref{elas1}. This implementation was
-inspired by \cite{ufl}, a functional language for finite element
-methods.
-
-The extended \href{https://github.com/devitocodes/devito}{Devito}
-language introduces two new types, \texttt{VectorFunction}/
-\texttt{VectorTimeFunction} for vectorial objects such as the particle
-velocity, and \texttt{TensorFunction}/\texttt{TensorTimeFunction}
-for second-order tensor objects (matrices) such as the stress. These new
-objects are constructed in the same manner as scalar
-\texttt{Function} objects. They also automatically implement staggered
-grid and staggered finite-differences with the possibility of half-node
-averaging. Each component of a tensorial object -- a (scalar)
-\href{https://github.com/devitocodes/devito}{Devito} \texttt{Function}
--- is accessible via conventional vector notation (i.e.
+We have augmented the Devito language with tensorial objects to enable
+straightforward and mathematically rigorous definitions of
+high-dimensional PDEs, such as the elastic wave-equation defined in
+\eqref{elas1}. This implementation was inspired by \cite{ufl}, a
+functional language for finite element methods.
+
+The extended Devito language introduces two new types,
+\texttt{VectorFunction}/ \texttt{VectorTimeFunction} for vectorial objects
+such as the particle velocity, and
+\texttt{TensorFunction}/\texttt{TensorTimeFunction} for second-order
+tensor objects (matrices) such as the stress. These new objects are
+constructed in the same manner as scalar \texttt{Function} objects. They
+also automatically implement staggered grid and staggered
+finite-differences with the possibility of half-node averaging. Each
+component of a tensorial object -- a (scalar) Devito \texttt{Function} --
+is accessible via conventional vector notation (i.e.
 \texttt{v{[}0{]},\phantom{\ }t{[}0,1{]},....}).
@@
 
-The \texttt{SymPy} expressions created by these commands can be
-displayed with \texttt{sympy.pprint} as shown in
-Figure~\ref{PrettyElas}. This representation reflects perfectly the
-mathematics while maintaining computational portability and
-efficiency through the
-\href{https://github.com/devitocodes/devito}{Devito} compiler. The
-complete generated code for the elastic wave-equation with and without
-MPI is made available at
+The \texttt{SymPy} expressions created by these commands can be displayed
+with \texttt{sympy.pprint} as shown in Figure~\ref{PrettyElas}. This
+representation reflects perfectly the mathematics while maintaining
+computational portability and efficiency through the Devito compiler. The
+complete generated code for the elastic wave-equation with and without MPI
+is made available at
 \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
@@
 \caption{Particle velocity update stencil. The stress update
-is omitted for readability (the
-equation does not fit into a single page). However, it can be found in
-the \href{https://github.com/devitocodes/devito}{Devito} tutorial on
-elastic modelling on github.}\label{PrettyElas}
+is omitted for readability (the equation does not fit into a single page).
+However, it can be found in the Devito tutorial on elastic modelling
+on GitHub.}\label{PrettyElas}
 \end{figure*}
@@
 in the subsurface (at a depth of approximately 1km). Figures \ref{ElasWf}
-and \ref{ElasShot} demonstrate that this high-level
-\href{https://github.com/devitocodes/devito}{Devito} implementation
-of the elastic wave-equation is effective and accurate.
-Importantly, in constructing this model within the
-\href{https://github.com/devitocodes/devito}{Devito} DSL framework,
-computational technicalities such as the staggered grid analysis are
-abstracted away. We note that the shot records displayed in Figure~\ref{ElasShot}
-match the original data generated by the creator of this elastic model
+and \ref{ElasShot} demonstrate that this high-level Devito implementation
+of the elastic wave-equation is effective and accurate.  Importantly, in
+constructing this model within the Devito DSL framework, computational
+technicalities such as the staggered grid analysis are abstracted away. We
+note that the shot records displayed in Figure~\ref{ElasShot} match the
+original data generated by the creator of this elastic model
 (\cite{marmouelas}).
@@
 Finally, three dimensional elastic data was modelled in the Cloud to
-demonstrate the scalability of
-\href{https://github.com/devitocodes/devito}{Devito} to cluster-size
-problems. The model used in this experiment mimics a reference model in
-geophysics known as the SEAM model \cite{fehler2011seam}, a
-three dimensional extreme-scale synthetic representation of a
-subsurface. The physical dimensions of the model are
-\texttt{45kmx35kmx15km} discretized with a grid spacing of
+demonstrate the scalability of Devito to cluster-size problems. The model
+used in this experiment mimics a reference model in geophysics known as
+the SEAM model \cite{fehler2011seam}, a three dimensional extreme-scale
+synthetic representation of a subsurface. The physical dimensions of the
+model are \texttt{45kmx35kmx15km} discretized with a grid spacing of
 \texttt{20mx20mx10m} leading to a computational grid of
-\texttt{2250x1750x1500} grid points (5.9 billion grid points). One of
-the main challenges of elastic modelling is the extreme memory cost owing
-to the number of wavefields (a minimum of 21 fields in a three dimensional
+\texttt{2250x1750x1500} grid points (5.9 billion grid points). One of the
+main challenges of elastic modelling is the extreme memory cost owing to
+the number of wavefields (a minimum of 21 fields in a three dimensional
 propagator) that need to be stored:
@@
 Because earlier performance benchmarks focus mainly on roofline model
-analysis, we also compared the runtime of
-\href{https://github.com/devitocodes/devito}{Devito} with a reference
+analysis, we also compared the runtime of Devito with a reference
 open source hand-coded propagator in collaboration with its author for
@@
 
-The runtime obtained for this problem for both propagators were
-identical with less than one percent of difference. This similar runtime
-were obtained both with the Intel compiler and the GNU compiler and we
-ran the experiment with a fourth order and a sixth order discretization.
-The kernels were executed five time each to ensure consistency between
-the results and we consistently obtained similar runtimes. This
-comparison illustrates the performance achieved with
-\href{https://github.com/devitocodes/devito}{Devito} is at least on par
-with hand-coded propagators. Considering we do not take adantage of the
-\href{https://github.com/devitocodes/devito}{Devito} compiler full
-capabilities in the two dimensional case, we are confident that the
-generated code will be at least on par with hand-coded version for three
-dimensional probalems and this comparison will be part of future work.
+The runtime obtained for this problem for both propagators were identical
+with less than one percent of difference. This similar runtime were
+obtained both with the Intel compiler and the GNU compiler and we ran the
+experiment with a fourth order and a sixth order discretization.  The
+kernels were executed five time each to ensure consistency between the
+results and we consistently obtained similar runtimes. This comparison
+illustrates the performance achieved with Devito is at least on par with
+hand-coded propagators. Considering we do not take adantage of the Devito
+compiler full capabilities in the two dimensional case, we are confident
+that the generated code will be at least on par with hand-coded version
+for three dimensional probalems and this comparison will be part of future
+work.
 
@@
 acoustic wave-equation, to real-world application can be challenging in
-particular when attempted as an afterthought. In this work, we showed
-that thanks to design principles that aimed at this type of application
-from the beginning, \href{https://github.com/devitocodes/devito}{Devito}
-provides the scalability necessary. First, we showed that we provide a
-high-level interface not only for simple scalar equations but for
-coupled PDEs and allow the expression of non-trivial differential
-operators in a simple and concise way. Second and most importantly, we
-demonstrated that the compiler enables large-scale with state-of-the art
-computational performance and programming paradigm. The single-node
-performance is on par with hand-coded codes while providing the
-flexibility from the symbolic interface, and multi-node parallelism is
+particular when attempted as an afterthought. In this work, we showed that
+thanks to design principles that aimed at this type of application from
+the beginning, Devito provides the scalability necessary. First, we showed
+that we provide a high-level interface not only for simple scalar
+equations but for coupled PDEs and allow the expression of non-trivial
+differential operators in a simple and concise way. Second and most
+importantly, we demonstrated that the compiler enables large-scale with
+state-of-the art computational performance and programming paradigm. The
+single-node performance is on par with hand-coded codes while providing
+the flexibility from the symbolic interface, and multi-node parallelism is
 integrated in the compiler and interface in a conservative way. Finally,

### 2020-04-21 paper.tex
More polishing

@@
 in Devito; the interested reader should refer to the MPI tutorial at
-\href{https://github.com/devitocodes/devito/blob/master/examples/mpi/overview.ipynb}{mpi-notebook}
+\href{https://github.com/devitocodes/devito/blob/v4.2/examples/mpi/overview.ipynb}{mpi-notebook}
 for thorough explanations and practical examples.
@@
 
-We briefly describe the computational setup and the performance achieved for
-this anisotropic imaging problem. Due to time constraints, and because the
-resources we were given access to for this proof of concept with Azure were
-somewhat limited, we did not have access to HPC virtual machines (VM) nor
-Infiniband enabled ones. This experiment was carried out on
-\texttt{Standard\_E64\_v3} and \texttt{Standard\_E64s\_v3} nodes which, while
-not HPC VM nodes, are memory optimized thus allowing to the wavefield to be
-stored in memory for imaging (TTI adjoint state gradient \cite{virieux,
-louboutin2018segeow}). These VMs are Intel® Broadwell E5-2673 v4 2.3GH that are
-dual socket, 32 physical cores (with hyperthreading enabled) and 432Gb of DRAM.
-The overall inversion involved computing the image for 1500 source positions,
-i.e.~solving 1500 forward and 1500 adjoint TTI wave-equations. A single image
-required, in single precision, 600Gb of memory. Two VMs were used per source and
-MPI set with one rank per socket (4 MPI ranks per source) and 100 sources were
-imaged in parallel. The performance achieved was as follows:
+We briefly describe the computational setup and the performance achieved
+for this anisotropic imaging problem. Due to time constraints, and because
+the resources we were given access to for this proof of concept with Azure
+were somewhat limited, we did not have access to Infiniband-enabled
+virtual machines (VM). This experiment was carried out on
+\texttt{Standard\_E64\_v3} and \texttt{Standard\_E64s\_v3} nodes which,
+while not HPC VM nodes, are memory optimized thus allowing to the
+wavefield to be stored in memory for imaging (TTI adjoint state gradient
+\cite{virieux, louboutin2018segeow}). These VMs are Intel® Broadwell
+E5-2673 v4 2.3GH that are dual socket, 32 physical cores (with
+hyperthreading enabled) and 432Gb of DRAM.  The overall inversion involved
+computing the image for 1500 source positions, i.e.~solving 1500 forward
+and 1500 adjoint TTI wave-equations. A single image required, in single
+precision, 600Gb of memory. Two VMs were used per source and MPI set with
+one rank per socket (4 MPI ranks per source) and 100 sources were imaged
+in parallel. The performance achieved was as follows:
 
@@
 
-\cite{witte2019TPDedas} describe the serverless implementation of seismic
+\cite{witte2019TPDedas} describes the serverless implementation of seismic
 inverse problems in detail, including iterative algorithms for least-square
@@
 \href{https://batch-shipyard.readthedocs.io}{batch-shipyard} setup. This example
-can also be easily run on a traditional HPC cluster environment using, for
-example, \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or Dask~\cite{dask}
-for parallelization over sources.
+can also be easily run on a traditional HPC cluster environment using.
 
@@
 The extended Devito language introduces two new types,
-\texttt{VectorFunction}/ \texttt{VectorTimeFunction} for vectorial objects
+\texttt{VectorFunction}/\texttt{VectorTimeFunction} for vectorial objects
 such as the particle velocity, and
@@
 component of a tensorial object -- a (scalar) Devito \texttt{Function} --
-is accessible via conventional vector notation (i.e.
-\texttt{v{[}0{]},\phantom{\ }t{[}0,1{]},....}).
+is accessible via conventional vector notation (e.g.
+\texttt{v{[}0{]},\phantom{\ }t{[}0,1{]}}).
 
@@
 is omitted for readability (the equation does not fit into a single page).
-However, it can be found in the Devito tutorial on elastic modelling
-on GitHub.}\label{PrettyElas}
+    However, it can be found in a Devito
+    \href{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/tutorials/06_elastic.ipynb}{tutorial}
+    on elastic modelling.}\label{PrettyElas}
 \end{figure*}
@@
 \item
-  Three model parameters \texttt{lamda}, \texttt{mu} and \texttt{rho}
+  Three model parameters \texttt{lambda}, \texttt{mu} and \texttt{rho}
 \end{itemize}
@@
 minimum memory requirement of 461Gb for modelling alone. For this experiment,
-access was obtained for small HPC VMs (on azure) called \texttt{Standard\_H16r}.
+access was obtained for small HPC VMs (on Azure) called \texttt{Standard\_H16r}.
 These VMs contain 16 core Intel Xeon E5 2667 v3 chips, with no hyperthreading,
@@
 with MPI. Future work will involve working on InfiniBand enabled and
-true HPC VMs on azure to achieve Cloud performance on par with that of state of
+true HPC VMs on Azure to achieve Cloud performance on par with that of state of
 the art HPC clusters. Extrapolating from the performance obtained in this

### 2020-04-21 paper.tex
added new code samples

@@
 user to formulate wave-equations, and more generally time-dependent PDEs
-in a simple and mathematically coherent way. The Devito compiler then
+in a simple and mathematically coherent way. The\href{https://github.com/devitocodes/devito}{Devito}compiler then
 automatically generates finite-difference stencils from these mathematical
-expressions. One of the main advantages of Devito over other
+expressions. One of the main advantages of\href{https://github.com/devitocodes/devito}{Devito}over other
 finite-difference DSLs is that generic expressions such as sparse
@@
 fully supported and expressible in a high-level fashion. The second
-component of Devito is its compiler (c.f \cite{devito-compiler}) that
+component of\href{https://github.com/devitocodes/devito}{Devito}is its compiler (c.f \cite{devito-compiler}) that
 generates highly optimized C code. The generated code is then compiled at
@@
 forward model to be saved in memory). Secondly, through the elastic
-modelling example, we demonstrate that Devito now fully supports and
+modelling example, we demonstrate that\href{https://github.com/devitocodes/devito}{Devito}now fully supports and
 automates vectorial and second order tensorial staggered-grid
@@
 
-It is up to the Devito compiler to translate the symbolic specification
+It is up to the\href{https://github.com/devitocodes/devito}{Devito}compiler to translate the symbolic specification
 into C code. The lowering of the input language down to C consists of
@@
 optimizations (e.g., cache blocking, alignment, SIMD and OpenMP
-parallelism), Devito applies a series of FLOP-reducing transformations as
+parallelism),\href{https://github.com/devitocodes/devito}{Devito}applies a series of FLOP-reducing transformations as
 well as aggressive loop fusion. For a complete treatment, the interested
@@
 
-In this section we illustrate the Devito language showing how to implement
+In this section we illustrate the\href{https://github.com/devitocodes/devito}{Devito}language showing how to implement
 the acoustic wave-equation in isotropic media
@@
 \begin{python}
-from devito import Grid
+from\href{https://github.com/devitocodes/devito}{Devito}import Grid
 grid = Grid(shape=(nx, ny, nz),
@@
 \begin{python}
-from devito import Function, TimeFunction
+from\href{https://github.com/devitocodes/devito}{Devito}import Function, TimeFunction
 m = Function(name="m", grid=grid,
@@
 \begin{python}
-from devito import SparseTimeFunction
+from\href{https://github.com/devitocodes/devito}{Devito}import SparseTimeFunction
 s = SparseTimeFunction(name="src",
@@
 operation called injection, while the measurement is handled with an
-interpolation. By default, Devito initializes all \texttt{Function} data
+interpolation. By default,\href{https://github.com/devitocodes/devito}{Devito}initializes all \texttt{Function} data
 to 0, and thus automatically satisfies the zero Dirichlet condition at
 \texttt{t=0}. The isotropic acoustic wave-equation can then be implemented
-in Devito as below
+in\href{https://github.com/devitocodes/devito}{Devito}as below
 
 \begin{python}
-from devito import solve, Eq, Operator
+from\href{https://github.com/devitocodes/devito}{Devito}import solve, Eq, Operator
 eq = m * u.dt2 - u.laplace
@@
 \begin{python}
-from devito import Operator
+from\href{https://github.com/devitocodes/devito}{Devito}import Operator
 op = Operator(update + src_eqns + d_eqns)
@@
 
-The output of the Devito compiler for the running example used in this
+The output of the\href{https://github.com/devitocodes/devito}{Devito}compiler for the running example used in this
 section is available at
@@
 is such that users can almost entirely abstract away from it and reuse
-non-distributed code as is. Given any Devito code, just running it as
+non-distributed code as is. Given any\href{https://github.com/devitocodes/devito}{Devito}code, just running it as
 
@@
 \texttt{SparseFunction} objects the approach is different. Since a
-\texttt{SparseFunction} represents a sparse set of points, Devito looks at
+\texttt{SparseFunction} represents a sparse set of points,\href{https://github.com/devitocodes/devito}{Devito}looks at
 the physical coordinates of each point and, based on the \texttt{Grid}
@@
 
-When accessing or manipulating data in a Devito code, users have the
+When accessing or manipulating data in a\href{https://github.com/devitocodes/devito}{Devito}code, users have the
 illusion to be working with classic NumPy arrays, while underneath they
@@
 \texttt{(u\phantom{\ }+\phantom{\ }v.dx).dy}. As a consequence, the
-rotated derivative defined in~\ref{rot} is implemented with Devito in two
+rotated derivative defined in~\ref{rot} is implemented with\href{https://github.com/devitocodes/devito}{Devito}in two
 lines as:
@@
 rotating registers are used). While an explanation of the optimization
-strategies employed by Devito is beyond the scope of this paper (see
+strategies employed by\href{https://github.com/devitocodes/devito}{Devito}is beyond the scope of this paper (see
 \cite{devito-compiler} for details), what is emphasized here is that users
@@
 latest geophysical insights in tandem with those from high-performance
-computing. With Devito on the other hand, both the forward and adjoint
+computing. With\href{https://github.com/devitocodes/devito}{Devito}on the other hand, both the forward and adjoint
 equations can be implemented in a few lines of Python code as illustrated
@@
 
-We now demonstrate the scalability of Devito to real-world applications by
+We now demonstrate the scalability of\href{https://github.com/devitocodes/devito}{Devito}to real-world applications by
 imaging an industry-scale three-dimensional TTI subsurface model. This
@@
 described. This imaging project, in collaboration with Azure, demonstrates
-the scalability and robustness of Devito to large scale wave-equation
+the scalability and robustness of\href{https://github.com/devitocodes/devito}{Devito}to large scale wave-equation
 based inverse problems in combination with a cost-effective serverless
@@
 symbolic FLOP-reducing transformations have been performed during
-compilation.  Devito uses an in-house estimate of cost, rather than
+compilation. \href{https://github.com/devitocodes/devito}{Devito}uses an in-house estimate of cost, rather than
 \texttt{SymPy}'s estimate, to take care of some low-level intricacies. For
@@
 offset indexing into multi-dimensional arrays. To calculate the total
-number of FLOPs performed, Devito multiplies the floating-point operations
+number of FLOPs performed,\href{https://github.com/devitocodes/devito}{Devito}multiplies the floating-point operations
 calculated at compile time by the size of the iteration space, and it does
@@
 code motion, the amount of innermost-loop-invariant sub-expressions in an
-\texttt{Operator} is typically negligible and therefore the Devito estimate
+\texttt{Operator} is typically negligible and therefore the\href{https://github.com/devitocodes/devito}{Devito}estimate
 does not suffer from this issue, or at least not, to the best of our
@@
 checked against those produced by Intel Advisor on several single-node
-experiments: the differences -- typically Devito underestimating the
+experiments: the differences -- typically\href{https://github.com/devitocodes/devito}{Devito}underestimating the
 achieved performance -- were always at most in the order of units, and
@@
 ignored in acoustic models) are the most hazardous. In this section, we
-exploit the tensor algebra language introduced in Devito v4.0 to express
+exploit the tensor algebra language introduced in\href{https://github.com/devitocodes/devito}{Devito}v4.0 to express
 an elastic model with compact and elegant notation.
@@
 
-We have augmented the Devito language with tensorial objects to enable
+We have augmented the\href{https://github.com/devitocodes/devito}{Devito}language with tensorial objects to enable
 straightforward and mathematically rigorous definitions of
@@
 
-The extended Devito language introduces two new types,
+The extended\href{https://github.com/devitocodes/devito}{Devito}language introduces two new types,
 \texttt{VectorFunction}/\texttt{VectorTimeFunction} for vectorial objects
@@
 finite-differences with the possibility of half-node averaging. Each
-component of a tensorial object -- a (scalar) Devito \texttt{Function} --
+component of a tensorial object -- a (scalar)\href{https://github.com/devitocodes/devito}{Devito}\texttt{Function} --
 is accessible via conventional vector notation (e.g.
@@
 \begin{python}
-from devito import VectorTimeFunction,
+from\href{https://github.com/devitocodes/devito}{Devito}import VectorTimeFunction,
                    TensorTimeFunction
@@
 representation reflects perfectly the mathematics while maintaining
-computational portability and efficiency through the Devito compiler. The
+computational portability and efficiency through the\href{https://github.com/devitocodes/devito}{Devito}compiler. The
 complete generated code for the elastic wave-equation with and without MPI
@@
 \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
-in \texttt{elastic-so8.c} and \texttt{elastic-so8-mpi.c}.
+in \texttt{elastic-so12.c} and \texttt{elastic-so8-mpi.c}.
 
@@
     However, it can be found in a Devito
-    \href{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/tutorials/06_elastic.ipynb}{tutorial}
+    \href{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/tutorials/06_elastic_varying_parameters.ipynb}{tutorial}
     on elastic modelling.}\label{PrettyElas}
@@
 in the subsurface (at a depth of approximately 1km). Figures \ref{ElasWf}
-and \ref{ElasShot} demonstrate that this high-level Devito implementation
+and \ref{ElasShot} demonstrate that this high-level\href{https://github.com/devitocodes/devito}{Devito}implementation
 of the elastic wave-equation is effective and accurate.  Importantly, in
-constructing this model within the Devito DSL framework, computational
+constructing this model within the\href{https://github.com/devitocodes/devito}{Devito}DSL framework, computational
 technicalities such as the staggered grid analysis are abstracted away. We
@@
 Finally, three dimensional elastic data was modelled in the Cloud to
-demonstrate the scalability of Devito to cluster-size problems. The model
+demonstrate the scalability of\href{https://github.com/devitocodes/devito}{Devito}to cluster-size problems. The model
 used in this experiment mimics a reference model in geophysics known as
@@
 Because earlier performance benchmarks focus mainly on roofline model
-analysis, we also compared the runtime of Devito with a reference
+analysis, we also compared the runtime of\href{https://github.com/devitocodes/devito}{Devito}with a reference
 open source hand-coded propagator in collaboration with its author for
@@
 results and we consistently obtained similar runtimes. This comparison
-illustrates the performance achieved with Devito is at least on par with
+illustrates the performance achieved with\href{https://github.com/devitocodes/devito}{Devito}is at least on par with
 hand-coded propagators. Considering we do not take adantage of the Devito
@@
 thanks to design principles that aimed at this type of application from
-the beginning, Devito provides the scalability necessary. First, we showed
+the beginning,\href{https://github.com/devitocodes/devito}{Devito}provides the scalability necessary. First, we showed
 that we provide a high-level interface not only for simple scalar

### 2020-04-21 gencode/elastic-so12-mpi.c

@@ -36,23 +36,19 @@ struct profiler
 void sendrecv_txyz(struct dataobj *restrict a_vec, const int buf_x_size, const int buf_y_size, const int buf_z_size, int ogtime, int ogx, int ogy, int ogz, int ostime, int osx, int osy, int osz, int fromrank, int torank, MPI_Comm comm, const int nthreads);
 void gather_txyz(float *restrict buf_vec, const int buf_x_size, const int buf_y_size, const int buf_z_size, struct dataobj *restrict a_vec, int otime, int ox, int oy, int oz, const int nthreads);
 void scatter_txyz(float *restrict buf_vec, const int buf_x_size, const int buf_y_size, const int buf_z_size, struct dataobj *restrict a_vec, int otime, int ox, int oy, int oz, const int nthreads);
-void haloupdate7(struct dataobj *restrict a_vec, MPI_Comm comm, struct neighborhood * nb, int otime, const int nthreads);
-void haloupdate0(struct dataobj *restrict a_vec, MPI_Comm comm, struct neighborhood * nb, int otime, const int nthreads);
-void haloupdate1(struct dataobj *restrict a_vec, MPI_Comm comm, struct neighborhood * nb, int otime, const int nthreads);
-void haloupdate2(struct dataobj *restrict a_vec, MPI_Comm comm, struct neighborhood * nb, int otime, const int nthreads);
-void haloupdate3(struct dataobj *restrict a_vec, MPI_Comm comm, struct neighborhood * nb, int otime, const int nthreads);
-void haloupdate4(struct dataobj *restrict a_vec, MPI_Comm comm, struct neighborhood * nb, int otime, const int nthreads);
-void haloupdate5(struct dataobj *restrict a_vec, MPI_Comm comm, struct neighborhood * nb, int otime, const int nthreads);
-void haloupdate6(struct dataobj *restrict a_vec, MPI_Comm comm, struct neighborhood * nb, int otime, const int nthreads);
[…]

### 2020-04-21 gencode/elastic-so12.c

@@ -1,0 +1,475 @@
+#define _POSIX_C_SOURCE 200809L
+#include "stdlib.h"
+#include "math.h"
+#include "sys/time.h"
+#include "xmmintrin.h"
+#include "pmmintrin.h"
+#include "omp.h"
+
+struct dataobj
+{
+  void *restrict data;
+  int * size;
+  int * npsize;
+  int * dsize;
+  int * hsize;
+  int * hofs;
+  int * oofs;
+} ;
+
+struct profiler
+{
+  double section0;
+  double section1;
+  double section2;
+  double section3;
+} ;
+
+void bf0(struct dataobj *restrict damp_vec, struct dataobj *restrict irho_vec, struct dataobj *restrict tau_xx_vec, struct dataobj *restrict tau_xy_vec, struct dataobj *restrict tau_xz_vec, struct dataobj *restrict tau_yy_vec, struct dataobj *restrict tau_yz_vec, struct dataobj *restrict tau_zz_vec, struct dataobj *restrict v_x_vec, struct dataobj *restrict v_y_vec, struct dataobj *restrict v_z_vec, const int t0, const int t1, const int x0_blk0_size, const int x_M, const int x_m, const int y0_blk0_size, const int y_M, const int y_m, const int z_M, const int z_m, const int nthreads);
[…]

### 2020-04-21 gencode/elastic-so8.c

@@ -1,476 +1,0 @@
-#define _POSIX_C_SOURCE 200809L
-#include "stdlib.h"
-#include "math.h"
-#include "sys/time.h"
-#include "xmmintrin.h"
-#include "pmmintrin.h"
-#include "omp.h"
-
-struct dataobj
-{
-  void *restrict data;
-  int * size;
-  int * npsize;
-  int * dsize;
-  int * hsize;
-  int * hofs;
-  int * oofs;
-} ;
-
-struct profiler
-{
-  double section0;
-  double section1;
-  double section2;
-  double section3;
-} ;
-
-void bf0(struct dataobj *restrict damp_vec, struct dataobj *restrict irho_vec, struct dataobj *restrict tau_xx_vec, struct dataobj *restrict tau_xy_vec, struct dataobj *restrict tau_xz_vec, struct dataobj *restrict tau_yy_vec, struct dataobj *restrict tau_yz_vec, struct dataobj *restrict tau_zz_vec, struct dataobj *restrict v_x_vec, struct dataobj *restrict v_y_vec, struct dataobj *restrict v_z_vec, const int t0, const int t1, const int x0_blk0_size, const int x_M, const int x_m, const int y0_blk0_size, const int y_M, const int y_m, const int z_M, const int z_m, const int nthreads);
[…]

### 2020-04-21 paper.tex
affiliations but small changes

--- 2020-04-21 paper.tex
@@
 \IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
-\IEEEauthorblockA{\textit{Devito Codes}}
+\IEEEauthorblockA{\textit{Devito Codes}\\
+London, UK \\
+fabio@devitocodes.com}
 \and
 \IEEEauthorblockN{3\textsuperscript{th} Rhodri Nelson}
-\IEEEauthorblockA{\textit{Imperial College London}}
+\IEEEauthorblockA{\textit{Imperial College London}\\
+London, UK \\
+rhodri.nelson@imperial.ac.uk}
 \and
 \IEEEauthorblockN{4\textsuperscript{rd} Philipp Witte}
-\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
+\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
+Atlanta, GA \\
+pwitte3@gatech.edu}
 \and
 \IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
-\IEEEauthorblockA{\textit{Imperial College London}}
+\IEEEauthorblockA{\textit{Imperial College London}\\
+London, UK \\
+g.bisbas18@imperial.ac.uk}
 \and
 \IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
-\IEEEauthorblockA{\textit{TU-Delft}}
+\IEEEauthorblockA{\textit{TU-Delft}\\
+Delft, NL \\
+J.W.Thorbecke@tudelft.nl}
 \and
 \IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
-\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
+\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
+Atlanta, GA\\
+felix.herrmann@gatech.edu}
 \and
 \IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
-\IEEEauthorblockA{\textit{Imperial College London}}
+\IEEEauthorblockA{\textit{Imperial College London}\\
+London, UK \\
+g.gorma@imperial.ac.uk&}
 }
@@
 \cite{tarantola}, among the first successful
-expositions on real 3D data was presented in \cite{sirgue2009}. Other
+expositions on real 3D data was presented in \cite{sirgue}. Other
 studies utilizing FDM within the FWI workflow include
@@
 user to formulate wave-equations, and more generally time-dependent PDEs
-in a simple and mathematically coherent way. The Devito compiler then
+in a simple and mathematically coherent way. The \href{https://github.com/devitocodes/devito}{Devito} compiler then
 automatically generates finite-difference stencils from these mathematical
-expressions. One of the main advantages of Devito over other
+expressions. One of the main advantages of \href{https://github.com/devitocodes/devito}{Devito} over other
 finite-difference DSLs is that generic expressions such as sparse
@@
 fully supported and expressible in a high-level fashion. The second
-component of Devito is its compiler (c.f \cite{devito-compiler}) that
+component of \href{https://github.com/devitocodes/devito}{Devito} is its compiler (c.f \cite{devito-compiler}) that
 generates highly optimized C code. The generated code is then compiled at
@@
 forward model to be saved in memory). Secondly, through the elastic
-modelling example, we demonstrate that Devito now fully supports and
+modelling example, we demonstrate that \href{https://github.com/devitocodes/devito}{Devito} now fully supports and
 automates vectorial and second order tensorial staggered-grid
@@
 
-It is up to the Devito compiler to translate the symbolic specification
+It is up to the \href{https://github.com/devitocodes/devito}{Devito} compiler to translate the symbolic specification
 into C code. The lowering of the input language down to C consists of
@@
 optimizations (e.g., cache blocking, alignment, SIMD and OpenMP
-parallelism), Devito applies a series of FLOP-reducing transformations as
+parallelism), \href{https://github.com/devitocodes/devito}{Devito} applies a series of FLOP-reducing transformations as
 well as aggressive loop fusion. For a complete treatment, the interested
@@
 
-In this section we illustrate the Devito language showing how to implement
+In this section we illustrate the \href{https://github.com/devitocodes/devito}{Devito} language showing how to implement
 the acoustic wave-equation in isotropic media
@@
 operation called injection, while the measurement is handled with an
-interpolation. By default, Devito initializes all \texttt{Function} data
+interpolation. By default, \href{https://github.com/devitocodes/devito}{Devito} initializes all \texttt{Function} data
 to 0, and thus automatically satisfies the zero Dirichlet condition at
 \texttt{t=0}. The isotropic acoustic wave-equation can then be implemented
-in Devito as below
+in \href{https://github.com/devitocodes/devito}{Devito} as below
 
@@
 
-The output of the Devito compiler for the running example used in this
+The output of the \href{https://github.com/devitocodes/devito}{Devito} compiler for the running example used in this
 section is available at
@@
 is such that users can almost entirely abstract away from it and reuse
-non-distributed code as is. Given any Devito code, just running it as
+non-distributed code as is. Given any \href{https://github.com/devitocodes/devito}{Devito} code, just running it as
 
@@
 \texttt{SparseFunction} objects the approach is different. Since a
-\texttt{SparseFunction} represents a sparse set of points, Devito looks at
+\texttt{SparseFunction} represents a sparse set of points, \href{https://github.com/devitocodes/devito}{Devito} looks at
 the physical coordinates of each point and, based on the \texttt{Grid}
@@
 
-When accessing or manipulating data in a Devito code, users have the
+When accessing or manipulating data in a \href{https://github.com/devitocodes/devito}{Devito} code, users have the
 illusion to be working with classic NumPy arrays, while underneath they
@@
 \texttt{(u\phantom{\ }+\phantom{\ }v.dx).dy}. As a consequence, the
-rotated derivative defined in~\ref{rot} is implemented with Devito in two
+rotated derivative defined in~\ref{rot} is implemented with \href{https://github.com/devitocodes/devito}{Devito} in two
 lines as:
@@
 rotating registers are used). While an explanation of the optimization
-strategies employed by Devito is beyond the scope of this paper (see
+strategies employed by \href{https://github.com/devitocodes/devito}{Devito} is beyond the scope of this paper (see
 \cite{devito-compiler} for details), what is emphasized here is that users
@@
 latest geophysical insights in tandem with those from high-performance
-computing. With Devito on the other hand, both the forward and adjoint
+computing. With \href{https://github.com/devitocodes/devito}{Devito} on the other hand, both the forward and adjoint
 equations can be implemented in a few lines of Python code as illustrated
@@
 
-We now demonstrate the scalability of Devito to real-world applications by
+We now demonstrate the scalability of \href{https://github.com/devitocodes/devito}{Devito} to real-world applications by
 imaging an industry-scale three-dimensional TTI subsurface model. This
@@
 described. This imaging project, in collaboration with Azure, demonstrates
-the scalability and robustness of Devito to large scale wave-equation
+the scalability and robustness of \href{https://github.com/devitocodes/devito}{Devito} to large scale wave-equation
 based inverse problems in combination with a cost-effective serverless
@@
 symbolic FLOP-reducing transformations have been performed during
-compilation.  Devito uses an in-house estimate of cost, rather than
+compilation.  \href{https://github.com/devitocodes/devito}{Devito} uses an in-house estimate of cost, rather than
 \texttt{SymPy}'s estimate, to take care of some low-level intricacies. For
@@
 offset indexing into multi-dimensional arrays. To calculate the total
-number of FLOPs performed, Devito multiplies the floating-point operations
+number of FLOPs performed, \href{https://github.com/devitocodes/devito}{Devito} multiplies the floating-point operations
 calculated at compile time by the size of the iteration space, and it does
@@
 code motion, the amount of innermost-loop-invariant sub-expressions in an
-\texttt{Operator} is typically negligible and therefore the Devito estimate
+\texttt{Operator} is typically negligible and therefore the \href{https://github.com/devitocodes/devito}{Devito} estimate
 does not suffer from this issue, or at least not, to the best of our
@@
 checked against those produced by Intel Advisor on several single-node
-experiments: the differences -- typically Devito underestimating the
+experiments: the differences -- typically \href{https://github.com/devitocodes/devito}{Devito} underestimating the
 achieved performance -- were always at most in the order of units, and
@@
 \href{https://batch-shipyard.readthedocs.io}{batch-shipyard} setup. This example
-can also be easily run on a traditional HPC cluster environment using.
+can also be easily run on a traditional HPC cluster environment using, for example, 
+\href{https://github.com/slimgroup/JUDI.jl}{JUDI}~\cite{witte2018alf} or Dask~\cite{dask}
+for parallelization over sources.
 
@@
 ignored in acoustic models) are the most hazardous. In this section, we
-exploit the tensor algebra language introduced in Devito v4.0 to express
+exploit the tensor algebra language introduced in \href{https://github.com/devitocodes/devito}{Devito} v4.0 to express
 an elastic model with compact and elegant notation.
@@
 
-We have augmented the Devito language with tensorial objects to enable
+We have augmented the \href{https://github.com/devitocodes/devito}{Devito} language with tensorial objects to enable
 straightforward and mathematically rigorous definitions of
@@
 
-The extended Devito language introduces two new types,
-\texttt{VectorFunction}/\texttt{VectorTimeFunction} for vectorial objects
-such as the particle velocity, and
+The extended \href{https://github.com/devitocodes/devito}{Devito} language introduces two new types, 
+\texttt{VectorFunction}/\texttt{VectorTimeFunction} for vectorial objects such as the particle velocity, and
 \texttt{TensorFunction}/\texttt{TensorTimeFunction} for second-order
@@
 finite-differences with the possibility of half-node averaging. Each
-component of a tensorial object -- a (scalar) Devito \texttt{Function} --
+component of a tensorial object -- a (scalar) \href{https://github.com/devitocodes/devito}{Devito} \texttt{Function} --
 is accessible via conventional vector notation (e.g.
@@
 representation reflects perfectly the mathematics while maintaining
-computational portability and efficiency through the Devito compiler. The
+computational portability and efficiency through the \href{https://github.com/devitocodes/devito}{Devito} compiler. The
 complete generated code for the elastic wave-equation with and without MPI
@@
 \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
-in \texttt{elastic-so8.c} and \texttt{elastic-so8-mpi.c}.
+in \texttt{elastic-so12.c} and \texttt{elastic-so12-mpi.c}.
 
@@
 in the subsurface (at a depth of approximately 1km). Figures \ref{ElasWf}
-and \ref{ElasShot} demonstrate that this high-level Devito implementation
+and \ref{ElasShot} demonstrate that this high-level \href{https://github.com/devitocodes/devito}{Devito} implementation
 of the elastic wave-equation is effective and accurate.  Importantly, in
-constructing this model within the Devito DSL framework, computational
+constructing this model within the \href{https://github.com/devitocodes/devito}{Devito} DSL framework, computational
 technicalities such as the staggered grid analysis are abstracted away. We
@@
 Finally, three dimensional elastic data was modelled in the Cloud to
-demonstrate the scalability of Devito to cluster-size problems. The model
+demonstrate the scalability of \href{https://github.com/devitocodes/devito}{Devito} to cluster-size problems. The model
 used in this experiment mimics a reference model in geophysics known as
@@
 Because earlier performance benchmarks focus mainly on roofline model
-analysis, we also compared the runtime of Devito with a reference
+analysis, we also compared the runtime of \href{https://github.com/devitocodes/devito}{Devito} with a reference
 open source hand-coded propagator in collaboration with its author for
@@
 results and we consistently obtained similar runtimes. This comparison
-illustrates the performance achieved with Devito is at least on par with
+illustrates the performance achieved with \href{https://github.com/devitocodes/devito}{Devito} is at least on par with
 hand-coded propagators. Considering we do not take adantage of the Devito
@@
 thanks to design principles that aimed at this type of application from
-the beginning, Devito provides the scalability necessary. First, we showed
+the beginning, \href{https://github.com/devitocodes/devito}{Devito} provides the scalability necessary. First, we showed
 that we provide a high-level interface not only for simple scalar
@@
 
+\section{Code availability}
+
+The code to reproduce the different examples presented in this work is available online in the fllwing repositories:
+
+\begin{itemize}
+  \item The complete code for TTI imaging is available at \url{https://github.com/slimgroup/Azure2019} and includes 
+  the TTI propagators, the Azure setup for shot parallelism and a documentation.
+  \item The elastic modeling can be run with the elastic example available in Devito at
+  \url{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/elastic/elastic_example.py} 
+  and can be run with any size and spatial order.
+\end{itemize}
+
 \bibliography{sc20_paper}

### 2020-04-21 paper.tex
more code and renamed

@@
 section is available at
-\href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
+\href{https://github.com/mloubout/SC20Paper/tree/master/codesamples}{CodeSample}
 in \texttt{acou-so8.c}.
@@
 generated code for tti modeling with and without MPI is made available
-at \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
+at \href{https://github.com/mloubout/SC20Paper/tree/master/codesamples}{CodeSample}
 in \texttt{tti-so8-unoptimized.c}, \texttt{tti-so8.c} and \texttt{tti-so8-mpi.c}.
@@
 is made available at
-\href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
+\href{https://github.com/mloubout/SC20Paper/tree/master/codesamples}{CodeSample}
 in \texttt{elastic-so12.c} and \texttt{elastic-so12-mpi.c}.
@@
   \url{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/elastic/elastic_example.py} 
-  and can be run with any size and spatial order.
+  and can be run with any size and spatial order. A standalone script to run the lare 3D elastic modeling
+  is also available at {https://github.com/mloubout/SC20Paper/tree/master/codesamples}
 \end{itemize}

### 2020-04-21 codesamples/tti-so8-unoptimized.c

@@ -1,0 +1,348 @@
+#define _POSIX_C_SOURCE 200809L
+#include "stdlib.h"
+#include "math.h"
+#include "sys/time.h"
+#include "xmmintrin.h"
+#include "pmmintrin.h"
+#include "omp.h"
+
+struct dataobj
+{
+  void *restrict data;
+  int * size;
+  int * npsize;
+  int * dsize;
+  int * hsize;
+  int * hofs;
+  int * oofs;
+} ;
+
+struct profiler
+{
+  double section0;
+  double section1;
+  double section2;
+  double section3;
+} ;
+
+void bf0(struct dataobj *restrict damp_vec, const float dt, struct dataobj *restrict epsilon_vec, float *restrict r49_vec, float *restrict r50_vec, float *restrict r51_vec, float *restrict r52_vec, float *restrict r53_vec, struct dataobj *restrict u_vec, struct dataobj *restrict v_vec, struct dataobj *restrict vp_vec, const int x0_blk0_size, const int x_size, const int y0_blk0_size, const int y_size, const int z_size, const int t0, const int t1, const int t2, const int x_M, const int x_m, const int y_M, const int y_m, const int z_M, const int z_m, const int nthreads);
+
[…]

### 2020-04-21 paper.tex
math typo
update tti code

@@
 \author{
-\IEEEauthorblockN{1\textsuperscript{st} Louboutin Mathias}
+\IEEEauthorblockN{1\textsuperscript{st} Mathias Louboutin}
 \IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
@@
 bubesatti2016, chu2011, duveneck, fletcher, fowlertti2010,
-louboutin2018segeow, whitmore1983, witte2016segpve, xu2014, ,zhang2005,
+louboutin2018segeow, whitmore1983, witte2016segpve, xu2014, zhang2005,
 zhang2011, zhan2013}. The TTI wave-equation is an acoustic, low dimensional (4
@@
 Cloud. Our experiments are reproducible using the instructions in a public
-repository \href{https://github.com/slimgroup/Azure2019}{AzureTTI}, which
+repository \href{https://github.com/slimgroup/Azure2019/tree/v1.0}{AzureTTI}, which
 contains, among the other things, the Dockerfiles and Azure
@@
 \begin{aligned}
-    \tau = \begin{bmatrix}\tau_{xx}(t, x, y) & \tau_{xy}(t, x, y)\\\tau_{xy}t, x, y) & \tau_{yy}(t, x, y)\end{bmatrix}.
+    \tau = \begin{bmatrix}\tau_{xx}(t, x, y) & \tau_{xy}(t, x, y)\\\tau_{xy}(t, x, y) & \tau_{yy}(t, x, y)\end{bmatrix}.
 \end{aligned}
@@
 \begin{itemize}
-  \item The complete code for TTI imaging is available at \url{https://github.com/slimgroup/Azure2019} and includes 
+  \item The complete code for TTI imaging is available at \url{https://github.com/slimgroup/Azure2019/tree/v1.0} and includes 
   the TTI propagators, the Azure setup for shot parallelism and a documentation.

### 2020-04-21 codesamples/tti-so8-unoptimized.c

@@ -2,8 +2,6 @@
 #include "stdlib.h"
 #include "math.h"
 #include "sys/time.h"
-#include "xmmintrin.h"
-#include "pmmintrin.h"
 #include "omp.h"
 
 struct dataobj
@@ -22,14 +20,14 @@ struct profiler
   double section0;
   double section1;
   double section2;
-  double section3;
 } ;
 
-void bf0(struct dataobj *restrict damp_vec, const float dt, struct dataobj *restrict epsilon_vec, float *restrict r49_vec, float *restrict r50_vec, float *restrict r51_vec, float *restrict r52_vec, float *restrict r53_vec, struct dataobj *restrict u_vec, struct dataobj *restrict v_vec, struct dataobj *restrict vp_vec, const int x0_blk0_size, const int x_size, const int y0_blk0_size, const int y_size, const int z_size, const int t0, const int t1, const int t2, const int x_M, const int x_m, const int y_M, const int y_m, const int z_M, const int z_m, const int nthreads);
 
-int ForwardTTI(struct dataobj *restrict damp_vec, struct dataobj *restrict delta_vec, const float dt, struct dataobj *restrict epsilon_vec, const float o_x, const float o_y, const float o_z, struct dataobj *restrict phi_vec, struct dataobj *restrict rec_vec, struct dataobj *restrict rec_coords_vec, struct dataobj *restrict src_vec, struct dataobj *restrict src_coords_vec, struct dataobj *restrict theta_vec, struct dataobj *restrict u_vec, struct dataobj *restrict v_vec, struct dataobj *restrict vp_vec, const int x0_blk0_size, const int x_M, const int x_m, const int x_size, const int y0_blk0_size, const int y_M, const int y_m, const int y_size, const int z_M, const int z_m, const int z_size, const int p_rec_M, const int p_rec_m, const int p_src_M, const int p_src_m, const int time_M, const int time_m, struct profiler * timers, const int nthreads, const int nthreads_nonaffine)
[…]

### 2020-04-21 codesamples/tti-so8-unoptimized.c
format unoptimized tti

@@ -1,4 +1,4 @@
-#define _POSIX_C_SOURCE 200809L
+t#define _POSIX_C_SOURCE 200809L
 #include "stdlib.h"
 #include "math.h"
 #include "sys/time.h"
@@ -51,8 +51,166 @@ int ForwardTTI(struct dataobj *restrict damp_vec, struct dataobj *restrict delta
         {
           for (int z = z_m; z <= z_M; z += 1)
           {
-            u[t1][x + 8][y + 8][z + 8] = 1.0F*(2.0F*pow(dt, 2)*(sqrt(2*delta[x + 8][y + 8][z + 8] + 1)*(-8.33333346e-3F*(-(8.33333346e-3F*v[t0][x + 4][y + 8][z + 8] - 6.66666677e-2F*v[t0][x + 5][y + 8][z + 8] + 6.66666677e-2F*v[t0][x + 7][y + 8][z + 8] - 8.33333346e-3F*v[t0][x + 8][y + 8][z + 8])*sin(theta[x + 6][y + 8][z + 8])*cos(phi[x + 6][y + 8][z + 8]) - (8.33333346e-3F*v[t0][x + 6][y + 6][z + 8] - 6.66666677e-2F*v[t0][x + 6][y + 7][z + 8] + 6.66666677e-2F*v[t0][x + 6][y + 9][z + 8] - 8.33333346e-3F*v[t0][x + 6][y + 10][z + 8])*sin(phi[x + 6][y + 8][z + 8])*sin(theta[x + 6][y + 8][z + 8]) - (8.33333346e-3F*v[t0][x + 6][y + 8][z + 6] - 6.66666677e-2F*v[t0][x + 6][y + 8][z + 7] + 6.66666677e-2F*v[t0][x + 6][y + 8][z + 9] - 8.33333346e-3F*v[t0][x + 6][y + 8][z + 10])*cos(theta[x + 6][y + 8][z + 8]))*sin(theta[x + 6][y + 8][z + 8])*cos(phi[x + 6][y + 8][z + 8]) + 6.66666677e-2F*(-(8.33333346e-3F*v[t0][x + 5][y + 8][z + 8] - 6.66666677e-2F*v[t0][x + 6][y + 8][z + 8] + 6.66666677e-2F*v[t0][x + 8][y + 8][z + 8] - 8.33333346e-3F*v[
[…]

### 2020-04-21 paper.tex
misc fixup

@@
     \lstset{#1}
-}
-{}
+}{}
 
@@
 \newcommand{\NormalTok}[1]{{#1}}
+\newcommand{\devito}{\href{https://github.com/devitocodes/devito}{Devito} }
 
@@
 
-\title{Devito for large scale elastic modelling and anisotropic inversion}
+\title{Scaling through abstraction – high-performance vectorial wave simulations for seismic inversion with Devito}
 
@@
 London, UK \\
-g.gorma@imperial.ac.uk&}
+g.gorman@imperial.ac.uk}
 }
@@
 
- \href{https://github.com/devitocodes/devito}{Devito}  is an open-source
+\devito is an open-source
 Python project based on domain-specific language and compiler
@@
 user to formulate wave-equations, and more generally time-dependent PDEs
-in a simple and mathematically coherent way. The \href{https://github.com/devitocodes/devito}{Devito} compiler then
+in a simple and mathematically coherent way. The \devito compiler then
 automatically generates finite-difference stencils from these mathematical
-expressions. One of the main advantages of \href{https://github.com/devitocodes/devito}{Devito} over other
+expressions. One of the main advantages of \devito over other
 finite-difference DSLs is that generic expressions such as sparse
@@
 fully supported and expressible in a high-level fashion. The second
-component of \href{https://github.com/devitocodes/devito}{Devito} is its compiler (c.f \cite{devito-compiler}) that
+component of \devito is its compiler (c.f \cite{devito-compiler}) that
 generates highly optimized C code. The generated code is then compiled at
@@
 forward model to be saved in memory). Secondly, through the elastic
-modelling example, we demonstrate that \href{https://github.com/devitocodes/devito}{Devito} now fully supports and
+modelling example, we demonstrate that \devito now fully supports and
 automates vectorial and second order tensorial staggered-grid
@@
 
-It is up to the \href{https://github.com/devitocodes/devito}{Devito} compiler to translate the symbolic specification
+It is up to the \devito compiler to translate the symbolic specification
 into C code. The lowering of the input language down to C consists of
@@
 optimizations (e.g., cache blocking, alignment, SIMD and OpenMP
-parallelism), \href{https://github.com/devitocodes/devito}{Devito} applies a series of FLOP-reducing transformations as
+parallelism), \devito applies a series of FLOP-reducing transformations as
 well as aggressive loop fusion. For a complete treatment, the interested
@@
 
-In this section we illustrate the \href{https://github.com/devitocodes/devito}{Devito} language showing how to implement
+In this section we illustrate the \devito language showing how to implement
 the acoustic wave-equation in isotropic media
@@
 
-\begin{python}
+\begin{python}[label=grid, caption=Grid creation]
 from devito import Grid
@@
 \end{python}
-
+\noindent
 where \texttt{(nx,\phantom{\ }ny,\phantom{\ }nz)} are the number of grid
@@
 
-\begin{python}
+\begin{python}[label=function, caption=Function definition]
 from devito import Function, TimeFunction
@@
 \end{python}
-
+\noindent
 where \texttt{so} is the spatial discretization order and \texttt{to}
@@
 
-\begin{python}
+\begin{python}[label=sparse, caption=SparseFunction deifnition]
 from devito import SparseTimeFunction
@@
 operation called injection, while the measurement is handled with an
-interpolation. By default, \href{https://github.com/devitocodes/devito}{Devito} initializes all \texttt{Function} data
+interpolation. By default, \devito initializes all \texttt{Function} data
 to 0, and thus automatically satisfies the zero Dirichlet condition at
 \texttt{t=0}. The isotropic acoustic wave-equation can then be implemented
-in \href{https://github.com/devitocodes/devito}{Devito} as below
-
-\begin{python}
+in \devito as below
+
+\begin{python}[label=WE, caption=Wave-equation symbolic definition]
 from devito import solve, Eq, Operator
@@
 
-\begin{python}
+\begin{python}[label=op, caption=Operator creation]
 from devito import Operator
@@
 
-The output of the \href{https://github.com/devitocodes/devito}{Devito} compiler for the running example used in this
+The output of the \devito compiler for the running example used in this
 section is available at
@@
 is such that users can almost entirely abstract away from it and reuse
-non-distributed code as is. Given any \href{https://github.com/devitocodes/devito}{Devito} code, just running it as
+non-distributed code as is. Given any \devito code, just running it as
 
@@
 
+\noindent
 triggers the generation of code with routines for halo exchanges. The
@@
 \texttt{SparseFunction} objects the approach is different. Since a
-\texttt{SparseFunction} represents a sparse set of points, \href{https://github.com/devitocodes/devito}{Devito} looks at
+\texttt{SparseFunction} represents a sparse set of points, \devito looks at
 the physical coordinates of each point and, based on the \texttt{Grid}
@@
 
-When accessing or manipulating data in a \href{https://github.com/devitocodes/devito}{Devito} code, users have the
+When accessing or manipulating data in a \devito code, users have the
 illusion to be working with classic NumPy arrays, while underneath they
@@
 \end{Shaded}
-
+\noindent
 which is distributed across 4 ranks such that \texttt{rank\phantom{\ }0}
@@
 \end{Shaded}
-
+\noindent
 such that now \texttt{rank\phantom{\ }0} contains the elements
@@
 $\theta(x), \phi(x)$ that define the rotation of the vertical and horizontal
-axis around the cartesian directions. However, unlike the scalar isotropic
+axes around the cartesian directions. However, unlike the scalar isotropic
 acoustic wave-equation itself, the TTI wave-equation is extremely
@@
 \texttt{(u\phantom{\ }+\phantom{\ }v.dx).dy}. As a consequence, the
-rotated derivative defined in~\ref{rot} is implemented with \href{https://github.com/devitocodes/devito}{Devito} in two
+rotated derivative defined in~\ref{rot} is implemented with \devito in two
 lines as:
 
-\begin{python}
+\begin{python}[label=rotxpy, caption=Rotated finite-difference with symbolic shortcuts]
 dx_u = cos(theta) * cos(phi) * u.dx +
@@
 rotating registers are used). While an explanation of the optimization
-strategies employed by \href{https://github.com/devitocodes/devito}{Devito} is beyond the scope of this paper (see
+strategies employed by \devito is beyond the scope of this paper (see
 \cite{devito-compiler} for details), what is emphasized here is that users
-can easily take full advantage of these optimization without needed to
+can easily take full advantage of these optimizations without needed to
 concern themselves with the details.
@@
 that practitioners of seismic inversion are often geoscientists, not computer
-scientists/programmers. Low level implementations from non-specialists can often
+scientists/programmers. Low level implementations from non-specialists can
 often lead to poorly performing code. However, if research codes are passed to
@@
 latest geophysical insights in tandem with those from high-performance
-computing. With \href{https://github.com/devitocodes/devito}{Devito} on the other hand, both the forward and adjoint
+computing. With \devito on the other hand, both the forward and adjoint
 equations can be implemented in a few lines of Python code as illustrated
-with the rotated operator in \#rotxpy. The low level optimization element
+with the rotated operator in Listing~\ref{rotxpy}. The low level optimization element
 of the development is then taken care of under the hood by the Devito
@@
 
-We now demonstrate the scalability of \href{https://github.com/devitocodes/devito}{Devito} to real-world applications by
+We now demonstrate the scalability of \devito to real-world applications by
 imaging an industry-scale three-dimensional TTI subsurface model. This
@@
 described. This imaging project, in collaboration with Azure, demonstrates
-the scalability and robustness of \href{https://github.com/devitocodes/devito}{Devito} to large scale wave-equation
+the scalability and robustness of \devito to large scale wave-equation
 based inverse problems in combination with a cost-effective serverless
@@
 using two nodes per source experiment.
-
+\vskip 0.1in
+\noindent
 \textbf{\emph{Computational performance}}
-
+\vskip 0.1in
 We briefly describe the computational setup and the performance achieved
@@
 theoretically leads to a performance of 0.4PFLOP/s.
-
+\vskip 0.1in
+\noindent
 \textbf{\emph{How performance was measured}}
-
+\vskip 0.1in
 The execution time is computed through Python-level timers prefixed by an
@@
 symbolic FLOP-reducing transformations have been performed during
-compilation.  \href{https://github.com/devitocodes/devito}{Devito} uses an in-house estimate of cost, rather than
+compilation.  \devito uses an in-house estimate of cost, rather than
 \texttt{SymPy}'s estimate, to take care of some low-level intricacies. For
@@
 offset indexing into multi-dimensional arrays. To calculate the total
-number of FLOPs performed, \href{https://github.com/devitocodes/devito}{Devito} multiplies the floating-point operations
+number of FLOPs performed, \devito multiplies the floating-point operations
 calculated at compile time by the size of the iteration space, and it does
@@
 code motion, the amount of innermost-loop-invariant sub-expressions in an
-\texttt{Operator} is typically negligible and therefore the \href{https://github.com/devitocodes/devito}{Devito} estimate
+\texttt{Operator} is typically negligible and therefore the \devito estimate
 does not suffer from this issue, or at least not, to the best of our
@@
 checked against those produced by Intel Advisor on several single-node
-experiments: the differences -- typically \href{https://github.com/devitocodes/devito}{Devito} underestimating the
+experiments: the differences -- typically \devito underestimating the
 achieved performance -- were always at most in the order of units, and
@@
 
+\vskip 0.1in
+\noindent
 \textbf{\emph{Imaging result}}
-
+\vskip 0.1in
 The subsurface velocity model used in this study is an artificial anisotropic
@@
 ignored in acoustic models) are the most hazardous. In this section, we
-exploit the tensor algebra language introduced in \href{https://github.com/devitocodes/devito}{Devito} v4.0 to express
+exploit the tensor algebra language introduced in \devito v4.0 to express
 an elastic model with compact and elegant notation.
@@
 %
+\noindent
  where $v$ is a vector valued function with one component per cartesian
@@
 equations in two dimensions (two equations for the particle velocity and
-three for the stress) and nine equations in three dimensions (three for velocity
-and six for stress). However, the mathematical
+three for the stress) and nine equations in three dimensions (three for the particle velocity
+and six for the stress). However, the mathematical
 definition only require two coupled vector/tensor-valued equations for
@@
 
-We have augmented the \href{https://github.com/devitocodes/devito}{Devito} language with tensorial objects to enable
+We have augmented the \devito language with tensorial objects to enable
 straightforward and mathematically rigorous definitions of
@@
 
-The extended \href{https://github.com/devitocodes/devito}{Devito} language introduces two new types, 
+The extended \devito language introduces two new types, 
 \texttt{VectorFunction}/\texttt{VectorTimeFunction} for vectorial objects such as the particle velocity, and
@@
 finite-differences with the possibility of half-node averaging. Each
-component of a tensorial object -- a (scalar) \href{https://github.com/devitocodes/devito}{Devito} \texttt{Function} --
+component of a tensorial object -- a (scalar) \devito \texttt{Function} --
 is accessible via conventional vector notation (e.g.
@@
 
-\begin{python}
+\begin{python}[label=elastic, caption=Vectorial elastic wave-equation with Devito]
 from devito import VectorTimeFunction,
@@
 representation reflects perfectly the mathematics while maintaining
-computational portability and efficiency through the \href{https://github.com/devitocodes/devito}{Devito} compiler. The
+computational portability and efficiency through the \devito compiler. The
 complete generated code for the elastic wave-equation with and without MPI
@@
 wavefield is, as expected, purely acoustic in the water layer
-(\texttt{t\_\{xy\}=0}) and transitions correctly at the ocean bottom to
+($\tau_{xy}=0$) and transitions correctly at the ocean bottom to
 an elastic wavefield. We can also clearly see the shear wave-front
 in the subsurface (at a depth of approximately 1km). Figures \ref{ElasWf}
-and \ref{ElasShot} demonstrate that this high-level \href{https://github.com/devitocodes/devito}{Devito} implementation
+and \ref{ElasShot} demonstrate that this high-level \devito implementation
 of the elastic wave-equation is effective and accurate.  Importantly, in
-constructing this model within the \href{https://github.com/devitocodes/devito}{Devito} DSL framework, computational
+constructing this model within the \devito DSL framework, computational
 technicalities such as the staggered grid analysis are abstracted away. We
@@
 Finally, three dimensional elastic data was modelled in the Cloud to
-demonstrate the scalability of \href{https://github.com/devitocodes/devito}{Devito} to cluster-size problems. The model
+demonstrate the scalability of \devito to cluster-size problems. The model
 used in this experiment mimics a reference model in geophysics known as
@@
 Because earlier performance benchmarks focus mainly on roofline model
-analysis, we also compared the runtime of \href{https://github.com/devitocodes/devito}{Devito} with a reference
+analysis, we also compared the runtime of \devito with a reference
 open source hand-coded propagator in collaboration with its author for
@@
 results and we consistently obtained similar runtimes. This comparison
-illustrates the performance achieved with \href{https://github.com/devitocodes/devito}{Devito} is at least on par with
-hand-coded propagators. Considering we do not take adantage of the Devito
+illustrates the performance achieved with \devito is at least on par with
+hand-coded propagators. Considering we do not take advantage of the Devito
 compiler full capabilities in the two dimensional case, we are confident
 that the generated code will be at least on par with hand-coded version
-for three dimensional probalems and this comparison will be part of future
+for three dimensional problems and this comparison will be part of future
 work.
@@
 The transition from academic size problems, such as the two-dimensional
-acoustic wave-equation, to real-world application can be challenging in
+acoustic wave-equation, to real-world applications can be challenging in
 particular when attempted as an afterthought. In this work, we showed that
 thanks to design principles that aimed at this type of application from
-the beginning, \href{https://github.com/devitocodes/devito}{Devito} provides the scalability necessary. First, we showed
+the beginning, \devito provides the scalability necessary. First, we showed
 that we provide a high-level interface not only for simple scalar
@@
 
-The code to reproduce the different examples presented in this work is available online in the fllwing repositories:
+The code to reproduce the different examples presented in this work is available online in the follwing repositories:
 
@@
   \url{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/elastic/elastic_example.py} 
-  and can be run with any size and spatial order. A standalone script to run the lare 3D elastic modeling
-  is also available at {https://github.com/mloubout/SC20Paper/tree/master/codesamples}
+  and can be run with any size and spatial order. A standalone script to run the large 3D elastic modeling
+  is also available at \url{https://github.com/mloubout/SC20Paper/tree/master/codesamples}
 \end{itemize}

### 2020-04-21 paper.tex

@@
             pdfauthor={},
-            pdftitle={Devito for large scale elastic modelling and anisotropic inversion.},
+            pdftitle={Scaling through abstraction – high-performance vectorial wave simulations for seismic inversion with Devito.},
             colorlinks=true,
@@
 inversion workflow must satisfy the following criteria
-\cite{virieuxmodelling}: - Efficient for multiple-source modeling - The
-memory requirement of the modeling - The ability of a parallel algorithm
+\cite{virieuxmodelling}: - Efficient for multiple-source modelling - The
+memory requirement of the modelling - The ability of a parallel algorithm
 to use an increasing number of processors - Ability of the method to
 process models of arbitrary levels of heterogeneity - Reduce the
-nonlinearity of FWI - Feasibility of the extension of the modeling
+nonlinearity of FWI - Feasibility of the extension of the modelling
 approach to more realistic physical descriptions of the media.
@@
 Devito and its symbolic API and present the distributed memory
-implementation that allows large-scale modeling and inversion by means of
+implementation that allows large-scale modelling and inversion by means of
 domain decomposition. We then provide a brief comparison with a state of
@@
 measurement-based inverse problems. Several additional features are
-supported, among which staggered grids, sub-domains, and stencils with
-custom coefficients. The last major building block of a good PDE solver
-are the boundary conditions, that for finite-difference methods are
-notoriously various and often complicated. The system is, however,
-sufficiently general to express them through composition of core
+supported, among which are staggered grids, sub-domains, and stencils with
+custom coefficients. The last major building block of a solid PDE solver
+are the boundary conditions which for finite-difference methods are
+notoriously diverse and often complicated. The system is, however,
+sufficiently general to express them through a composition of core
 mechanisms. For example, free surface and perfectly-matched layers
@@
 
-It is up to the \devito compiler to translate the symbolic specification
-into C code. The lowering of the input language down to C consists of
-several compilation passes, some of which introducing performance
-optimizations that are the key to fast code. Next to classic stencil
+It is the job of the \devito compiler to translate the symbolic specification
+into C code. The lowering of the input language to C consists of
+several compilation passes, some of which introduce performance
+optimizations that are the key to rapid code. Next to classic stencil
 optimizations (e.g., cache blocking, alignment, SIMD and OpenMP
@@
 
-In this section we illustrate the \devito language showing how to implement
-the acoustic wave-equation in isotropic media
+In this section we illustrate the \devito language by demonstrating
+an implementation of the acoustic wave-equation in isotropic media
 %
@@
 \noindent
-where \texttt{so} is the spatial discretization order and \texttt{to}
-the time discretization order that is used for the generation of the
-finite-difference stencil. Second, we define point-wise objects such as
-point sources \texttt{src}, located at physical coordinates $x_s$, and
-receiver (measurement) objects \texttt{d}, with sensors located at the
-physical locations $x_r$
-
-\begin{python}[label=sparse, caption=SparseFunction deifnition]
+where \texttt{so} is the order of the spatial discretization and \texttt{to}
+the time discretization order used when generating the
+finite-difference stencil. Next, we define point source objects (\texttt{src}) located at the physical
+coordinates $x_r$, and the receiver (measurement) objects (\texttt{d})
+located at the physical locations $x_r$
+
+\begin{python}[label=sparse, caption=SparseFunction definition]
 from devito import SparseTimeFunction
@@
 
-The source term is handled separately from the PDE as a pointwise
-operation called injection, while the measurement is handled with an
+The source term is handled separately from the PDE as a point-wise
+operation called injection, while measurement is handled via
 interpolation. By default, \devito initializes all \texttt{Function} data
@@
 \texttt{t=0}. The isotropic acoustic wave-equation can then be implemented
-in \devito as below
+in \devito as follows
 
@@
 
-The first compilation pass processes the equations individually. The
+The first compilation pass processes equations individually. The
 equations are lowered to an enriched representation, while the
@@
 FLOP-reducing transformations. It is then further lowered into an
-abstract syntax tree, and it is on such representation that parallelisms
+abstract syntax tree, and it is on such representation that parallelism
 is introduced (SIMD, shared-memory, MPI). Finally, all remaining
 low-level aspects of code generation are handled, among which the most
-relevant one is the data management (e.g., definition of variables,
+relevant is data management (e.g., definition of variables,
 transfers between host and device).
@@
 
-The domain decomposition occurs in Python upon creation of a \texttt{Grid}
+Domain decomposition occurs in Python upon creation of a \texttt{Grid}
 object. Exploiting the MPI Cartesian topology abstraction, Devito
 logically splits a grid based on the number of available MPI processes
-(users are given an ``escape hatch'' to override Devito's default
+(noting that users are given an ``escape hatch'' to override Devito's default
 decomposition strategy).  \texttt{Function} and \texttt{TimeFunction}
@@
 data manipulation, Devito
-does not support natively any mechanisms for parallel I/O. The
-distributed NumPy arrays, however, provide a generic and flexible
-infrastructure for the implementation of parallel I/O for any sort of
-file format (e.g., see \cite{witte2018alf}).
+does not natively support any mechanisms for parallel I/O. However, the
+distributed NumPy arrays along with the ability to seamlessly transfer any
+desired slice of data between ranks provides a generic and flexible
+infrastructure for the implementation of any form of parallel I/O
+(e.g., see \cite{witte2018alf}).
 
@@
 
-One of the main applications of seismic finite-difference modeling in
+One of the main applications of seismic finite-difference modelling in
 exploration geophysics is reverse-time migration (RTM), a wave-equation based
@@
 In our seismic imaging case study, we use an anisotropic representation of the
-physics called Tilted Transverse Isotropic (TTI) modeling \cite{thomsen1986}. This
+physics called Tilted Transverse Isotropic (TTI) modelling \cite{thomsen1986}. This
 representation for wave motion is one of the most widely used in exploration
@@
 (c.f. \cite{louboutin2016ppf}). For illustrative purposes, the complete
-generated code for tti modeling with and without MPI is made available
+generated code for tti modelling with and without MPI is made available
 at \href{https://github.com/mloubout/SC20Paper/tree/master/codesamples}{CodeSample}
@@
 
-Because earlier performance benchmarks focus mainly on roofline model
-analysis, we also compared the runtime of \devito with a reference
-open source hand-coded propagator in collaboration with its author for
-completeness. This propagator, described in \cite{thorbecke} is a
-state of the art elastic kernel (Equation~\ref{elas1}). For a fair
-comparison, we ensure, in collaboration with the author of
-\href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} that the
-physical and computational settings were identical. The setting were the
-following:
+Earlier performance benchmarks mainly focused on roofline model analysis.
+In this study, for completeness, the runtime of \devito is therefore
+compared to that of the open source hand-coded propagator
+\href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc}. This propagator,
+described in \cite{thorbecke}, is a state of the art elastic kernel
+(Equation~\ref{elas1}) and the comparisons presented here were carried out
+in collaboration with its author. To ensure a fair
+comparison, we ensured that the
+physical and computational settings were identical. The settings were as
+follows:
 
@@
 
-The runtime obtained for this problem for both propagators were identical
-with less than one percent of difference. This similar runtime were
-obtained both with the Intel compiler and the GNU compiler and we ran the
-experiment with a fourth order and a sixth order discretization.  The
-kernels were executed five time each to ensure consistency between the
-results and we consistently obtained similar runtimes. This comparison
+The runtimes observed for this problem were essentially identical,
+showing less than a one percent of difference. Such similar runtimes were
+obtained with both the Intel and GNU compilers and the
+experiment was performed with both fourth and sixth order discretizations.
+Kernels were executed five times each and the
+runtimes observed were consistently very similar. This comparison
 illustrates the performance achieved with \devito is at least on par with
 hand-coded propagators. Considering we do not take advantage of the Devito
-compiler full capabilities in the two dimensional case, we are confident
-that the generated code will be at least on par with hand-coded version
-for three dimensional problems and this comparison will be part of future
+compilers full capabilities in two dimensional cases, we are confident
+that the code generated will be at least on par with the hand-coded version
+for three dimensional problems and this comparison will be part of our future
 work.
@@
 
-The transition from academic size problems, such as the two-dimensional
-acoustic wave-equation, to real-world applications can be challenging in
-particular when attempted as an afterthought. In this work, we showed that
-thanks to design principles that aimed at this type of application from
-the beginning, \devito provides the scalability necessary. First, we showed
-that we provide a high-level interface not only for simple scalar
-equations but for coupled PDEs and allow the expression of non-trivial
-differential operators in a simple and concise way. Second and most
-importantly, we demonstrated that the compiler enables large-scale with
+Transitioning from academic toy problems, such as the two-dimensional
+acoustic wave-equation, to real-world applications can be challenging,
+particularly if this transition is carried out as an afterthought.
+Owing to the fundamental design principles of \devito such scaling,
+however, becomes trivial. In this work we demonstrated
+the high-level interface provided by \devito not only for simple scalar
+equations but also for coupled PDEs. This interface allows, in a simple, concise
+and consistent manner, the expression of all-kinds of non-trivial
+differential operators. Next, and most
+importantly, we demonstrated that the compiler enables large-scale modelling with
 state-of-the art computational performance and programming paradigm. The
-single-node performance is on par with hand-coded codes while providing
-the flexibility from the symbolic interface, and multi-node parallelism is
-integrated in the compiler and interface in a conservative way. Finally,
-we demonstrated that our abstraction provides to performance portability
-that enables both on-premise HPC and Cloud HPC.
+single-node performance is on par with state of the art hand-coded models, but
+packaged with this performance comes
+the flexibility of the symbolic interface and multi-node parallelism, which is
+integrated in the compiler and interface in a accessible way. Finally,
+we demonstrated that our abstractions provide the necessary portability
+to enable both on-premise and Cloud based HPC.
 
@@
 
-The code to reproduce the different examples presented in this work is available online in the follwing repositories:
+The code to reproduce the different examples presented in this work is available online in the following repositories:
 
@@
   the TTI propagators, the Azure setup for shot parallelism and a documentation.
-  \item The elastic modeling can be run with the elastic example available in Devito at
+  \item The elastic modelling can be run with the elastic example available in Devito at
   \url{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/elastic/elastic_example.py} 
-  and can be run with any size and spatial order. A standalone script to run the large 3D elastic modeling
+  and can be run with any size and spatial order. A standalone script to run the large 3D elastic modelling
   is also available at \url{https://github.com/mloubout/SC20Paper/tree/master/codesamples}

### 2020-04-21 paper.tex
final version

--- 2020-04-21 paper.tex
@@
-\documentclass[conference]{IEEEtran}
+\documentclass[10pt, conference]{IEEEtran}
 \IEEEoverridecommandlockouts
@@
             pdfauthor={},
-            pdftitle={Devito for large scale elastic modelling and anisotropic inversion.},
+            pdftitle={Scaling through abstractions – high-performance vectorial wave simulations for seismic inversion with Devito},
             colorlinks=true,
@@
 
-\title{Scaling through abstraction – high-performance vectorial wave simulations for seismic inversion with Devito}
+\title{Scaling through abstractions – high-performance vectorial wave simulations for seismic inversion with Devito}
 
@@
 \and
-\IEEEauthorblockN{3\textsuperscript{th} Rhodri Nelson}
+\IEEEauthorblockN{3\textsuperscript{rd} Rhodri Nelson}
 \IEEEauthorblockA{\textit{Imperial College London}\\
@@
 \and
-\IEEEauthorblockN{4\textsuperscript{rd} Philipp Witte}
+\IEEEauthorblockN{4\textsuperscript{th} Philipp Witte}
 \IEEEauthorblockA{\textit{Georgia Institute of Technology}\\

### 2020-04-21 paper.tex
final

--- 2020-04-21 paper.tex
@@
 \IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
-Antlanta, GA \\
+Atlanta, GA \\
 mlouboutin3@gatech.edu}
@@
 }
+
+% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
+% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID}
+% \and
+% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID}
+% \and
+% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
+% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID}
+% \and
+% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
+% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID}
+% \and
+% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
+% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID}
+% \and
+% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
+% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID}
+% }
 
@@
 source experiments, computing an elastic synthetic dataset would require 322
-EFLOPs (23k time-steps x 2.8TFLOP/time-step x 500
+EFLOPs (23k time-steps x 2.8TFLOP/time-step x 5000
 sources), or utilizing the full scalability and computing all sources in

### 2020-04-23 paper.tex
Update paper.tex

@@
 specific language (DSL) and compiler for the automatic generation of
-finite-difference stencils, has been developed. Originally deigned to
+finite-difference stencils, has been developed. Originally designed to
 accelerate research and development in exploration geophysics, the
@@
 note that the shot records displayed in Figure~\ref{ElasShot} match the
-original data generated by the creator of this elastic model
-(\cite{marmouelas}).
+original data generated by the creator of this elastic model \cite{marmouelas}.
 

### 2020-05-26 paper.tex
add reviews

@@
 
+\textcolor{red}{Wave-equation based geophysical inversion does not only require fast and accurate wave-equation sovers, 
+but also adavnced algorithms for non-convex optimization for large scale PDE constraints inverse problem.
+The high-level interface \devito provdes enables research and developement for these specific problem in a framework that scales
+to industry size problem automatically through its compiler. Projects using \devito for the wave-equation solves include 
+FWI \cite{witte2018alf}, compressived least-square RTM \cite{witte2018cls} or 
+time-domain WRI \cite{rizzuti2020EAGEtwri, louboutin2020SEGtwri} that tackles the cycle-skipping issue in seismic inversion.}
+
 This paper is organized as follows: first, we provide a brief overview of
@@
 \href{https://github.com/mloubout/SC20Paper/tree/master/codesamples}{CodeSample}
-in \texttt{acou-so8.c}.
+in \texttt{acou-so8.c} \textcolor{red}{and the accuracy of the generated code is verified through our
+continuous integration pipeline against analytical solutions \cite{devito-api}.}
 
@@
 (e.g., see \cite{witte2018alf}).
+
+\textcolor{red}{
+\subsubsection{Strong scaling}
+Strong scaling results from CX2.
+}
 
@@
 
+\textcolor{red}{One of the main challenges in gephysical exploration is to obtain access to reference implementations
+of various wave-equations propagator. While we are in the prcess of comparing hand coded production code from high end contractor companies,
+we here compare the performance to one of the few open source software available. It is extremely important to note that are not only 
+interested in pure stencil framework, such as YASK \cite{yask}, that usually do no handle sparse operation for the source injection 
+and receiver interpolation and are therefore nota dapted to measurement based inverse problems.}
+
 Earlier performance benchmarks mainly focused on roofline model analysis.

### 2020-06-01 response.md
Add draft response.md

Dear Editor,

First of all, we would like to thank the reviewers for their work and constructive feedback. Below, we address all of their questions and observations.

# Review of pap429s2 by Reviewer 1
Detailed Comments for Authors
    This is an interesting paper, well-written, logically ordered and shows interesting results. I only have small questions having to do with scaling results and validation.
Comments for Revision
    If you run your model with a variable number of cores, how is the weak and strong scaling? Can you comment on that? Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality? How is that done?
Scored Review Questions ACCEPT (4)

At the time of the experimentation, we ran into two main issues:
1. Limited resources on Azure, including the access to an Infiniband-enabled system.
2. Most of the MPI-level optimizations that are today in Devito were still in a tuning and consolidation phase.

We subsequently managed to work around point one above; we have published some scaling experiments in [..., page ..., figure ...], although they still don't exploit the latest advancements in Devito v4.2 (e.g., computation/communication overlap). What we want to remark is that our manuscript centers on the development of HPC production-grade wave propagators from a high-level language, and while performance is one of the cornerstones of our work, it was not the objective of this paper to show how powerful Devito is in an extreme scaling setting. As -- we are sure -- you have understood, this is completely sensible for a venue such as SC. We could have added scaling plots showing what, with the latest Devito v4.2, appears to be a remarkable parallel efficiency, but we feel that this would have been premature, and our philosophy is not to publish any results that has not been adequately validated and reproduced over multiple architectures. This attitude is at the hearth of the success of Devito, and it's not our intention to undermine it for a mere publication.

About validation of simulation quality -- [TODO]

# Review of pap429s2 by Reviewer 2
Detailed Comments for Authors
    This paper presents a methodology to use abstractions and compiler technology effectively in the field of exploration seismology. It is a very nicely written paper that puts down all the motivation and technical details quite well. However, it is not very suitable for SC. Parallelization is pretty straightforward, uniform decomposition of relatively straightforward operation. The computational kernels have really high arithmetic intensity that is great for the domain, but not as challenging for optimization, and therefore not very interesting to this community. The parallelism applied so far is straight up MPI which is a bit behind times these days.
Comments for Revision
    I don't think any revision will change the nature of the work, which, as I mentioned in comments section is nicely done but not best suited for SC.
Scored Review Questions REJECT (1)

With all due respect, we strongly disagree with all the critiques moved to the manuscript. Let's go through them one at a time.

* However, it is not very suitable for SC

As partly explained in our reply to R1, this was conceived as an application oriented paper exploiting high-level abstractions and compilation technology. Computational science is evolving rapidly, with production-level DSL-based systems replacing unmaintainable legacy codes, and our work is just yet another small brick in this direction. Unlike the majority of DSL-based tools that often get published, our work builds on and contributes to an established framework for seismic imaging, so we believe that this should be of strong interest for the SC community.

*  Parallelization is pretty straightforward, uniform decomposition of relatively straightforward operation.

First, we observe that this is not just manual implementation of some MPI-based domain decomposition for some pre-existing code. This is rather _generation_ of MPI parallel code from a purely mathematical description of solvers for partial differential equations. The software engineering required is considerable, especially for an open source project such as Devito that was born in academia. We're talking about thousands of lines of carefully engineered code, plus tests and examples, which allow users to just run `mpirun -n X python ...` to run over MPI with virtually no changes to code. Also, as we explained in the paper, it's not grid decomposition only -- sparse functions, distributed numpy arrays, compiler-level optimazion for effective halo exchange -- these are just a few examples of features that are extremely far from being regarded as "pretty straightforward".

* The computational kernels have really high arithmetic intensity that is great for the domain, but not as challenging for optimization

This is simply an error. A high arithmetic intensity does not imply simple optimization. In fact, it's entirely the opposite, and this returns over and over again in computational science. For example, with high-order finite (spectral) elements -- you may get an extremely good OI, but if you're not using tensor-product basis functions and suitable techniques such as sum-factorization, yes, you'll perform loads of flops and obtain a beautiful GFlops/s performance, but unfortunately a huge fraction of those flops will be totally useless, as redundant. The same story happens with our TTI. As we explain in the paper, without having captured all those cross-iteration common sub-expressions,  we would have obtained remarkable GFlops/s performance, but terrible GPoints/s. And capturing those redundancies is extremely difficult -- in fact, this is one of the reasons several companies are today evaluating Devito: machines (compilers) can be better than humans at detecting redundancies, and definitely way faster than them. In the paper we spend a few paragraphs describing how this is achieved. We think this has potentially great interest for the SC community.

* The parallelism applied so far is straight up MPI which is a bit behind times these days.

As far as we know, in 2020 MPI is still the most widely used approach for domain decomposition.

# Review of pap429s2 by Reviewer 3
Detailed Comments for Authors
    The authors in the paper describe their Devito, open-source Python project targeted in the area of seismic exploration. The idea of this particular piece of work is the ease of use through Devito to address, in a Domain Specific Language, creating code for seismic exploration problems which due to being computational intensive often require the use of distributed parallel systems. This is built upon MPI but appears to be hidden from the end user. It was a little confusing but the MPI parallelism appears to essentially be a domain decomposition style parallelism. There was some mention of threads, in particular there seems to be several compilation passes that abstracts out the parallelism which can include SIMD, shared-memory and MPI. Only the MPI is detailed. Through Devito, the authors demonstrate its utility on two problems; (1) Reverse Time Migration (RTM) for a pseudo-acoustic wave problem and (2) an elastic wave problem. With the RTM, they consider titled-transverse isotropic (TTI) wave equation formulation. In RTM, the forward problem is calculated and then the adjoint problem is solve backwards in time to get the imaging. RTM tends to require significant memory and/or fast access to storage of the forward problem during solves for the adjoint problem. Devito is used to set these problems up in VM (Virtual Machines) to be run in a Cloud (Azure) on 32 nodes. If this reviewer understands correctly, Devito is set up to calculate the FLOPs count which is how the authors get the performance. One question that arises, in the performance is how does Devito account for FMA (Fused Multiple Adds)? A comparison was made with an open source hand-code propagator "fdelmode" for the elastic wave equation which seems to be on a 2D problem. The result of this comparison are "essentially identical". A table comparison might be helpful. Scalability performance would also be interesting.
    It appears much of the performance result are provided in the references so it is unclear as to what the purpose of this paper is for an SC audience. For actual seismic surveys, the imaging (interpretation) part is only 10% of the survey. Of course, one does us the RTM in developing the velocity model which adds to complexity in a way. Ease of use coupled with excellent scalable performance while important doesn't address the real "industrial" problems, it helps a little.
Comments for Revision
    Perhaps restricting the discussion of both ease of use of Devito, with a comparison including scalability on one problem, say the 3D TTI problem against a known method and doing this in a cloud setting up to a significant number of nodes would make this a more appropriate paper for the SC audience. This might be too tall an order here. Dropping the discussion on the 3D elastic wave and including a table of comparison with the fdelmode for the 2D elastic wave problem or even just going through scalability comparison on the TTI problem against another code might demonstrate the value of Devito better.
Scored Review Questions WEAK REJECT (2)

# Review of pap429s2 by Reviewer 4 
Detailed Comments for Authors
    This manuscript describes optimizations for Devito, a symbolic DSL designed for geophysics, to extend its use for industrial scale modeling of seismic inversion problems.
    Strengths:
        added support for vector and tensor objects for staggered-grid finite-differences
        demonstrated solving large-scale inverse problems (requiring large memory and computation)
        demonstrated its use on Cloud-based HPC
        paper is well-written
    Weaknesses:
        One issue is that I cannot tell how much of this work is new
        contributions. The paper reads a lot like an overview of previous work in many parts. The parts that I am sure are new work (e.g., adding support for vector and tensors), I am unclear of the amount of effort or innovation involved.
        I like this paper and the work (the ease-of-use of Devito is impressive), but I think for SC we would need more information on its scalability and performance.
Comments for Revision
        Please address the weaknesses above
        Is section III describing new work? It is unclear to me if this is an overview or new work for this paper…
        Is the parallelism just via MPI? What are the possibilities for this code in terms of heterogeneous clusters? Will it be able to take advantage of GPUs?
    Minor issues:
    (1) section 3, 1st paragraph: "must to be" => "must be"
    (2) section 1, 3rd paragraph: make an itemized list?
Scored Review Questions WEAK REJECT (2)

### 2020-06-01 response.md
response headings

@@
 
-# Review of pap429s2 by Reviewer 1
+## Review of pap429s2 by Reviewer 1
 Detailed Comments for Authors
@@
 Scored Review Questions ACCEPT (4)
+
+### Reply
 
@@
 
-# Review of pap429s2 by Reviewer 2
+## Review of pap429s2 by Reviewer 2
 Detailed Comments for Authors
@@
 Scored Review Questions REJECT (1)
+
+### Reply
 
@@
 
-# Review of pap429s2 by Reviewer 3
+## Review of pap429s2 by Reviewer 3
 Detailed Comments for Authors
@@
 
-# Review of pap429s2 by Reviewer 4 
+### Reply
+
+...
+
+## Review of pap429s2 by Reviewer 4 
 Detailed Comments for Authors
@@
 Scored Review Questions WEAK REJECT (2)
+
+### Reply
+
+...

### 2020-06-01 response.md

@@
 
-...
+* One question that arises, in the performance is how does Devito account for FMA (Fused Multiple Adds)?
+
+The operations determining the performance of a code are the source level ones, while FMA is a lower level concept (ie, a scalar FMA would account for two operations).
+This is not Devito, but rather then model adopted everywhere, otherwise two performance values from two different architectures, one with FMAs the other without, would not even be comparable.
+
+* Scalability performance would also be interesting.
+
+[TODO] Reuse part of replies above
+
+*     It appears much of the performance result are provided in the references so it is unclear as to what the purpose of this paper is for an SC audience.
+For actual seismic surveys, the imaging (interpretation) part is only 10% of the survey.
+Of course, one does us the RTM in developing the velocity model which adds to complexity in a way.
+Ease of use coupled with excellent scalable performance while important doesn't address the real "industrial" problems, it helps a little.
+
+[TODO] Felix, Gerard, Mathias...
 

### 2020-06-01 response.md

@@
 
-About validation of simulation quality -- [TODO]
+About validation of simulation quality -- we did not include it in this paper as we think it is outside of the scope of the presented results but the simulation quality is verified in two ways.
+As described in [@devito-api] The numerical accuracy is compared to the analytical solution for the acoustic wave equation as well as the convergence rate of the finite-difference discretization.
+The verification of the TTI equation is not as simple as we  do not have an analytical solution but the implementation of the method of manufactured solution is in progress.
+All parts such as finite-difference accuracy, comparison with isotropicr esult for zeros anisotropic parameters are part of the continuous integration.
 
@@
 
-[TODO] Felix, Gerard, Mathias...
+The industrial problem does indeed involve muliple parts, such as data acquisition then processing, and mst importatly velocity model building.
+We do not claim here to solve all the problems, but demonstrate that the combination of a high-elvel DSL and a highly efficient compiler provides the necessary tool for sesimic modeling and imaging at (industrial) scale.
+The complexity of model building requires advance optimization methods, and the easy access to computationally and easy to develop wave-equation solvers enable research and development for seismic inversion.
+Devito is therfore used for the development of inital background model building [@mojica2019tab] or cycle skipping mitigating methods [@rizzuti2020EAGEtwri; @louboutin2020SEGtwri] and compressive least square migration [@witte2018cls] based on a linear algebra framework for seismic inversion built on top of devito [@witte2018alf]
+We added a clarification in the introduction.
 
@@
 
+* One issue is that I cannot tell how much of this work is new
+
 ...
+
+*   but I think for SC we would need more information on its scalability and performance.
+
+See previous responses
+
+*    Is section III describing new work? It is unclear to me if this is an overview or new work for this paper…
+
+Yes. We have refined the txt to make it more explicit and detailed.
+
+*    Is the parallelism just via MPI?
+
+The domain decomposition is via MPI, however, the overall task parallelism is not.
+Seismic inversion is a fullys eparable problem on terms of surce experiment, and the paralleization over these sources is done using `batch-shipyard` in a serverless settting.
+THis allow optimal usage of Cloud ressource rather than firing up a cluster aht will be IDLE most of time but still payed for bytt he hour in the cloud.
+
+*   What are the possibilities for this code in terms of heterogeneous clusters? Will it be able to take advantage of GPUs?
+
+GPU support is in development and currently supported for some applications. Its capabilities and related applications will be fully covered in a future dedicated paper.

### 2020-06-01 paper.tex
rhodrin mods + tidying.

--- 2020-04-23 paper.tex
@@
 
+\textcolor{red}{It should be emphasized that unlike in,
+e.g., earthquake simulation models \cite{fu201718} (more?), where
+a single shot is computed, wave-equation based
+geophysical inversion not only requires
+fast and accurate solvers but also adavnced algorithms for
+non-convex optimization of large scale constrained inverse problems. The 
+high-level interface provided by \devito enables research and developement for these 
+specific problem in a framework that scales to industry size problem 
+automatically through its compiler. \devito has been used previously
+for the solution of the wave equation within imaging frameworks including
+FWI \cite{witte2018alf}, compressived least-square RTM \cite{witte2018cls} and
+time-domain Wavefield Reconstruction Inversion (WRI)~\cite{rizzuti2020EAGEtwri, 
+louboutin2020SEGtwri}, the later of which takles the cycle-skipping issue
+present in seismic inversion.}
+
 Previous work focused on the DSL and compiler to highlight the potential
@@
 benchmarked with the roofline model (\cite{patterson, devito-compiler,
-devito-api, louboutin2016ppf}).  Before concluding, results from the
-Cloud-based experiments discussed above are presented, highlighting the
-vectorial and tensorial capabilities of Devito.
+devito-api, louboutin2016ppf}). \textcolor{red}{Before concluding, results from the
+Cloud-based experiments introduced above are presented}.
 
@@
 sufficiently general to express them through a composition of core
-mechanisms. For example, free surface and perfectly-matched layers
-(PMLs) boundary conditions can be expressed as equations -- just like
+mechanisms. For example, free surface and perfectly-matched \textcolor{red}{layer
+(PML) conditions} can be expressed as equations -- just like
 any other PDE equations -- over a suitable sub-domain.
@@
 \begin{cases}
- m \frac{d^2 u(t, x)}{dt^2} - \Delta u(t, x) = \delta(xs) q(t) \\
- u(0, .) = \frac{d u(t, x)}{dt}(0, .) = 0 \\
+ m \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(xs) q(t) \\
+ u(0, .) = \frac{\partial u(t, x)}{\partial t}(0, .) = 0 \\
  d(t, .) = u(t, xr).
@@
 %
+\textcolor{red}{FIXME: Define the terms appearing in the above equation.}
  The core of the Devito
@@
 
-A \texttt{Grid} represents a discretized finite n-dimensional space and
+A \texttt{Grid} represents a discretized finite $n$-dimensional space and
 is created as follows
@@
 points in each direction,
-\texttt{(ext\_x,\phantom{\ }ext\_y,\phantom{\ }ext\_z)} is the physical
+\texttt{(ext\_x,\phantom{\ }ext\_y,\phantom{\ }ext\_z)} is the
 extent of the domain in physical units (i.e \texttt{m}) and
 \texttt{(o\_x,\phantom{\ }o\_y,\phantom{\ }o\_z)} is the origin of the
-domain in the same physical units. The object \texttt{grid} contains all
-the information related to the discretization such as the grid spacing.
+domain in the same physical units. \textcolor{red}{The \texttt{grid} object contains all
+information pertinent to the discretization such as the grid spacing}.
 We use \texttt{grid} to create the symbolic objects that will be used to
@@
   G_{\bar{x}\bar{x}} &= D_{\bar{x}}^T D_{\bar{x}} \\
-  D_{\bar{x}} &= \cos(\mathbf{\theta})\cos(\mathbf{\phi})\frac{\mathrm{d}}{\mathrm{d}x} + \cos(\mathbf{\theta})\sin(\mathbf{\phi})\frac{\mathrm{d}}{\mathrm{d}y} - \sin(\mathbf{\theta})\frac{\mathrm{d}}{\mathrm{d}z}.
+  D_{\bar{x}} &= \cos(\mathbf{\theta})\cos(\mathbf{\phi})\frac{\partial}{\partial x} + \cos(\mathbf{\theta})\sin(\mathbf{\phi})\frac{\partial}{\partial y} - \sin(\mathbf{\theta})\frac{\partial}{\partial z}.
 \end{aligned}
@@
 \begin{aligned}
-&\frac{1}{\rho}\frac{dv}{dt} = \nabla . \tau \\
-&\frac{d \tau}{dt} = \lambda \mathrm{tr}(\nabla v) \mathbf{I}  + \mu (\nabla v + (\nabla v)^T)
+&\frac{1}{\rho}\frac{\partial v}{\partial t} = \nabla . \tau \\
+&\frac{\partial \tau}{\partial t} = \lambda \mathrm{tr}(\nabla v) \mathbf{I}  + \mu (\nabla v + (\nabla v)^T)
 \end{aligned}
@@
 %
- The discretization of \eqref{elas1} and \eqref{stress} requires five
-equations in two dimensions (two equations for the particle velocity and
-three for the stress) and nine equations in three dimensions (three for the particle velocity
-and six for the stress). However, the mathematical
+ Note that the discretization of \eqref{elas1} and \eqref{stress} requires five
+scalar equations in two dimensions (two for the particle velocity and
+three for stress) and nine equations in three dimensions (three for particle velocity
+and six for stress). However, the mathematical
 definition only require two coupled vector/tensor-valued equations for
@@
 The extended \devito language introduces two new types, 
-\texttt{VectorFunction}/\texttt{VectorTimeFunction} for vectorial objects such as the particle velocity, and
+\texttt{VectorFunction}/\texttt{VectorTimeFunction} for vectorial objects
+such as the particle velocity, and
 \texttt{TensorFunction}/\texttt{TensorTimeFunction} for second-order
@@
 
-Earlier performance benchmarks mainly focused on roofline model analysis.
+Earlier performance benchmarks mainly focused on roofline model analysis
+\textcolor{red}{[TODO: Add reference]}.
 In this study, for completeness, the runtime of \devito is therefore
@@
 (Equation~\ref{elas1}) and the comparisons presented here were carried out
-in collaboration with its author. To ensure a fair
-comparison, we ensured that the
+in collaboration with its author.
+\textcolor{red}{It should be emphasized that direct model comparison in
+the area of exploration geophysics is challenging. The majority of production
+codes are either proprioty or are simply not publicly available. Further,
+\devito should not be compared against pure stencil frameworks,
+such as YASK \cite{yask}, since they are unable to compute the sparse operations
+required for source injection and receiver interpolation (and are therefore not
+suitable for use in measurement based inversion problems). The comparison carried out
+here is conducted in two dimensions. Whilst large scale 3D scaling comparisons are theoretically
+possible, compute time on the Cloud is expensive. Such a comparison is not feasible at this
+time and will therefore be left for future work.}
+
+To ensure a fair comparison, we ensured that the
 physical and computational settings were identical. The settings were as
@@
 hand-coded propagators. Considering we do not take advantage of the Devito
-compilers full capabilities in two dimensional cases, we are confident
+compilers full capabilities in \textcolor{red}{the two dimensional case}, we are confident
 that the code generated will be at least on par with the hand-coded version
@@
 differential operators. Next, and most
-importantly, we demonstrated that the compiler enables large-scale modelling with
+importantly, \textcolor{red}{it was} demonstrated that the compiler enables large-scale modelling with
 state-of-the art computational performance and programming paradigm. The

### 2020-06-01 paper.tex
add refs and FIXME

@@
 \textcolor{red}{It should be emphasized that unlike in,
-e.g., earthquake simulation models \cite{fu201718} (more?), where
+e.g., earthquake simulation models \cite{fu201718}, where
 a single shot is computed, wave-equation based
@@
 \begin{cases}
- m \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(xs) q(t) \\
+ m(x) \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(xs) q(t) \\
  u(0, .) = \frac{\partial u(t, x)}{\partial t}(0, .) = 0 \\
@@
 %
-\textcolor{red}{FIXME: Define the terms appearing in the above equation.}
+\textcolor{red}{where $u(t, x)$ is the pressure wavefield, $xs$ is the point source position, 
+$q(t)$ is thes ource time signature, $d(t, .)$ is the measured data at positions $xr$ and $m(x)$ is the 
+squared slowness.}
  The core of the Devito
@@
 for thorough explanations and practical examples.
+\textcolor{red}{We highlugh the main feature of the distributed framework tha enable the industry scale
+applications we describe in this paper. A more thorough study of its implementation, including `mpi4numpy' and
+a complete scaling analysis will be presented in a later paper. Initial results assciated with early version of \devito's
+distributed computing capabilities can be found in \cite{luporini2019adp} on a conventional cluster and 
+in \cite{witte2019ecl} that contains results on AWS.}
 
@@
 Earlier performance benchmarks mainly focused on roofline model analysis
-\textcolor{red}{[TODO: Add reference]}.
+\textcolor{red}{\cite{sevendwarfs, patterson, williams2009roofline, louboutin2016ppf}}.
 In this study, for completeness, the runtime of \devito is therefore

### 2020-06-01 paper.tex
add refs and FIXME

@@
 for thorough explanations and practical examples.
-\textcolor{red}{We highlugh the main feature of the distributed framework tha enable the industry scale
-applications we describe in this paper. A more thorough study of its implementation, including `mpi4numpy' and
-a complete scaling analysis will be presented in a later paper. Initial results assciated with early version of \devito's
-distributed computing capabilities can be found in \cite{luporini2019adp} on a conventional cluster and 
-in \cite{witte2019ecl} that contains results on AWS.}
 

### 2020-06-02 response.md
Update response.md

@@
 ## Review of pap429s2 by Reviewer 1
-Detailed Comments for Authors
-    This is an interesting paper, well-written, logically ordered and shows interesting results. I only have small questions having to do with scaling results and validation.
-Comments for Revision
-    If you run your model with a variable number of cores, how is the weak and strong scaling?
+
+> If you run your model with a variable number of cores, how is the weak and strong scaling?
 Can you comment on that?
@@
 How is that done?
-Scored Review Questions ACCEPT (4)
-
-### Reply
 
@@
 1. Limited resources on Azure, including the access to an Infiniband-enabled system.
-2. Most of the MPI-level optimizations that are today in Devito were still in a tuning and consolidation phase.
+2. Most of the MPI-level optimizations that are today in Devito still needed tuning and consolidation.
 
-We subsequently managed to work around point one above; we have published some scaling experiments in [..., page ..., figure ...], although they still don't exploit the latest advancements in Devito v4.2 (e.g., computation/communication overlap).
-What we want to remark is that our manuscript centers on the development of HPC production-grade wave propagators from a high-level language, and while performance is one of the cornerstones of our work, it was not the objective of this paper to show how powerful Devito is in an extreme scaling setting.
-As -- we are sure -- you have understood, this is completely sensible for a venue such as SC.
-We could have added scaling plots showing what, with the latest Devito v4.2, appears to be a remarkable parallel efficiency, but we feel that this would have been premature, and our philosophy is not to publish any results that has not been adequately validated and reproduced over multiple architectures.
-This attitude is at the hearth of the success of Devito, and it's not our intention to undermine it for a mere publication.
+We subsequently managed to work around point one above.
+Initial scaling experiments have been published in [..., page ..., figure ...], although this work still doesn't exploit the latest advancements in Devito v4.2 (e.g., computation/communication overlap).
+Our manuscript centers on the development of HPC production-grade wave propagators from a high-level language, and while performance is one of the cornerstones of our work, showing off Devito at extreme scale was not an objective.
+We have preliminary evidence that the latest Devito v4.2 has remarkable parallel efficiency, but we felt that publishing these results today would be premature -- we at least need testing and reproducibility on several architectures and applications.
 
-About validation of simulation quality -- we did not include it in this paper as we think it is outside of the scope of the presented results but the simulation quality is verified in two ways.
-As described in [@devito-api] The numerical accuracy is compared to the analytical solution for the acoustic wave equation as well as the convergence rate of the finite-difference discretization.
-The verification of the TTI equation is not as simple as we  do not have an analytical solution but the implementation of the method of manufactured solution is in progress.
-All parts such as finite-difference accuracy, comparison with isotropicr esult for zeros anisotropic parameters are part of the continuous integration.
+About validation of simulation quality -- we did not include it in this paper as we think it is outside of the scope of the presented results, but the simulation quality is verified in two ways.
+As described in [@devito-api], the numerical accuracy is compared to the analytical solution for the acoustic wave equation as well as the convergence rate of the finite-difference discretization.
+The verification of TTI is not as simple as we do not have an analytical solution; we will be using the method of manufactured solution (MMS), but this is still work in progress.
+However, all of the key parts, such as finite-difference accuracy and comparison with isotropic results for zero-ed anisotropic parameters are implemented and part of our continuous integration system.
 
 ## Review of pap429s2 by Reviewer 2
-Detailed Comments for Authors
-    This paper presents a methodology to use abstractions and compiler technology effectively in the field of exploration seismology.
-It is a very nicely written paper that puts down all the motivation and technical details quite well.
-However, it is not very suitable for SC.
-Parallelization is pretty straightforward, uniform decomposition of relatively straightforward operation.
-The computational kernels have really high arithmetic intensity that is great for the domain, but not as challenging for optimization, and therefore not very interesting to this community.
-The parallelism applied so far is straight up MPI which is a bit behind times these days.
-Comments for Revision
-    I don't think any revision will change the nature of the work, which, as I mentioned in comments section is nicely done but not best suited for SC.
-Scored Review Questions REJECT (1)
-
-### Reply
-
-With all due respect, we strongly disagree with all the critiques moved to the manuscript. Let's go through them one at a time.
 
@@
 
-As partly explained in our reply to R1, this was conceived as an application oriented paper exploiting high-level abstractions and compilation technology.
-Computational science is evolving rapidly, with production-level DSL-based systems replacing unmaintainable legacy codes, and our work is just yet another small brick in this direction.
-Unlike the majority of DSL-based tools that often get published, our work builds on and contributes to an established framework for seismic imaging, so we believe that this should be of strong interest for the SC community.
+As partly explained in our reply to R1, we conceived this work as an application-oriented paper based on high-level abstractions and compilation technology.
+We're in the middle of an epochal change for computational science, with legacy codes being superseded, gradually, by DSLs; our work is just yet another small step in this direction.
+However, unlike many other works centered on DSLs, ours builds on -- and contributes to -- what is nowadays an established framework for seismic imaging.
+We believe that this should be of strong interest for the SC community, especially since -- we believe -- the methodology and philosophy underpinning Devito is reusable in computational science.
 
@@
 
-First, we observe that this is not just manual implementation of some MPI-based domain decomposition for some pre-existing code.
-This is rather _generation_ of MPI parallel code from a purely mathematical description of solvers for partial differential equations.
-The software engineering required is considerable, especially for an open source project such as Devito that was born in academia.
-We're talking about thousands of lines of carefully engineered code, plus tests and examples, which allow users to just run `mpirun -n X python ...` to run over MPI with virtually no changes to code.
-Also, as we explained in the paper, it's not grid decomposition only -- sparse functions, distributed numpy arrays, compiler-level optimazion for effective halo exchange -- these are just a few examples of features that are extremely far from being regarded as "pretty straightforward".
+We remark that our work does _not_ describe some ad-hoc MPI-based domain decomposition; what Devito can do is _generation_ of MPI parallel code from a mathematical language for solving partial differential equations.
+This requires considerable software engineering (especially for an open source project such as Devito that was born in academia): we're talking about thousands of lines of carefully engineered code, plus tests and examples, to enable users to just type `mpirun -n X python ...` and get their code to run on a distributed-memory system with virtually no changes to their code.
+Secondly, as we explained in the paper [section ...], it's not just "classic domain decomposition": Devito orchestrates sparse functions, implements distributed numpy arrays, and performs several compiler-level optimazions for effective halo exchange.
+Based on our experience, this is a fairly sophisticated system, quite far from being "straightforward".
 
@@
 
-This is simply an error.
-A high arithmetic intensity does not imply simple optimization.
-In fact, it's entirely the opposite, and this returns over and over again in computational science.
-For example, with high-order finite (spectral) elements -- you may get an extremely good OI, but if you're not using tensor-product basis functions and suitable techniques such as sum-factorization, yes, you'll perform loads of flops and obtain a beautiful GFlops/s performance, but unfortunately a huge fraction of those flops will be totally useless, as redundant.
+High arithmetic intensity does not imply "non challenging optimization".
+In fact, it's the exact opposite, and there are several examples in computational science that we could use as evidence, even outside of seismic.
+Take high-order finite (spectral) elements for instance -- you may get an extremely good arithmetic intensity, but without tensor-product basis functions and suitable optimizations such as sum-factorization you'll likely end up with a notable, yet totally misleasing, GFlops/s performance, as a huge fraction of flops will stem from redundant computation (e.g., common sub-expressions, lack of code hoisting, distributivity).
 The same story happens with our TTI.
-As we explain in the paper, without having captured all those cross-iteration common sub-expressions,  we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
-And capturing those redundancies is extremely difficult -- in fact, this is one of the reasons several companies are today evaluating Devito: machines (compilers) can be better than humans at detecting redundancies, and definitely way faster than them.
-In the paper we spend a few paragraphs describing how this is achieved.
-We think this has potentially great interest for the SC community.
+As we explain in the paper [section ...], without capturing the cross-iteration common sub-expressions (due to high order derivatives), we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
+And capturing such redundancies is indeed quite challenging.
+In fact, this is one of the reasons several companies are today so much interested in Devito: machines (compilers) can be better than humans at detecting redundancies, and definitely way faster than them.
 

### 2020-06-02 response.md
Update response.md

@@
 
-* However, it is not very suitable for SC
+> However, it is not very suitable for SC
 
@@
 
-*  Parallelization is pretty straightforward, uniform decomposition of relatively straightforward operation.
+> Parallelization is pretty straightforward, uniform decomposition of relatively straightforward operation.
 
@@
 
-* The computational kernels have really high arithmetic intensity that is great for the domain, but not as challenging for optimization
+> The computational kernels have really high arithmetic intensity that is great for the domain, but not as challenging for optimization
 
@@
 
-* The parallelism applied so far is straight up MPI which is a bit behind times these days.
+> The parallelism applied so far is straight up MPI which is a bit behind times these days.
 

### 2020-06-02 response.md
Update response.md

@@
 
-* One question that arises, in the performance is how does Devito account for FMA (Fused Multiple Adds)?
+> One question that arises, in the performance is how does Devito account for FMA (Fused Multiple Adds)?
 
@@
 
-* Scalability performance would also be interesting.
+> Scalability performance would also be interesting.
 
-[TODO] Reuse part of replies above
+See reply to R1 -- basically, (i) lack of resources and (ii) Devito not ready yet for extreme scale at the time we ran the experiments.
 
-*     It appears much of the performance result are provided in the references so it is unclear as to what the purpose of this paper is for an SC audience.
-For actual seismic surveys, the imaging (interpretation) part is only 10% of the survey.
+> It appears much of the performance result are provided in the references so it is unclear as to what the purpose of this paper is for an SC audience.
+
+We are showing performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations.
+These results are, respectively, in sections [...].
+As we explain in our reply to R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
+To achieve this, we're using several bleeding edge technologies -- an established DSL-based system such as Devito (used in companies and academia), the Azure cloud, Docker (for ease of installation, use and reproducibility), the rapidly-spreading GitHub Actions for continuos integration, and much more.
+The use and orchestration of these technologies is thouroughly described in the paper, and this should be, in our opinion, of strong interest for the SC audience.
+
+> For actual seismic surveys, the imaging (interpretation) part is only 10% of the survey.
 Of course, one does us the RTM in developing the velocity model which adds to complexity in a way.
@@
 
-The industrial problem does indeed involve muliple parts, such as data acquisition then processing, and mst importatly velocity model building.
-We do not claim here to solve all the problems, but demonstrate that the combination of a high-elvel DSL and a highly efficient compiler provides the necessary tool for sesimic modeling and imaging at (industrial) scale.
-The complexity of model building requires advance optimization methods, and the easy access to computationally and easy to develop wave-equation solvers enable research and development for seismic inversion.
-Devito is therfore used for the development of inital background model building [@mojica2019tab] or cycle skipping mitigating methods [@rizzuti2020EAGEtwri; @louboutin2020SEGtwri] and compressive least square migration [@witte2018cls] based on a linear algebra framework for seismic inversion built on top of devito [@witte2018alf]
-We added a clarification in the introduction.
+The industrial problem does indeed involve muliple parts, such as data acquisition, processing, and, most importantly, velocity model building.
+We do not claim here to solve all these problems; however, we demonstrate that the combination of advanced technolgies such as high-level DSLs and compilers is key to highly efficient sesimic modeling and imaging at industrial scale.
+The complexity of model building requires advanced optimization methods.
+The straightforward access to computationally and easy-to-develop wave-equation solvers enable research and development for seismic inversion.
+Devito is therfore used for the development of inital background model building [@mojica2019tab] or cycle skipping mitigating methods [@rizzuti2020EAGEtwri; @louboutin2020SEGtwri] and compressive least square migration [@witte2018cls] based on a linear algebra framework for seismic inversion built on top of devito [@witte2018alf].
+We have added a clarification in the introduction.
 

### 2020-06-02 paper.tex
Typos

@@
 for thorough explanations and practical examples.
+\textcolor{red}{We highlight the main features of the distributed framework tha enable the industry scale
+applications we describe in this paper. A more thorough study of its implementation, including `mpi4py' and
+a complete scaling analysis will be presented in a later paper. Initial results associated with early version of \devito's
+distributed computing capabilities can be found in \cite{luporini2019adp} on a conventional cluster and 
+in \cite{witte2019ecl} that contains results on AWS.}
 
@@
 the area of exploration geophysics is challenging. The majority of production
-codes are either proprioty or are simply not publicly available. Further,
+codes are either proprietary or are simply not publicly available. Further,
 \devito should not be compared against pure stencil frameworks,

### 2020-06-02 response.md
Update response.md

@@
 
-## Review of pap429s2 by Reviewer 1
+## R1
 
@@
 
-## Review of pap429s2 by Reviewer 2
+## R2
 
@@
 
-As far as we know, in 2020 MPI is still the most widely used approach for domain decomposition.
+In 2020 MPI is still de facto a widely adopted approach for domain decomposition.
 
-## Review of pap429s2 by Reviewer 3
-Detailed Comments for Authors
-    The authors in the paper describe their Devito, open-source Python project targeted in the area of seismic exploration.
-The idea of this particular piece of work is the ease of use through Devito to address, in a Domain Specific Language, creating code for seismic exploration problems which due to being computational intensive often require the use of distributed parallel systems.
-This is built upon MPI but appears to be hidden from the end user.
-It was a little confusing but the MPI parallelism appears to essentially be a domain decomposition style parallelism.
-There was some mention of threads, in particular there seems to be several compilation passes that abstracts out the parallelism which can include SIMD, shared-memory and MPI.
+## R3
+
+> It was a little confusing but the MPI parallelism appears to essentially be a domain decomposition style parallelism.
+
+As we elaborate in R2 (with references to the paper), it's not just about decomposing a grid.
+Devito deals with distribution of sparse functions, python-level distributed arrays, decomposition over sub-domains and much more.
+
+> There was some mention of threads, in particular there seems to be several compilation passes that abstracts out the parallelism which can include SIMD, shared-memory and MPI.
 Only the MPI is detailed.
-Through Devito, the authors demonstrate its utility on two problems; (1) Reverse Time Migration (RTM) for a pseudo-acoustic wave problem and (2) an elastic wave problem.
-With the RTM, they consider titled-transverse isotropic (TTI) wave equation formulation.
-In RTM, the forward problem is calculated and then the adjoint problem is solve backwards in time to get the imaging.
-RTM tends to require significant memory and/or fast access to storage of the forward problem during solves for the adjoint problem.
-Devito is used to set these problems up in VM (Virtual Machines) to be run in a Cloud (Azure) on 32 nodes.
-If this reviewer understands correctly, Devito is set up to calculate the FLOPs count which is how the authors get the performance.
-One question that arises, in the performance is how does Devito account for FMA (Fused Multiple Adds)?
-A comparison was made with an open source hand-code propagator "fdelmode" for the elastic wave equation which seems to be on a 2D problem.
-The result of this comparison are "essentially identical".
-A table comparison might be helpful.
-Scalability performance would also be interesting.
-    It appears much of the performance result are provided in the references so it is unclear as to what the purpose of this paper is for an SC audience.
-For actual seismic surveys, the imaging (interpretation) part is only 10% of the survey.
-Of course, one does us the RTM in developing the velocity model which adds to complexity in a way.
-Ease of use coupled with excellent scalable performance while important doesn't address the real "industrial" problems, it helps a little.
-Comments for Revision
-    Perhaps restricting the discussion of both ease of use of Devito, with a comparison including scalability on one problem, say the 3D TTI problem against a known method and doing this in a cloud setting up to a significant number of nodes would make this a more appropriate paper for the SC audience.
-This might be too tall an order here.
-Dropping the discussion on the 3D elastic wave and including a table of comparison with the fdelmode for the 2D elastic wave problem or even just going through scalability comparison on the TTI problem against another code might demonstrate the value of Devito better.
-Scored Review Questions WEAK REJECT (2)
 
-### Reply
+Correct. We refer to other papers for all these other aspects.
 
@@
 
-> For actual seismic surveys, the imaging (interpretation) part is only 10% of the survey.
-Of course, one does us the RTM in developing the velocity model which adds to complexity in a way.
-Ease of use coupled with excellent scalable performance while important doesn't address the real "industrial" problems, it helps a little.
+> For actual seismic surveys, the imaging (interpretation) part is only 10% of the survey. [...] while important doesn't address the real "industrial" problems, it helps a little.
 
@@
 
-## Review of pap429s2 by Reviewer 4 
+> Dropping the discussion on the 3D elastic wave and [...] might demonstrate the value of Devito better.
+
+We believe that showing the new capabilities of Devito -- the tensor language -- for a different physics and discretization is as important to the SC audience as a strong scaling experiment.
+In fact, strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:
+
+* even with high-frequency FWI, a typical shot-level computation requires no more than a bunch of nodes (often order of units);
+* there is an outer-level of parallelism for inversion, essentially a farm (master-slave) where the individual workers and internally MPI-parallel.
+
+## R4 
 Detailed Comments for Authors

### 2020-06-02 response.md
Update response.md

@@
 ## R4
-Detailed Comments for Authors
-    This manuscript describes optimizations for Devito, a symbolic DSL designed for geophysics, to extend its use for industrial scale modeling of seismic inversion problems.
-    Strengths:
-        added support for vector and tensor objects for staggered-grid finite-differences
-        demonstrated solving large-scale inverse problems (requiring large memory and computation)
-        demonstrated its use on Cloud-based HPC
-        paper is well-written
-    Weaknesses:
-        One issue is that I cannot tell how much of this work is new
-        contributions.
-The paper reads a lot like an overview of previous work in many parts.
-The parts that I am sure are new work (e.g., adding support for vector and tensors), I am unclear of the amount of effort or innovation involved.
-        I like this paper and the work (the ease-of-use of Devito is impressive), but I think for SC we would need more information on its scalability and performance.
-Comments for Revision
-        Please address the weaknesses above
-        Is section III describing new work? It is unclear to me if this is an overview or new work for this paper…
-        Is the parallelism just via MPI? What are the possibilities for this code in terms of heterogeneous clusters? Will it be able to take advantage of GPUs?
-    Minor issues:
-    (1) section 3, 1st paragraph: "must to be" => "must be"
-    (2) section 1, 3rd paragraph: make an itemized list?
-Scored Review Questions WEAK REJECT (2)
-
-### Reply
 
@@
 
-...
+We have improved the manuscript to clarify what the contributions are. These are also now reiterated throughout the whole paper. TODO -- need to add pointers to what we added?
+
+* [...] vector and tensors [...] I am unclear of the amount of effort or innovation involved.
+
+The number of additions to the codebase is notable: https://github.com/devitocodes/devito/pull/873
+And for Devito and his users, this was an invaluable addition.
 
@@
 
-See previous responses
+The single-node performance is discussed in other (cited) articles. In the previous responses (R1, R2, R3), we have elaborated why the lack of scalability experiments.
 
@@
 
-Yes. We have refined the txt to make it more explicit and detailed.
+Yes. We have improved the text to make it more explicit and detailed.
 
@@
 
-The domain decomposition is via MPI, however, the overall task parallelism is not.
-Seismic inversion is a fullys eparable problem on terms of surce experiment, and the paralleization over these sources is done using `batch-shipyard` in a serverless settting.
-THis allow optimal usage of Cloud ressource rather than firing up a cluster aht will be IDLE most of time but still payed for bytt he hour in the cloud.
+The inversion (outer-level of parallelism) requires what is essentially task parallelism; each task is internally MPI-parallel (domain decomposition).
+In this context task-parallelism means paralleization over sources, which we have accomplished using `batch-shipyard` in a serverless settting (see Section TODO).
+We remark that this allows optimal usage of the typical cloud infrastructure, where one pays based on actual usage of computational resources.
 
@@
 
-GPU support is in development and currently supported for some applications. Its capabilities and related applications will be fully covered in a future dedicated paper.
+GPU support is in development and currently (Devito v4.2) supported for some applications. Its capabilities and related applications will be fully covered in a future dedicated paper.
+
+* Minor issues [...]
+
+Fixed, thanks.

### 2020-06-02 response.md
more rephrasing

@@
 The operations determining the performance of a code are the source level ones, while FMA is a lower level concept (ie, a scalar FMA would account for two operations).
-This is not Devito, but rather then model adopted everywhere, otherwise two performance values from two different architectures, one with FMAs the other without, would not even be comparable.
+This is not Devito, but rather the model adopted everywhere, otherwise two performance values from two different architectures, one with FMAs the other without, would not even be comparable.
 
@@
 * even with high-frequency FWI, a typical shot-level computation requires no more than a bunch of nodes (often order of units);
-* there is an outer-level of parallelism for inversion, essentially a farm (master-slave) where the individual workers and internally MPI-parallel.
+* there is an outer-level of parallelism for inversion, essentially a task farm where the individual workers is internally MPI-parallel.
 
@@
 
-* One issue is that I cannot tell how much of this work is new
+> One issue is that I cannot tell how much of this work is new
 
@@
 
-* [...] vector and tensors [...] I am unclear of the amount of effort or innovation involved.
+> [...] vector and tensors [...] I am unclear of the amount of effort or innovation involved.
 
@@
 
-* but I think for SC we would need more information on its scalability and performance.
+> but I think for SC we would need more information on its scalability and performance.
 
@@
 
-* Is section III describing new work? It is unclear to me if this is an overview or new work for this paper…
+> Is section III describing new work? It is unclear to me if this is an overview or new work for this paper…
 
@@
 
-* Is the parallelism just via MPI?
+> Is the parallelism just via MPI?
 
 The inversion (outer-level of parallelism) requires what is essentially task parallelism; each task is internally MPI-parallel (domain decomposition).
-In this context task-parallelism means paralleization over sources, which we have accomplished using `batch-shipyard` in a serverless settting (see Section TODO).
+In this context task-parallelism means paralleization over sources, which we have accomplished using `batch-shipyard` in a serverless setting (see Section II end).
 We remark that this allows optimal usage of the typical cloud infrastructure, where one pays based on actual usage of computational resources.
 
-* What are the possibilities for this code in terms of heterogeneous clusters? Will it be able to take advantage of GPUs?
+> What are the possibilities for this code in terms of heterogeneous clusters? Will it be able to take advantage of GPUs?
 
@@
 
-* Minor issues [...]
+> Minor issues [...]
 

### 2020-06-02 paper.tex
more rephrasing

--- 2020-06-01 paper.tex
@@
 
+\usepackage[normalem]{ulem}
 \usepackage{color}
@@
 This paper is organized as follows: first, we provide a brief overview of
-Devito and its symbolic API and present the distributed memory
+Devito and its symbolic API and present the \textcolor{red}{newly developped} distributed memory
 implementation that allows large-scale modelling and inversion by means of
@@
 for thorough explanations and practical examples.
-\textcolor{red}{We highlugh the main feature of the distributed framework tha enable the industry scale
+\textcolor{red}{We highlight the main features of the distributed framework tha enable the industry scale
 applications we describe in this paper. A more thorough study of its implementation, including `mpi4numpy' and
@@
 important in global seismology since shear surface waves (which are
-ignored in acoustic models) are the most hazardous. In this section, we
+ignored in acoustic models) are the most hazardous. \textcolor{red}{In this section, we 
+introduce the new tensor algebra language implemented in \devito that enables the simple and 
+high-level definition of vectorial and tensorial physics. This API extension also implements 
+fully automatized staggered grid and staggered ginite-difference generally required for a stable discretization
+of vectorial PDEs. This extension allows us to express an elastic model with compact and elegant notation 
+and model its solution computationally efficiently with the \devito compiler}
+
+\sout{\textcolor{green}{In this section, we
 exploit the tensor algebra language introduced in \devito v4.0 to express
-an elastic model with compact and elegant notation.
+an elastic model with compact and elegant notation.}}
 

### 2020-06-02 paper.tex
Update paper.tex

@@
 applications we describe in this paper. A more thorough study of its implementation, including `mpi4numpy' and
-a complete scaling analysis will be presented in a later paper. Initial results assciated with early version of \devito's
+a complete scaling analysis will be presented in a later paper. Initial results associated with early version of \devito's
 distributed computing capabilities can be found in \cite{luporini2019adp} on a conventional cluster and 

### 2020-06-02 response.md
update response

@@
 
-> If you run your model with a variable number of cores, how is the weak and strong scaling?
-Can you comment on that?
-Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality?
-How is that done?
+> If you run your model ... How is that done?
 
@@
 We subsequently managed to work around point one above.
-Initial scaling experiments have been published in [..., page ..., figure ...], although this work still doesn't exploit the latest advancements in Devito v4.2 (e.g., computation/communication overlap).
-Our manuscript centers on the development of HPC production-grade wave propagators from a high-level language, and while performance is one of the cornerstones of our work, showing off Devito at extreme scale was not an objective.
-We have preliminary evidence that the latest Devito v4.2 has remarkable parallel efficiency, but we felt that publishing these results today would be premature -- we at least need testing and reproducibility on several architectures and applications.
+Initial scaling experiments have been published in [@witte2019ecl;@luporini2019adp], although this work still doesn't exploit the latest advancements in Devito v4.2 (e.g., computation/communication overlap).
+Our manuscript centers on the development of HPC production-grade wave propagators from a high-level language, and while performance is one of the cornerstones of our work, the application was the focus.
+We have preliminary evidence that the latest Devito v4.2 has remarkable parallel efficiency, but we felt that publishing these results today would be premature.
 
@@
 As partly explained in our reply to R1, we conceived this work as an application-oriented paper based on high-level abstractions and compilation technology.
-We're in the middle of an epochal change for computational science, with legacy codes being superseded, gradually, by DSLs; our work is just yet another small step in this direction.
+We are in the middle of an epochal change for computational science, with legacy codes being superseded, gradually, by DSLs; our work is just yet another small step in this direction.
 However, unlike many other works centered on DSLs, ours builds on -- and contributes to -- what is nowadays an established framework for seismic imaging.
@@
 This requires considerable software engineering (especially for an open source project such as Devito that was born in academia): we're talking about thousands of lines of carefully engineered code, plus tests and examples, to enable users to just type `mpirun -n X python ...` and get their code to run on a distributed-memory system with virtually no changes to their code.
-Secondly, as we explained in the paper [section ...], it's not just "classic domain decomposition": Devito orchestrates sparse functions, implements distributed numpy arrays, and performs several compiler-level optimazions for effective halo exchange.
+Secondly, as we explained in the paper [section ], it's not just "classic domain decomposition": Devito orchestrates sparse functions, implements distributed numpy arrays, and performs several compiler-level optimazions for effective halo exchange.
 Based on our experience, this is a fairly sophisticated system, quite far from being "straightforward".
@@
 The same story happens with our TTI.
-As we explain in the paper [section ...], without capturing the cross-iteration common sub-expressions (due to high order derivatives), we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
+As we explain in the paper [section III] without capturing the cross-iteration common sub-expressions (due to high order derivatives), we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
 And capturing such redundancies is indeed quite challenging.
@@
 
-As we elaborate in R2 (with references to the paper), it's not just about decomposing a grid.
-Devito deals with distribution of sparse functions, python-level distributed arrays, decomposition over sub-domains and much more.
+As we elaborate in R2, it's not just about decomposing a grid. Devito deals with distribution of sparse functions, python-level distributed arrays, decomposition over sub-domains and much more.
 
@@
 
-See reply to R1 -- basically, (i) lack of resources and (ii) Devito not ready yet for extreme scale at the time we ran the experiments.
+See reply to R1: (i) lack of resources and (ii) Devito not ready yet for scalability tests at the time we ran the experiments.
 
@@
 We are showing performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations.
-These results are, respectively, in sections [...].
-As we explain in our reply to R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
-To achieve this, we're using several bleeding edge technologies -- an established DSL-based system such as Devito (used in companies and academia), the Azure cloud, Docker (for ease of installation, use and reproducibility), the rapidly-spreading GitHub Actions for continuos integration, and much more.
+These results are, respectively, in sections II, III and I.
+As explained  to R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
+To achieve this, we are using several bleeding edge technologies -- an established DSL-based system such as Devito (used in companies and academia), the Azure cloud, Docker (for ease of installation, use and reproducibility), the rapidly-spreading GitHub Actions for continuos integration, and much more.
 The use and orchestration of these technologies is thouroughly described in the paper, and this should be, in our opinion, of strong interest for the SC audience.
@@
 
-We believe that showing the new capabilities of Devito -- the tensor language -- for a different physics and discretization is as important to the SC audience as a strong scaling experiment.
+We believe that showing the new capabilities of Devito -- the tensor language, distributed omputing via high-level abstractions -- for a different physics and discretization is as important to the SC audience as a strong scaling experiment.
 In fact, strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:
@@
 
-We have improved the manuscript to clarify what the contributions are. These are also now reiterated throughout the whole paper. TODO -- need to add pointers to what we added?
+We have improved the manuscript to clarify what the contributions are. These are also now reiterated throughout the whole paper (edits in red).
 

### 2020-06-03 response.md
update response

@@
 
-> If you run your model ... How is that done?
+> If you run your model [...] How is that done?
 
@@
 
-We subsequently managed to work around point one above.
-Initial scaling experiments have been published in [@witte2019ecl;@luporini2019adp], although this work still doesn't exploit the latest advancements in Devito v4.2 (e.g., computation/communication overlap).
-Our manuscript centers on the development of HPC production-grade wave propagators from a high-level language, and while performance is one of the cornerstones of our work, the application was the focus.
-We have preliminary evidence that the latest Devito v4.2 has remarkable parallel efficiency, but we felt that publishing these results today would be premature.
+We subsequently managed to work around point one.
+Initial scaling experiments have been published in [@witte2019ecl;@luporini2019adp], although it doesn't exploit the latest advancements in Devito(e.g., computation/communication overlap).
+We present here and application oriented paper for the development of HPC production-grade wave propagators from a high-level language.
+We have preliminary evidence that the latest Devito has remarkable parallel efficiency, but too premature to be included.
 
-About validation of simulation quality -- we did not include it in this paper as we think it is outside of the scope of the presented results, but the simulation quality is verified in two ways.
-As described in [@devito-api], the numerical accuracy is compared to the analytical solution for the acoustic wave equation as well as the convergence rate of the finite-difference discretization.
-The verification of TTI is not as simple as we do not have an analytical solution; we will be using the method of manufactured solution (MMS), but this is still work in progress.
-However, all of the key parts, such as finite-difference accuracy and comparison with isotropic results for zero-ed anisotropic parameters are implemented and part of our continuous integration system.
+Concerning the accuracy -- these were not included as out of the scope of this work, but the simulation quality is verified.
+In [@devito-api], we analylsed the numerical accuracy against the analytical solution and the convergence rate of the finite-difference discretization.
+The verification of TTI is not straightfoward (not analytical solution); we will use the method of manufactured solution (MMS).
+Our continuous integration also tests numerical accuracy in various ways including in the TTI case.
 
@@
 This requires considerable software engineering (especially for an open source project such as Devito that was born in academia): we're talking about thousands of lines of carefully engineered code, plus tests and examples, to enable users to just type `mpirun -n X python ...` and get their code to run on a distributed-memory system with virtually no changes to their code.
-Secondly, as we explained in the paper [section ], it's not just "classic domain decomposition": Devito orchestrates sparse functions, implements distributed numpy arrays, and performs several compiler-level optimazions for effective halo exchange.
+Secondly, as we explained in the paper [section I], it's not just "classic domain decomposition": Devito orchestrates sparse functions, implements distributed numpy arrays, and performs several compiler-level optimazions for effective halo exchange.
 Based on our experience, this is a fairly sophisticated system, quite far from being "straightforward".
@@
 
-> It was a little confusing but the MPI parallelism appears to essentially be a domain decomposition style parallelism.
+> It was a little confusing but [...] style parallelism.
 
@@
 
-> There was some mention of threads, in particular there seems to be several compilation passes that abstracts out the parallelism which can include SIMD, shared-memory and MPI.
-Only the MPI is detailed.
+> There was some mention of threads, [...]. Only the MPI is detailed.
 
@@
 
-> One question that arises, in the performance is how does Devito account for FMA (Fused Multiple Adds)?
+> [...] FMA (Fused Multiple Adds)?
 
@@
 
-> It appears much of the performance result are provided in the references so it is unclear as to what the purpose of this paper is for an SC audience.
+> It appears much of the performance [...]
 
 We are showing performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations.
-These results are, respectively, in sections II, III and I.
 As explained to R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
-To achieve this, we are using several bleeding edge technologies -- an established DSL-based system such as Devito (used in companies and academia), the Azure cloud, Docker (for ease of installation, use and reproducibility), the rapidly-spreading GitHub Actions for continuos integration, and much more.
-The use and orchestration of these technologies is thouroughly described in the paper, and this should be, in our opinion, of strong interest for the SC audience.
+To achieve this, we are using bleeding edge technologies -- an established DSL-based system such as Devito (used in companies and academia), Azure cloud, Docker (portability), the rapidly-spreading GitHub-Actions for CI, ....
+The use and orchestration of these technologies is thouroughly described in the paper, and is, in our opinion, of strong interest for the SC audience.
 
-> For actual seismic surveys, the imaging (interpretation) part is only 10% of the survey. [...] while important doesn't address the real "industrial" problems, it helps a little.
+> For actual seismic [...] it helps a little.
 
@@
 We do not claim here to solve all these problems; however, we demonstrate that the combination of advanced technolgies such as high-level DSLs and compilers is key to highly efficient sesimic modeling and imaging at industrial scale.
-The complexity of model building requires advanced optimization methods.
-The straightforward access to computationally and easy-to-develop wave-equation solvers enable research and development for seismic inversion.
-Devito is therfore used for the development of inital background model building [@mojica2019tab] or cycle skipping mitigating methods [@rizzuti2020EAGEtwri; @louboutin2020SEGtwri] and compressive least square migration [@witte2018cls] based on a linear algebra framework for seismic inversion built on top of devito [@witte2018alf].
-We have added a clarification in the introduction.
+Devito is used both in academia/industry for various applications.
+We have added the references and clarification in the introduction.
 
@@
 
-> Is section III describing new work? It is unclear to me if this is an overview or new work for this paper…
+> Is section III describing new work?[...]
 
@@
 
-> Is the parallelism just via MPI?
+> Is the parallelism just via MPI?[...]
 
@@
 
-> What are the possibilities for this code in terms of heterogeneous clusters? Will it be able to take advantage of GPUs?
+> What are the possibilities [...]
 

### 2020-06-04 response.md
Update response.md

@@
 
-We subsequently managed to work around point one.
-Initial scaling experiments have been published in [@witte2019ecl;@luporini2019adp], although it doesn't exploit the latest advancements in Devito(e.g., computation/communication overlap).
-We present here and application oriented paper for the development of HPC production-grade wave propagators from a high-level language.
-We have preliminary evidence that the latest Devito has remarkable parallel efficiency, but too premature to be included.
+We subsequently managed to work around 1.: initial scaling experiments have been published in [@witte2019ecl;@luporini2019adp], although with the latest advancements in Devito (e.g., computation/communication overlap) still unexploited.
+Today, we have preliminary evidence that Devito 4.2 has remarkable parallel efficiency, but we feel publication would be premature.
 
-Concerning the accuracy -- these were not included as out of the scope of this work, but the simulation quality is verified.
-In [@devito-api], we analylsed the numerical accuracy against the analytical solution and the convergence rate of the finite-difference discretization.
-The verification of TTI is not straightfoward (not analytical solution); we will use the method of manufactured solution (MMS).
-Our continuous integration also tests numerical accuracy in various ways including in the TTI case.
+Concerning the accuracy -- it was out of the scope of this work, but the simulation quality is verified.
+In [@devito-api], we have analyzed both the numerical accuracy against the analytical solution and the convergence rate of the finite-difference discretization.
+In Devito, we systematically test the numerical accuracy of all of the implemented models; this is automated through continuous integration.
 
@@
 
-> However, it is not very suitable for SC
+> [...] not very suitable for SC
 
-As partly explained in our reply to R1, we conceived this work as an application-oriented paper based on high-level abstractions and compilation technology.
-We are in the middle of an epochal change for computational science, with legacy codes being superseded, gradually, by DSLs; our work is just yet another small step in this direction.
-However, unlike many other works centered on DSLs, ours builds on -- and contributes to -- what is nowadays an established framework for seismic imaging.
-We believe that this should be of strong interest for the SC community, especially since -- we believe -- the methodology and philosophy underpinning Devito is reusable in computational science.
+This work was conceived as an application-oriented paper based on high-level abstractions and compilation technology.
+However, unlike many other works centered on DSLs, ours builds on -- and contributes to -- what is nowadays an established framework for seismic imaging, used in industry and academia.
+Our work shows the benefits (productivity, HPC, ...) of raising the level of abstraction, so it should be relevant for SC.
 
@@
 
-We remark that our work does _not_ describe some ad-hoc MPI-based domain decomposition; what Devito can do is _generation_ of MPI parallel code from a mathematical language for solving partial differential equations.
-This requires considerable software engineering (especially for an open source project such as Devito that was born in academia): we're talking about thousands of lines of carefully engineered code, plus tests and examples, to enable users to just type `mpirun -n X python ...` and get their code to run on a distributed-memory system with virtually no changes to their code.
-Secondly, as we explained in the paper [section I], it's not just "classic domain decomposition": Devito orchestrates sparse functions, implements distributed numpy arrays, and performs several compiler-level optimazions for effective halo exchange.
-Based on our experience, this is a fairly sophisticated system, quite far from being "straightforward".
+Our work does _not_ describe some ad-hoc MPI-based domain decomposition, but rather the _generation_ MPI parallel code from a mathematical language for solving partial differential equations.
+This requires considerable software engineering -- thousands of lines of carefully engineered code, plus tests and examples, which allow users to run on a distributed-memory system with virtually no changes to their code.
+Secondly, as explained in the paper [sections I and II-B], it's not just "classic domain decomposition": Devito also decomposes sparse functions, implements distributed numpy arrays, and performs several compiler-level optimazions for effective halo exchange.
+This results in a fairly sophisticated system.
 
-> The computational kernels have really high arithmetic intensity that is great for the domain, but not as challenging for optimization
+> kernels have really high arithmetic intensity [...] not as challenging for optimization
 
@@
 
-> The parallelism applied so far is straight up MPI which is a bit behind times these days.
+> [...] MPI which is a bit behind times these days.
 

### 2020-06-04 response.md
Update response.md

@@
 
-High arithmetic intensity does not imply "non challenging optimization".
-In fact, it's the exact opposite, and there are several examples in computational science that we could use as evidence, even outside of seismic.
-Take high-order finite (spectral) elements for instance -- you may get an extremely good arithmetic intensity, but without tensor-product basis functions and suitable optimizations such as sum-factorization you'll likely end up with a notable, yet totally misleasing, GFlops/s performance, as a huge fraction of flops will stem from redundant computation (e.g., common sub-expressions, lack of code hoisting, distributivity).
+High arithmetic intensity does not imply "non-challenging optimization".
+In fact, it's the opposite, even outside of seismic.
+Take high-order finite (spectral) elements for instance -- you may get an extremely high arithmetic intensity, but without tensor-product basis functions and optimizations such as sum-factorization, ultimately you'll get a notable yet misleasing GFlops/s performance, as a huge fraction of flops will stem from redundant computation (e.g., common sub-expressions, lack of code hoisting, distributivity).
 The same story happens with our TTI.
-As we explain in the paper [section III] without capturing the cross-iteration common sub-expressions (due to high order derivatives), we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
-And capturing such redundancies is indeed quite challenging.
-In fact, this is one of the reasons several companies are today so much interested in Devito: machines (compilers) can be better than humans at detecting redundancies, and definitely way faster than them.
+As we explain in the paper [section III], without capturing the cross-iteration common sub-expressions (due to high order derivatives), we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
+Detecting and scheduling such redundancies is cumbersome.
+This is also one of the reasons there is today so much interest in Devito by companies: machines (compilers) can be better than humans at optimizing away redundancies, and definitely way faster than them.
 
@@
 
-In 2020 MPI is still de facto a widely adopted approach for domain decomposition.
+MPI is still a widely adopted approach for domain decomposition.
 
@@
 
-> There was some mention of threads, [...]. Only the MPI is detailed.
+> some mention of threads, [...]. Only the MPI is detailed.
 
@@
 
-> Scalability performance would also be interesting.
+> Scalability performance
 
@@
 
-> Dropping the discussion on the 3D elastic wave and [...] might demonstrate the value of Devito better.
+> Dropping the discussion on the 3D elastic [...] might demonstrate the value of Devito better.
 
@@
 
-The number of additions to the codebase is notable: https://github.com/devitocodes/devito/pull/873
-And for Devito and his users, this was an invaluable addition.
+The number of additions to the codebase is significant: https://github.com/devitocodes/devito/pull/873
 
-> but I think for SC we would need more information on its scalability and performance.
+> for SC we would need [...] scalability and performance.
 
@@
 
-> Is section III describing new work?[...]
+> Is section III describing new work?
 
-Yes. We have improved the text to make it more explicit and detailed.
+Yes. We have improved the text to make this clear.
 
-> Is the parallelism just via MPI?[...]
+> Is the parallelism just via MPI?
 
 The inversion (outer-level of parallelism) requires what is essentially task parallelism; each task is internally MPI-parallel (domain decomposition).
-In this context task-parallelism means paralleization over sources, which we have accomplished using `batch-shipyard` in a serverless setting (see Section II end).
-We remark that this allows optimal usage of the typical cloud infrastructure, where one pays based on actual usage of computational resources.
+In this context task-parallelism means parallelization over sources, which we have accomplished using `batch-shipyard` in a serverless setting (see Section II).
+This allowed optimal usage of the cloud, where one pays based on actual usage of computational resources.
 
@@
 
-GPU support is in development and currently (Devito v4.2) supported for some applications. Its capabilities and related applications will be fully covered in a future dedicated paper.
+GPU support is in development and currently (Devito v4.2) supported for some applications. Its capabilities and related applications will be fully covered in a future paper.
 

### 2020-06-04 response.md
Update response.md

@@
 
-> If you run your model [...] How is that done?
+> you run your model [...] How is that done?
 
-At the time of the experimentation, we ran into two main issues:
+During experimentation, we had two issues:
 1. Limited resources on Azure, including the access to an Infiniband-enabled system.
@@
 
-We subsequently managed to work around 1.: initial scaling experiments have been published in [@witte2019ecl;@luporini2019adp], although with the latest advancements in Devito (e.g., computation/communication overlap) still unexploited.
+We later managed to perform initial scaling experiments [@witte2019ecl;@luporini2019adp], although still without the latest advancements in Devito (e.g., computation/communication overlap).
 Today, we have preliminary evidence that Devito 4.2 has remarkable parallel efficiency, but we feel publication would be premature.
@@
 In [@devito-api], we have analyzed both the numerical accuracy against the analytical solution and the convergence rate of the finite-difference discretization.
-In Devito, we systematically test the numerical accuracy of all of the implemented models; this is automated through continuous integration.
+In Devito, we systematically test, through continuous integration, the numerical accuracy of all of the implemented models.
 
@@
 
-> Parallelization is pretty straightforward, uniform decomposition of relatively straightforward operation.
+> Parallelization is pretty straightforward, [...] straightforward operation.
 
-Our work does _not_ describe some ad-hoc MPI-based domain decomposition, but rather the _generation_ MPI parallel code from a mathematical language for solving partial differential equations.
-This requires considerable software engineering -- thousands of lines of carefully engineered code, plus tests and examples, which allow users to run on a distributed-memory system with virtually no changes to their code.
-Secondly, as explained in the paper [sections I and II-B], it's not just "classic domain decomposition": Devito also decomposes sparse functions, implements distributed numpy arrays, and performs several compiler-level optimazions for effective halo exchange.
+Our work does _not_ describe some ad-hoc MPI-based domain decomposition, but rather the _generation_ of MPI-parallel code from a mathematical language for solving PDEs.
+This requires considerable software engineering -- thousands of lines of carefully engineered code, which allow users to run on a distributed-memory system with virtually no changes to their code.
+Secondly, as explained in the paper [sections I and II-B], it's not just "classic domain decomposition": Devito also decomposes sparse functions, implements distributed numpy arrays, and applies several optimazions for effective halo exchange.
 This results in a fairly sophisticated system.
 
-> kernels have really high arithmetic intensity [...] not as challenging for optimization
+> really high arithmetic intensity [...] not as challenging for optimization
 
@@
 In fact, it's the opposite, even outside of seismic.
-Take high-order finite (spectral) elements for instance -- you may get an extremely high arithmetic intensity, but without tensor-product basis functions and optimizations such as sum-factorization, ultimately you'll get a notable yet misleasing GFlops/s performance, as a huge fraction of flops will stem from redundant computation (e.g., common sub-expressions, lack of code hoisting, distributivity).
+Take high-order finite (spectral) elements for instance -- you may get an extremely high arithmetic intensity, but without tensor-product basis functions and optimizations such as sum-factorization, ultimately you'll get a notable yet misleasing GFlops/s performance, since a huge fraction of flops are redundant (e.g., common sub-expressions, hoisting, contractions).
 The same story happens with our TTI.
-As we explain in the paper [section III], without capturing the cross-iteration common sub-expressions (due to high order derivatives), we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
-Detecting and scheduling such redundancies is cumbersome.
-This is also one of the reasons there is today so much interest in Devito by companies: machines (compilers) can be better than humans at optimizing away redundancies, and definitely way faster than them.
+As we explain in the paper [section III], without capturing the cross-iteration common sub-expressions due to high-order derivatives, we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
+Optimizing away such redundancies is cumbersome, and this is one of the reasons there is today so much interest in Devito by companies.
 
@@
 
-As we elaborate in R2, it's not just about decomposing a grid. Devito deals with distribution of sparse functions, python-level distributed arrays, decomposition over sub-domains and much more.
+As we elaborate in R2, it's not just "grid decomposition". Devito deals with distribution of sparse functions, distributed numpy arrays, decomposition over sub-domains and much more.
 
@@
 
-The operations determining the performance of a code are the source level ones, while FMA is a lower level concept (ie, a scalar FMA would account for two operations).
-This is not Devito, but rather the model adopted everywhere, otherwise two performance values from two different architectures, one with FMAs the other without, would not even be comparable.
+The operations determining the performance of a code are the source-level ones.
+FMA is a lower-level concept (ie, a scalar FMA would account for two operations).
+This is not Devito, but rather the model adopted everywhere.
 
@@
 
-See reply to R1: (i) lack of resources and (ii) Devito not ready yet for scalability tests at the time we ran the experiments.
+See reply to R1: (i) lack of resources and (ii) Devito not ready yet for scalability tests when we performed the experimentation.
 
@@
 We are showing performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations.
-As explained to R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
-To achieve this, we are using bleeding edge technologies -- an established DSL-based system such as Devito (used in companies and academia), Azure cloud, Docker (portability), the rapidly-spreading GitHub-Actions for CI, ....
-The use and orchestration of these technologies is thouroughly described in the paper, and is, in our opinion, of strong interest for the SC audience.
+Further, as explained to R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
+To achieve this, we are using bleeding edge technologies -- an established DSL-based system such as Devito, Azure cloud, Docker, the rapidly-spreading GitHub-Actions for CI, etc.
+The use and orchestration of these technologies is thouroughly described in the paper, and is, in our opinion, of strong interest for SC.
 
@@
 The industrial problem does indeed involve muliple parts, such as data acquisition, processing, and, most importantly, velocity model building.
-We do not claim here to solve all these problems; however, we demonstrate that the combination of advanced technolgies such as high-level DSLs and compilers is key to highly efficient sesimic modeling and imaging at industrial scale.
-Devito is used both in academia/industry for various applications.
+We do not claim here to solve all these problems; however, we demonstrate that the combination of advanced technolgies such as DSLs and compilers is key to highly-efficient sesimic modeling and imaging at industrial scale.
 We have added the references and clarification in the introduction.
 
-> Dropping the discussion on the 3D elastic [...] might demonstrate the value of Devito better.
+> Dropping [...] 3D elastic [...] might demonstrate the value of Devito better.
 
-We believe that showing the new capabilities of Devito -- the tensor language, distributed omputing via high-level abstractions -- for a different physics and discretization is as important to the SC audience as a strong scaling experiment.
+We believe that showing the new capabilities of Devito -- the tensor language, distributed computing from high-level abstractions -- for a different physics and discretization is as important to the SC audience as a strong scaling experiment.
 In fact, strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:

### 2020-06-04 response.md
Update response.md

@@
 
-First of all, we would like to thank the reviewers for their work and constructive feedback. Below, we address all of their questions and observations.
+First of all, we would like to thank the reviewers for their work and constructive feedback. Below, we address their questions and observations.
 
@@
 During experimentation, we had two issues:
-1. Limited resources on Azure, including the access to an Infiniband-enabled system.
-2. Most of the MPI-level optimizations that are today in Devito still needed tuning and consolidation.
+1-Limited Azure resources, including access to an Infiniband-enabled system.
+2-Most of the MPI-level optimizations today in Devito still needed tuning and consolidation.
 
 We later managed to perform initial scaling experiments [@witte2019ecl;@luporini2019adp], although still without the latest advancements in Devito (e.g., computation/communication overlap).
-Today, we have preliminary evidence that Devito 4.2 has remarkable parallel efficiency, but we feel publication would be premature.
+Today, we have preliminary evidence that Devito-v4.2 has remarkable parallel efficiency, but publication would be premature.
 
-Concerning the accuracy -- it was out of the scope of this work, but the simulation quality is verified.
+Concerning the accuracy -- the simulation quality was verified.
 In [@devito-api], we have analyzed both the numerical accuracy against the analytical solution and the convergence rate of the finite-difference discretization.
-In Devito, we systematically test, through continuous integration, the numerical accuracy of all of the implemented models.
+We have tested, and we keep doing so through continuous integration, the numerical accuracy of all implemented models.
 
@@
 However, unlike many other works centered on DSLs, ours builds on -- and contributes to -- what is nowadays an established framework for seismic imaging, used in industry and academia.
-Our work shows the benefits (productivity, HPC, ...) of raising the level of abstraction, so it should be relevant for SC.
+Our work shows the benefits (productivity, HPC, ...) of raising the level of abstraction, so it is relevant for SC.
 
-> Parallelization is pretty straightforward, [...] straightforward operation.
+> Parallelization is pretty straightforward [...]
 
@@
 This requires considerable software engineering -- thousands of lines of carefully engineered code, which allow users to run on a distributed-memory system with virtually no changes to their code.
-Secondly, as explained in the paper [sections I and II-B], it's not just "classic domain decomposition": Devito also decomposes sparse functions, implements distributed numpy arrays, and applies several optimazions for effective halo exchange.
+Secondly, as explained in the paper [Sections I and II-B], it's not just "classic domain decomposition": Devito also decomposes sparse functions, implements distributed numpy arrays, and applies several optimazions for effective halo exchange.
 This results in a fairly sophisticated system.
@@
 The same story happens with our TTI.
-As we explain in the paper [section III], without capturing the cross-iteration common sub-expressions due to high-order derivatives, we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
-Optimizing away such redundancies is cumbersome, and this is one of the reasons there is today so much interest in Devito by companies.
+As we explain in the paper [Section III], without capturing the cross-iteration common sub-expressions due to high-order derivatives, we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
+Optimizing away such redundancies is cumbersome, and this is one of the reasons companies are so much interested in Devito.
 
@@
 
-MPI is still a widely adopted approach for domain decomposition.
+MPI is de facto the standard approach for domain decomposition.
 
@@
 
-As we elaborate in R2, it's not just "grid decomposition". Devito deals with distribution of sparse functions, distributed numpy arrays, decomposition over sub-domains and much more.
+See R2 -- it's not just "grid decomposition", Devito deals with distribution of sparse functions, distributed numpy arrays, decomposition over sub-domains, etc.
 
-> some mention of threads, [...]. Only the MPI is detailed.
+> some mention of threads, [...]. Only MPI is detailed.
 
-Correct. We refer to other papers for all these other aspects.
+Correct. We explicitly refer to other articles for these other aspects.
 
@@
 
-The operations determining the performance of a code are the source-level ones.
-FMA is a lower-level concept (ie, a scalar FMA would account for two operations).
-This is not Devito, but rather the model adopted everywhere.
+The operations determining the performance of a code are the source-level ones. FMA is a lower-level concept (ie, a scalar FMA would account for two operations). This model is adopted everywhere.
 
@@
 
-See reply to R1: (i) lack of resources and (ii) Devito not ready yet for scalability tests when we performed the experimentation.
+See R1 -- (i) lack of resources and (ii) Devito not ready yet for scalability tests when the experimentation was performed.
 
@@
 
-We are showing performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations.
-Further, as explained to R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
+We show performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations.
+Further, as explained in R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
 To achieve this, we are using bleeding edge technologies -- an established DSL-based system such as Devito, Azure cloud, Docker, the rapidly-spreading GitHub-Actions for CI, etc.
@@
 The industrial problem does indeed involve muliple parts, such as data acquisition, processing, and, most importantly, velocity model building.
-We do not claim here to solve all these problems; however, we demonstrate that the combination of advanced technolgies such as DSLs and compilers is key to highly-efficient sesimic modeling and imaging at industrial scale.
-We have added the references and clarification in the introduction.
+We do not claim to solve all these problems.
+However, we demonstrate that the combination of advanced technolgies such as DSLs and compilers is key to highly-efficient sesimic modeling and imaging at industrial scale.
+We have added references and clarification in the introduction.
 
@@
 
-We believe that showing the new capabilities of Devito -- the tensor language, distributed computing from high-level abstractions -- for a different physics and discretization is as important to the SC audience as a strong scaling experiment.
+We believe showing the new capabilities of Devito -- the tensor language, distributed computing from high-level abstractions -- for multiple physics and discretizations is as important as a strong scaling experiment.
 In fact, strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:
@@
 * even with high-frequency FWI, a typical shot-level computation requires no more than a bunch of nodes (often order of units);
-* there is an outer-level of parallelism for inversion, essentially a task farm where the individual workers is internally MPI-parallel.
+* there is an outer-level of parallelism for inversion, essentially a task farm where the individual workers are internally MPI-parallel.
 
@@
 
-> One issue is that I cannot tell how much of this work is new
+> One issue [...] how much of this work is new
 
@@
 
-> [...] vector and tensors [...] I am unclear of the amount of effort or innovation involved.
+> [...] vector and tensors [...] I am unclear [...] effort [...] involved.
 
@@
 
-> for SC we would need [...] scalability and performance.
+> for SC [...] sacalability and performance.
 
-The single-node performance is discussed in other (cited) articles. In the previous responses (R1, R2, R3), we have elaborated why the lack of scalability experiments.
+The single-node performance is discussed in other cited articles. See R1 about the lack of scalability experiments.
 
@@
 The inversion (outer-level of parallelism) requires what is essentially task parallelism; each task is internally MPI-parallel (domain decomposition).
-In this context task-parallelism means parallelization over sources, which we have accomplished using `batch-shipyard` in a serverless setting (see Section II).
+In this context task-parallelism means parallelization over sources, accomplished using `batch-shipyard` in a serverless setting [Section II].
 This allowed optimal usage of the cloud, where one pays based on actual usage of computational resources.
@@
 
-GPU support is in development and currently (Devito v4.2) supported for some applications. Its capabilities and related applications will be fully covered in a future paper.
+GPU support is in development and currently (Devito-v4.2) supported for some applications. Its capabilities and related applications will be fully covered in a future paper.
 

### 2020-06-04 paper.tex
update response
Minor typos

@@
 time-domain Wavefield Reconstruction Inversion (WRI)~\cite{rizzuti2020EAGEtwri, 
-louboutin2020SEGtwri}, the later of which takles the cycle-skipping issue
+louboutin2020SEGtwri}, the later of which tackles the cycle-skipping issue
 present in seismic inversion.}
@@
 \textcolor{red}{where $u(t, x)$ is the pressure wavefield, $xs$ is the point source position, 
-$q(t)$ is thes ource time signature, $d(t, .)$ is the measured data at positions $xr$ and $m(x)$ is the 
+$q(t)$ is the source time signature, $d(t, .)$ is the measured data at positions $xr$ and $m(x)$ is the
 squared slowness.}
@@
 for thorough explanations and practical examples.
-\textcolor{red}{We highlight the main features of the distributed framework tha enable the industry scale
-applications we describe in this paper. A more thorough study of its implementation, including `mpi4numpy' and
+\textcolor{red}{We highlight the main features of the distributed framework that enable the industry scale
+applications we describe in this paper. A more thorough study of its implementation, including MPI over numpy and
 a complete scaling analysis will be presented in a later paper. Initial results associated with early version of \devito's

### 2020-06-04 response.md
more text

@@
 We do not claim to solve all these problems.
-However, we demonstrate that the combination of advanced technolgies such as DSLs and compilers is key to highly-efficient sesimic modeling and imaging at industrial scale.
+However, we demonstrate that the combination of advanced technolgies such as DSLs and compilers is key to highly-efficient sesimic modelling and imaging at industrial scale.
 We have added references and clarification in the introduction.

### 2020-06-04 paper.tex
typo
more text

@@
 \textcolor{red}{We highlight the main features of the distributed framework that enable the industry scale
-applications we describe in this paper. A more thorough study of its implementation, including MPI over numpy and
+applications we describe in this paper. A more thorough study of its implementation, including MPI over NumPy and
 a complete scaling analysis will be presented in a later paper. Initial results associated with early version of \devito's
@@
 equations but also for coupled PDEs. This interface allows, in a simple, concise
-and consistent manner, the expression of all-kinds of non-trivial
-differential operators. Next, and most
+and consistent manner, the expression of all-kinds of non-trivial differential operators \textcolor{red}{, in this the vectorial elastic equation 
+and rotated finite-difference}. Next, and most
 importantly, \textcolor{red}{it was} demonstrated that the compiler enables large-scale modelling with
@@
 we demonstrated that our abstractions provide the necessary portability
-to enable both on-premise and Cloud based HPC.
+to enable both on-premise and Cloud based HPC. \textcolor{red}{WHile the work we 
+presented here does not solve all the problems in exploration geophysics, the high-level interface 
+and at (industry) scale performance we demonstrated convinced
+multiple academic and industrial project to use \devito.}
 

### 2020-06-04 response.md
add header for easy formating of refs
1000 words

@@
+---
+title: respons to reviews
+bibliography:
+	- sc20_paper.bib
+---
 Dear Editor,
 
-First of all, we would like to thank the reviewers for their work and constructive feedback. Below, we address their questions and observations.
+First of all, we would like to thank the reviewers for their work and constructive feedback. Below, we address their questions and observations. All edits are in red in the manuscript.
 
@@
 
-> you run your model [...] How is that done?
+> run your model [...] How is that done?
 
 During experimentation, we had two issues:
-1-Limited Azure resources, including access to an Infiniband-enabled system.
-2-Most of the MPI-level optimizations today in Devito still needed tuning and consolidation.
+1 - Limited Azure resources, including access to an Infiniband-enabled system.
+2 - Some MPI-level optimizations needed tuning and consolidation.
 
-We later managed to perform initial scaling experiments [@witte2019ecl;@luporini2019adp], although still without the latest advancements in Devito (e.g., computation/communication overlap).
+We managed to perform initial scaling experiments [@witte2019ecl;@luporini2019adp], although without the latest advancements in Devito (e.g., computation/communication overlap).
 Today, we have preliminary evidence that Devito-v4.2 has remarkable parallel efficiency, but publication would be premature.
 
-Concerning the accuracy -- the simulation quality was verified.
-In [@devito-api], we have analyzed both the numerical accuracy against the analytical solution and the convergence rate of the finite-difference discretization.
-We have tested, and we keep doing so through continuous integration, the numerical accuracy of all implemented models.
+Accuracy: details added to the text (in red).
+In [@devito-api], we analyze the numerical accuracy against the analytical solution and the convergence rate of the finite-differences.
+We also test through continuous integration the numerical accuracy of all implemented models.
 
@@
 
-Our work does _not_ describe some ad-hoc MPI-based domain decomposition, but rather the _generation_ of MPI-parallel code from a mathematical language for solving PDEs.
-This requires considerable software engineering -- thousands of lines of carefully engineered code, which allow users to run on a distributed-memory system with virtually no changes to their code.
-Secondly, as explained in the paper [Sections I and II-B], it's not just "classic domain decomposition": Devito also decomposes sparse functions, implements distributed numpy arrays, and applies several optimazions for effective halo exchange.
-This results in a fairly sophisticated system.
+Our work focuse on the generation of MPI-parallel code from a mathematical language for solving PDEs rather than the manual addition of domain decomposition toanexi  sting code.
+This requires considerable software engineering to allow users to run on a distributed-memory system with virtually no changes to their code.
+We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes unstructured sparse functions, implements distributed numpy arrays, and applies several optimazions for effective halo exchange.
 
-> really high arithmetic intensity [...] not as challenging for optimization
+> high arithmetic intensity [...] challenging for optimization
 
 High arithmetic intensity does not imply "non-challenging optimization".
-In fact, it's the opposite, even outside of seismic.
-Take high-order finite (spectral) elements for instance -- you may get an extremely high arithmetic intensity, but without tensor-product basis functions and optimizations such as sum-factorization, ultimately you'll get a notable yet misleasing GFlops/s performance, since a huge fraction of flops are redundant (e.g., common sub-expressions, hoisting, contractions).
-The same story happens with our TTI.
-As we explain in the paper [Section III], without capturing the cross-iteration common sub-expressions due to high-order derivatives, we would have obtained remarkable GFlops/s performance, but terrible GPoints/s.
-Optimizing away such redundancies is cumbersome, and this is one of the reasons companies are so much interested in Devito.
+High-order finite (spectral) elements for instance -- where an extremely high arithmetic intensity can be obtained, but without tensor-product basis functions and optimizations such as sum-factorization, notable yet misleasing GFlops/s performance is achieved since a huge fraction of flops are redundant (e.g., common sub-expressions).
+TTI is similar.
+As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtaine remarkable GFlops/s, but terrible GPoints/s.
+Optimizing away such redundancies is cumbersome, and is one of the reasons companies are so much interested in Devito.
 
@@
 
-MPI is de facto the standard approach for domain decomposition.
+MPI is still curently the most broadly accepted approach for domain decomposition.
 
@@
 
-> It was a little confusing but [...] style parallelism.
+> little confusing but [...] style parallelism.
 
-See R2 -- it's not just "grid decomposition", Devito deals with distribution of sparse functions, distributed numpy arrays, decomposition over sub-domains, etc.
+See R2 -- it's not just "grid decomposition", Devito deals with distribution of sparse functions, numpy arrays, decomposition over sub-domains, etc.
 
@@
 
-> [...] FMA (Fused Multiple Adds)?
+> [...] FMA?
 
-The operations determining the performance of a code are the source-level ones. FMA is a lower-level concept (ie, a scalar FMA would account for two operations). This model is adopted everywhere.
+The operations determining the performance of a code are the source-level ones. FMA is a lower-level concept (ie, a scalar FMA = two operations). This model is adopted everywhere.
 
@@
 
-> It appears much of the performance [...]
+> much of the performance [...]
 
@@
 To achieve this, we are using bleeding edge technologies -- an established DSL-based system such as Devito, Azure cloud, Docker, the rapidly-spreading GitHub-Actions for CI, etc.
-The use and orchestration of these technologies is thouroughly described in the paper, and is, in our opinion, of strong interest for SC.
+The use and orchestration of these technologies is thouroughly described in the paper, and is of strong interest for SC (strong industry interest for production scale).
 
@@
 
-> Dropping [...] 3D elastic [...] might demonstrate the value of Devito better.
+> Dropping [...] 3D elastic [...] might demonstrate [...].
 
-We believe showing the new capabilities of Devito -- the tensor language, distributed computing from high-level abstractions -- for multiple physics and discretizations is as important as a strong scaling experiment.
-In fact, strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:
+The new capabilities of Devito -- tensor language, MPI from high-level abstractions -- for multiple physics and discretizations is as important as a strong scaling experiment.
+Strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:
 
@@
 
-> One issue [...] how much of this work is new
+> [...] how much of this work is new
 
@@
 
-> [...] vector and tensors [...] I am unclear [...] effort [...] involved.
+> [...] vector and tensors [...] unclear [...] involved.
 
@@
 
-> Is section III describing new work?
+>  section III describing new work?
 
@@
 
-> Is the parallelism just via MPI?
+> parallelism just via MPI?
 

### 2020-06-05 response.md

@@
 ---
-title: respons to reviews
+title: response to reviews
 bibliography:
@@
 
-First of all, we would like to thank the reviewers for their work and constructive feedback. Below, we address their questions and observations. All edits are in red in the manuscript.
+We thank the reviewers for their work and constructive feedback. Below, we address their questions and observations. All edits are in red in the manuscript.
 
 ## R1
+> If you run your model with a variable number of cores, how is the weak and strong scaling? Can you comment on that? 
+We published and cited MPI scaling results in [@witte2019ecl;@luporini2019adp].
+While Devito MPI scaling has improved significantly since those publications, it is only a relatively minor consideration within the bigger picture.
+Applications such as FWI generally only use MPI for computing on a single data shot (i.e.
+a single observation).
+However, that computation unit of work sits within an embarrassingly parallel loop processing thousands of such shots as part of a gradient-decent algorithm.
+So while the computation of a single shot (the focus on Devito_ might only require TFLOPS, and iteration of the outer loop would be typically PFLOPS.
+A description of nested-parallelism in applications such as FWI is outside of the scope of this paper, but we have included several references to FWI papers that do describe the underlying algorithm.
 
-> run your model [...] How is that done?
-
-During experimentation, we had two issues:
-1 - Limited Azure resources, including access to an Infiniband-enabled system.
-2 - Some MPI-level optimizations needed tuning and consolidation.
-
-We managed to perform initial scaling experiments [@witte2019ecl;@luporini2019adp], although without the latest advancements in Devito (e.g., computation/communication overlap).
-Today, we have preliminary evidence that Devito-v4.2 has remarkable parallel efficiency, but publication would be premature.
-
-Accuracy: details added to the text (in red).
+> Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality? How is that done?
+Details on solver verification have been added to the paper (in red).
 In [@devito-api], we analyze the numerical accuracy against the analytical solution and the convergence rate of the finite-differences.
 We also test through continuous integration the numerical accuracy of all implemented models.
+Additionally, for inverse problems, we use the standard dot and gradient tests, "we focused on correct and testable implementations for the adjoint wave-equation and the gradient (action of the adjoint Jacobian) [@louboutin2018segeow; @louboutin2020THmfi]"
 
@@
 > [...] not very suitable for SC
-
-This work was conceived as an application-oriented paper based on high-level abstractions and compilation technology.
-However, unlike many other works centered on DSLs, ours builds on -- and contributes to -- what is nowadays an established framework for seismic imaging, used in industry and academia.
-Our work shows the benefits (productivity, HPC, ...) of raising the level of abstraction, so it is relevant for SC.
+There are few words in this review so we can only guess in what sense the reviewer feels the work is not suitable for SC.
+High productivity for HPC and performance-portability continue to be hot topics at SC.
+Many recent Gordon Bell finalists in recent years have used to some degree DSL's, and DSL's.
+Devito is one of the very examples of DSL's to have challenged the performance of handwritten HPC codes for real engineering applications at scale.
+While Devito is fully open source, we have not been able to publish any head-to-head comparisons with commercial codes in this domain because of the commercial sensitivity.
+However, one can conclude from the fact that Devito is being adopted by companies commercially to suggest comparisons were favourable.
+The research in Devito is highly relevant to SC as it demonstrates that the HPC cliche that one have to can have high-productivity languages or high-performance but not both is false.
 
@@
 
-Our work focuse on the generation of MPI-parallel code from a mathematical language for solving PDEs rather than the manual addition of domain decomposition toanexi  sting code.
-This requires considerable software engineering to allow users to run on a distributed-memory system with virtually no changes to their code.
-We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes unstructured sparse functions, implements distributed numpy arrays, and applies several optimazions for effective halo exchange.
+Whether or not parallelization is straightforward is not the issue. One can make the same comment about any finite difference, stencil code, anything involving linear algebra... 
+The novelty is that our work involves developing a layers of abstracts that allow finite difference based solvers be to implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism.
+This is not "straightforward" - the only comparable software technologies that we are aware of to have accomplished something similar are FEnICS and Firedrake for the finite-element method.
+We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed numpy arrays, and applies several optimizations for efficient halo exchange.
 
@@
 High arithmetic intensity does not imply "non-challenging optimization".
-High-order finite (spectral) elements for instance -- where an extremely high arithmetic intensity can be obtained, but without tensor-product basis functions and optimizations such as sum-factorization, notable yet misleasing GFlops/s performance is achieved since a huge fraction of flops are redundant (e.g., common sub-expressions).
-TTI is similar.
-As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtaine remarkable GFlops/s, but terrible GPoints/s.
-Optimizing away such redundancies is cumbersome, and is one of the reasons companies are so much interested in Devito.
+High-order finite (spectral) elements for instance -- where an extremely high arithmetic intensity can be obtained, but without tensor-product basis functions and optimizations such as sum-factorization results in an inflated GFlops/s measurement since a huge fraction of flops are redundant (e.g., common sub-expressions).
+TTI, a relatively sophisticated propagation model used in industry, is similar.
+As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtain very high GFlops/s, but terrible GPoints/s (and therefore poor time to solution).
+Optimizing away such redundancies, one of the compiler transformations performed by Devito,  is challenging to do by hand as there can be thousands of algebraic terms in these engineering applications.
 
@@
 
-MPI is still curently the most broadly accepted approach for domain decomposition.
+The authors are curious to know if the reviewer also believes all MPI focused papers are unsuitable for SC.
+Devito uses a range of parallel programming models, usually nested, depending on the context.
+For example, SIMD for SIMD registers, OpenMP for multi-core shared memory, OpenMP 5 offloading and OpenACC for accelerators, and frameworks such as Dask for parallelizing task-graphs.
+For domain-decomposition parallelism on distributed memory computers, MPI is still the most widely accepted solution.
 
@@
 
-See R2 -- it's not just "grid decomposition", Devito deals with distribution of sparse functions, numpy arrays, decomposition over sub-domains, etc.
+The complexity arises from the fact that there are several layers of parallelism and parallel programming patterns.
+While we focus on MPI  for domain decomposition for this paper, a single application will also use threading, accelerators, SIMD vectorization, irregular computation, distributed numpy arrays.
 
@@
 
-The operations determining the performance of a code are the source-level ones. FMA is a lower-level concept (ie, a scalar FMA = two operations). This model is adopted everywhere.
+The operations determining the performance of software are the source-level ones. FMA is a lower-level concept (i.e., a scalar FMA = two operations). This model is adopted everywhere.
 
@@
 
-See R1 -- (i) lack of resources and (ii) Devito not ready yet for scalability tests when the experimentation was performed.
+See response to R1.
 
@@
 Further, as explained in R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
-To achieve this, we are using bleeding edge technologies -- an established DSL-based system such as Devito, Azure cloud, Docker, the rapidly-spreading GitHub-Actions for CI, etc.
-The use and orchestration of these technologies is thouroughly described in the paper, and is of strong interest for SC (strong industry interest for production scale).
+To achieve this, we are using state-of-the-art edge technologies -- an established DSL-based system such as Devito, Azure cloud, Docker, GitHub-Actions for CI, etc.
+The use and orchestration of these technologies is thoroughly described in the paper and is of keen interest for SC (strong industry interest for production scale).
 
@@
 
-The industrial problem does indeed involve muliple parts, such as data acquisition, processing, and, most importantly, velocity model building.
+The industrial problem does indeed involve multiple parts, such as data acquisition, processing, and, most importantly, velocity model building.
 We do not claim to solve all these problems.
-However, we demonstrate that the combination of advanced technolgies such as DSLs and compilers is key to highly-efficient sesimic modelling and imaging at industrial scale.
+However, we demonstrate that the combination of advanced technologies such as DSLs and compilers is key to highly-efficient seismic modelling and imaging at an industrial scale.
 We have added references and clarification in the introduction.
@@
 
-The new capabilities of Devito -- tensor language, MPI from high-level abstractions -- for multiple physics and discretizations is as important as a strong scaling experiment.
+The new capabilities of Devito -- tensor language, MPI from high-level abstractions -- for multiple physics and discretizations is as essential as a strong scaling experiment.
 Strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:
@@
 
-> for SC [...] sacalability and performance.
+> for SC [...] scalability and performance.
 
@@
 The inversion (outer-level of parallelism) requires what is essentially task parallelism; each task is internally MPI-parallel (domain decomposition).
-In this context task-parallelism means parallelization over sources, accomplished using `batch-shipyard` in a serverless setting [Section II].
+In this context, task-parallelism means parallelization over sources, accomplished using `batch-shipyard` in a serverless setting [Section II].
 This allowed optimal usage of the cloud, where one pays based on actual usage of computational resources.
@@
 
-GPU support is in development and currently (Devito-v4.2) supported for some applications. Its capabilities and related applications will be fully covered in a future paper.
+GPU support is in development and currently (Devito-v4.2) supported for some applications. Its capabilities and related applications will be covered in a forthcoming paper.
 

### 2020-06-05 response.md
Fix some response typos.

@@
 > If you run your model with a variable number of cores, how is the weak and strong scaling? Can you comment on that?
+
 We published and cited MPI scaling results in [@witte2019ecl;@luporini2019adp].
@@
 However, that computation unit of work sits within an embarrassingly parallel loop processing thousands of such shots as part of a gradient-decent algorithm.
-So while the computation of a single shot (the focus on Devito_ might only require TFLOPS, and iteration of the outer loop would be typically PFLOPS.
+So while the computation of a single shot (the focus on Devito might only require TFLOPS, and iteration of the outer loop would be typically PFLOPS).
 A description of nested-parallelism in applications such as FWI is outside of the scope of this paper, but we have included several references to FWI papers that do describe the underlying algorithm.
@@
 > Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality? How is that done?
+
 Details on solver verification have been added to the paper (in red).
@@
 > [...] not very suitable for SC
+
 There are few words in this review so we can only guess in what sense the reviewer feels the work is not suitable for SC.
 High productivity for HPC and performance-portability continue to be hot topics at SC.
-Many recent Gordon Bell finalists in recent years have used to some degree DSL's, and DSL's.
+Many Gordon Bell finalists in recent years have used to some degree DSL's.
 Devito is one of the very examples of DSL's to have challenged the performance of handwritten HPC codes for real engineering applications at scale.
@@
 However, one can conclude from the fact that Devito is being adopted by companies commercially to suggest comparisons were favourable.
-The research in Devito is highly relevant to SC as it demonstrates that the HPC cliche that one have to can have high-productivity languages or high-performance but not both is false.
+The research in Devito is highly relevant to SC as it demonstrates the HPC cliche that one can have high-productivity languages or high-performance, but not both, is false.
 
@@
 Whether or not parallelization is straightforward is not the issue. One can make the same comment about any finite difference, stencil code, anything involving linear algebra... 
-The novelty is that our work involves developing a layers of abstracts that allow finite difference based solvers be to implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism.
+The novelty is that our work involves developing layers of abstraction that allow finite difference based solvers to be implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism.
 This is not "straightforward" - the only comparable software technologies that we are aware of to have accomplished something similar are FEnICS and Firedrake for the finite-element method.
-We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed numpy arrays, and applies several optimizations for efficient halo exchange.
+We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed NumPy arrays, and applies several optimizations for efficient halo exchange.
 
@@
 
-High arithmetic intensity does not imply "non-challenging optimization".
+High arithmetic intensity does not imply "not as challenging for optimization".
 High-order finite (spectral) elements for instance -- where an extremely high arithmetic intensity can be obtained, but without tensor-product basis functions and optimizations such as sum-factorization results in an inflated GFlops/s measurement since a huge fraction of flops are redundant (e.g., common sub-expressions).
@@
 The complexity arises from the fact that there are several layers of parallelism and parallel programming patterns.
-While we focus on MPI  for domain decomposition for this paper, a single application will also use threading, accelerators, SIMD vectorization, irregular computation, distributed numpy arrays.
+While we focus on MPI  for domain decomposition for this paper, a single application will also use threading, accelerators, SIMD vectorization, irregular computation and distributed numpy arrays.
 
@@
 
-GPU support is in development and currently (Devito-v4.2) supported for some applications. Its capabilities and related applications will be covered in a forthcoming paper.
+GPU support is in development and currently (Devito-v4.2) covers some applications. Its capabilities and related applications will be covered in a forthcoming paper.
 

### 2020-06-05 response.md
Trim response.

@@
 
-We thank the reviewers for their work and constructive feedback. Below, we address their questions and observations. All edits are in red in the manuscript.
+We thank the reviewers for their work and constructive feedback. Below we address their questions and observations. Edits marked in red.
 
 ## R1
-> If you run your model with a variable number of cores, how is the weak and strong scaling? Can you comment on that?
+> If you run your model with a variable number of cores, how is the weak and strong scaling?
 
@@
 While Devito MPI scaling has improved significantly since those publications, it is only a relatively minor consideration within the bigger picture.
-Applications such as FWI generally only use MPI for computing on a single data shot (i.e.
-a single observation).
-However, that computation unit of work sits within an embarrassingly parallel loop processing thousands of such shots as part of a gradient-decent algorithm.
-So while the computation of a single shot (the focus on Devito might only require TFLOPS, and iteration of the outer loop would be typically PFLOPS).
-A description of nested-parallelism in applications such as FWI is outside of the scope of this paper, but we have included several references to FWI papers that do describe the underlying algorithm.
+Applications such as FWI generally only use MPI for computing a shot (single observation).
+That unit of work sits within an embarrassingly parallel loop processing thousands of shots as part of a gradient-descent algorithm.
+While computing a single shot the focus on Devito might only require TFLOPS -- iteration of the outer loop would typically be PFLOPS.
+A description of nested-parallelism in applications such as FWI is outside of the scope of this paper, but we have included several references to FWI papers that describe the underlying algorithm.
 
-> Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality? How is that done?
+> Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality?
 
-Details on solver verification have been added to the paper (in red).
-In [@devito-api], we analyze the numerical accuracy against the analytical solution and the convergence rate of the finite-differences.
+Details on solver verification have been added to the paper.
+In [@devito-api], we analyze the numerical accuracy against analytical solutions and the convergence rate of the finite-differences.
 We also test through continuous integration the numerical accuracy of all implemented models.
-Additionally, for inverse problems, we use the standard dot and gradient tests, "we focused on correct and testable implementations for the adjoint wave-equation and the gradient (action of the adjoint Jacobian) [@louboutin2018segeow; @louboutin2020THmfi]"
+For inverse problems, we use standard dot and gradient tests, "we focused on correct and testable implementations for the adjoint wave-equation and the gradient (action of the adjoint Jacobian)" [@louboutin2018segeow; @louboutin2020THmfi].
 
@@
 
-There are few words in this review so we can only guess in what sense the reviewer feels the work is not suitable for SC.
+There are few words in this review so we can only guess in what sense the reviewer feels the work is not suitable.
 High productivity for HPC and performance-portability continue to be hot topics at SC.
 Many Gordon Bell finalists in recent years have used to some degree DSL's.
-Devito is one of the very examples of DSL's to have challenged the performance of handwritten HPC codes for real engineering applications at scale.
-While Devito is fully open source, we have not been able to publish any head-to-head comparisons with commercial codes in this domain because of the commercial sensitivity.
-However, one can conclude from the fact that Devito is being adopted by companies commercially to suggest comparisons were favourable.
-The research in Devito is highly relevant to SC as it demonstrates the HPC cliche that one can have high-productivity languages or high-performance, but not both, is false.
+Devito is one of the few DSL's to have challenged the performance of handwritten HPC codes for real engineering applications at scale.
+While Devito is fully open source, we have not been able to publish any head-to-head comparisons with commercial codes in this domain owing to commercial sensitivity.
+However, one can conclude from the fact that Devito is being adopted by companies commercially that comparisons were favourable.
+The research is highly relevant to SC as it demonstrates the HPC cliche that one can have high-productivity languages or high-performance, but not both, is false.
 
@@
 Whether or not parallelization is straightforward is not the issue. One can make the same comment about any finite difference, stencil code, anything involving linear algebra...
-The novelty is that our work involves developing layers of abstraction that allow finite difference based solvers to be implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism.
-This is not "straightforward" - the only comparable software technologies that we are aware of to have accomplished something similar are FEnICS and Firedrake for the finite-element method.
-We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed NumPy arrays, and applies several optimizations for efficient halo exchange.
+The novelty our work is that it involves developing layers of abstraction to allow finite difference based solvers to be implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism.
+This is not "straightforward" - the only comparable software technologies that we are aware of are FEnICS and Firedrake.
+We also point out in the paper [Sections I and II-B] that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed NumPy arrays, and applies several optimizations for efficient halo exchange.
 
@@
 High arithmetic intensity does not imply "not as challenging for optimization".
-High-order finite (spectral) elements for instance -- where an extremely high arithmetic intensity can be obtained, but without tensor-product basis functions and optimizations such as sum-factorization results in an inflated GFlops/s measurement since a huge fraction of flops are redundant (e.g., common sub-expressions).
+Take high-order finite elements for instance -- extremely high arithmetic intensity can be easily obtained, but without tensor-product basis functions and optimizations such as sum-factorization results (GFlops/s) are inflated since a huge fraction of flops are redundant (e.g., common sub-expressions).
 TTI, a relatively sophisticated propagation model used in industry, is similar.
 As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtain very high GFlops/s, but terrible GPoints/s (and therefore poor time to solution).
-Optimizing away such redundancies, one of the compiler transformations performed by Devito,  is challenging to do by hand as there can be thousands of algebraic terms in these engineering applications.
+Optimizing away such redundancies, one of the compiler transformations performed by Devito,  is challenging to do by hand since there can be thousands of algebraic terms.
 
@@
 Devito uses a range of parallel programming models, usually nested, depending on the context.
-For example, SIMD for SIMD registers, OpenMP for multi-core shared memory, OpenMP 5 offloading and OpenACC for accelerators, and frameworks such as Dask for parallelizing task-graphs.
+For example, SIMD registers, OpenMP for multi-core shared memory, OpenMP 5 offloading and OpenACC for accelerators, and frameworks such as Dask for parallelizing task-graphs.
 For domain-decomposition parallelism on distributed memory computers, MPI is still the most widely accepted solution.
@@
 The complexity arises from the fact that there are several layers of parallelism and parallel programming patterns.
-While we focus on MPI  for domain decomposition for this paper, a single application will also use threading, accelerators, SIMD vectorization, irregular computation and distributed numpy arrays.
+While we here focus on MPI for domain decomposition, a single application will also use threading, accelerators, SIMD vectorization, irregular computation and distributed numpy arrays.
 
@@
 
-Correct. We explicitly refer to other articles for these other aspects.
+Correct. We explicitly refer to other articles for other aspects.
 
@@
 
-The operations determining the performance of software are the source-level ones. FMA is a lower-level concept (i.e., a scalar FMA = two operations). This model is adopted everywhere.
+Operations determining the performance of software are the source-level operations. FMA is a lower-level concept (scalar FMA = two operations). This model is widely adopted.
 
@@
 
-See response to R1.
+See R1 response.
 
@@
 We show performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations.
-Further, as explained in R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
-To achieve this, we are using state-of-the-art edge technologies -- an established DSL-based system such as Devito, Azure cloud, Docker, GitHub-Actions for CI, etc.
-The use and orchestration of these technologies is thoroughly described in the paper and is of keen interest for SC (strong industry interest for production scale).
+As explained in R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
+To achieve this, we are using state-of-the-art edge technologies -- an established DSL-based system, Azure cloud, Docker, GitHub-Actions for CI, etc.
+The use and orchestration of these technologies is thoroughly described in the paper and is of keen interest to SC (strong industry interest for production scale).
 
@@
 
-The industrial problem does indeed involve multiple parts, such as data acquisition, processing, and, most importantly, velocity model building.
+The industrial problem does indeed involve multiple parts, including data acquisition, processing, and, most importantly, velocity model building.
 We do not claim to solve all these problems.
-However, we demonstrate that the combination of advanced technologies such as DSLs and compilers is key to highly-efficient seismic modelling and imaging at an industrial scale.
+However, we demonstrate that the combination of advanced technologies such as DSLs and compilers is key to highly-efficient seismic modelling and imaging at industrial scale.
 We have added references and clarification in the introduction.
@@
 
-We have improved the manuscript to clarify what the contributions are. These are also now reiterated throughout the whole paper (edits in red).
+We have improved the manuscript to clarify new contributions. These are also reiterated throughout the whole paper.
 
@@
 
-The number of additions to the codebase is significant: https://github.com/devitocodes/devito/pull/873
+Short answer is significant: https://github.com/devitocodes/devito/pull/873
 
@@
 
-The single-node performance is discussed in other cited articles. See R1 about the lack of scalability experiments.
+The single-node performance is discussed in other cited articles. See R1 re. lack of scalability experiments.
 
@@
 
-Yes. We have improved the text to make this clear.
+Yes. Text improved for clarity.
 
@@
 In this context, task-parallelism means parallelization over sources, accomplished using `batch-shipyard` in a serverless setting [Section II].
-This allowed optimal usage of the cloud, where one pays based on actual usage of computational resources.
 
@@
 
-GPU support is in development and currently (Devito-v4.2) covers some applications. Its capabilities and related applications will be covered in a forthcoming paper.
+GPU support is in development and currently (Devito-v4.2) covers some applications. GPU capabilities will be covered in forthcoming work.
 

### 2020-06-05 paper.tex
rhodrin final tweaks (typos + grammar).

--- 2020-06-04 paper.tex
@@
 distributed-memory parallelism and imaging based on the tilted transverse
-isotropic (TTI) wave-equation (\cite{virieux, thomsen1986, zhang-tti,
-duveneck, louboutin2018segeow}). These proof of concepts highlight two
+isotropic (TTI) wave-equation~\cite{virieux, thomsen1986, zhang-tti,
+duveneck, louboutin2018segeow}. These proof of concepts highlight two
 critical features: first, the ability of the symbolic interface and the
@@
 the art hand-coded wave propagator to validate the performance previously
-benchmarked with the roofline model (\cite{patterson, devito-compiler,
-devito-api, louboutin2016ppf}). \textcolor{red}{Before concluding, results from the
+benchmarked with the roofline model \cite{patterson, devito-compiler,
+devito-api, louboutin2016ppf}. \textcolor{red}{Before concluding, results from the
 Cloud-based experiments introduced above are presented}.
@@
 mathematical level and implements automatic discretization with the
-finite-difference method. The language is by design flexible enough to
-express other types of non finite-difference operators, such as
-interpolation and tensor contractions, that are inherent to
-measurement-based inverse problems. Several additional features are
+finite-difference method. \textcolor{red}{The language is by design flexible enough to
+express other non finite-difference operators, such as
+interpolation and tensor contractions, inherent in
+measurement-based inverse problems}. Several additional features are
 supported, among which are staggered grids, sub-domains, and stencils with
@@
 are the boundary conditions which for finite-difference methods are
-notoriously diverse and often complicated. The system is, however,
+notoriously diverse and often complicated. The \textcolor{red}{DSL} is, however,
 sufficiently general to express them through a composition of core
@@
 \begin{cases}
- m(x) \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(xs) q(t) \\
+ m(x) \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(x_s) q(t) \\
  u(0, .) = \frac{\partial u(t, x)}{\partial t}(0, .) = 0 \\
- d(t, .) = u(t, xr).
+ d(t, .) = u(t, x_r).
  \end{cases}
@@
 %
-\textcolor{red}{where $u(t, x)$ is the pressure wavefield, $xs$ is the point source position, 
-$q(t)$ is the source time signature, $d(t, .)$ is the measured data at positions $xr$ and $m(x)$ is the
+\textcolor{red}{where $u(t, x)$ is the pressure wavefield, $x_s$ is the point source position, 
+$q(t)$ is the source time signature, $d(t, .)$ is the measured data at positions $x_r$ and $m(x)$ is the
 squared slowness.}
@@
 
-\subsection{3D Imaging example on Azure}\label{d-imaging-example-on-azure}
+\subsection{3D Imaging on Azure}\label{d-imaging-example-on-azure}
 
@@
 high-level definition of vectorial and tensorial physics. This API extension also implements 
-fully automatized staggered grid and staggered ginite-difference generally required for a stable discretization
+fully automatized staggered grid and staggered finite-differences which are generally required for a stable discretization
 of vectorial PDEs. This extension allows us to express an elastic model with compact and elegant notation 
-and model its solution computationally efficiently with the \devito compiler}
+and model its solution computationally efficiently with the \devito compiler}.
 
@@
 equations but also for coupled PDEs. This interface allows, in a simple, concise
-and consistent manner, the expression of all-kinds of non-trivial
-differential operators. Next, and most
+and consistent manner, the expression of all-kinds of non-trivial differential operators\textcolor{red}{: the vectorial elastic equation 
+and TTI (rotated finite-difference) being two examples presented in the present study}. Next, and most
 importantly, \textcolor{red}{it was} demonstrated that the compiler enables large-scale modelling with
@@
 
+\textcolor{red}{We emphasize that the work
+presented here does not solve (and does not claim to solve) all problems encountered within exploration geophysics. The high-level interface 
+and at (industry) scale performance demonstrated has resulted in
+multiple academic and industrial projects adopting the use \devito within their larger imaging frameworks.}
+
 \section{Code availability}
@@
   \item The complete code for TTI imaging is available at \url{https://github.com/slimgroup/Azure2019/tree/v1.0} and includes 
-  the TTI propagators, the Azure setup for shot parallelism and a documentation.
-  \item The elastic modelling can be run with the elastic example available in Devito at
-  \url{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/elastic/elastic_example.py} 
-  and can be run with any size and spatial order. A standalone script to run the large 3D elastic modelling
+  the TTI propagators, the Azure setup for shot parallelism and documentation.
+  \item \textcolor{red}{Elastic simulations can be run using the elastic example} available in Devito at
+  \url{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/elastic/elastic_example.py}.
+A standalone script to run the large 3D elastic modelling \textcolor{red}{experiment presented here}
   is also available at \url{https://github.com/mloubout/SC20Paper/tree/master/codesamples}

### 2020-06-05 response.md
100 words

@@
----
-title: response to reviews
-bibliography:
-	- sc20_paper.bib
----
 Dear Editor,
@@
 
-## R1
-> If you run your model with a variable number of cores, how is the weak and strong scaling?
+R1
 
-We published and cited MPI scaling results in [@witte2019ecl;@luporini2019adp].
+    If you run your model with a variable number of cores, how is the weak and strong scaling?
+
+We published and cited MPI scaling results (P.
+A.
+Witte et al., 2019; F.
+Luporini et al., 2019).
 While Devito MPI scaling has improved significantly since those publications, it is only a relatively minor consideration within the bigger picture.
@@
 That unit of work sits within an embarrassingly parallel loop processing thousands of shots as part of a gradient-descent algorithm.
-While computing a single shot the focus on Devito might only require TFLOPS -- iteration of the outer loop would typically be PFLOPS.
+While computing a single shot the focus on Devito might only require TFLOPS – iteration of the outer loop would typically be PFLOPS.
 A description of nested-parallelism in applications such as FWI is outside of the scope of this paper, but we have included several references to FWI papers that describe the underlying algorithm.
 
-> Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality?
+    validation of simulation quality?
 
 Details on solver verification have been added to the paper.
-In [@devito-api], we analyze the numerical accuracy against analytical solutions and the convergence rate of the finite-differences.
+In (M.
+Louboutin et al., 2019), we analyze the numerical accuracy against analytical solutions and the convergence rate of the finite-differences.
 We also test through continuous integration the numerical accuracy of all implemented models.
-For inverse problems, we use standard dot and gradient tests, "we focused on correct and testable implementations for the adjoint wave-equation and the gradient (action of the adjoint Jacobian)" [@louboutin2018segeow; @louboutin2020THmfi].
+For inverse problems, we use standard dot and gradient tests.
 
-## R2
+R2
 
-> [...] not very suitable for SC
+    […] not very suitable for SC
 
-There are few words in this review so we can only guess in what sense the reviewer feels the work is not suitable.
  High productivity for HPC and performance-portability continue to be hot topics at SC.
-Many Gordon Bell finalists in recent years have used to some degree DSL's.
-Devito is one of the few DSL's to have challenged the performance of handwritten HPC codes for real engineering applications at scale.
+Many recent Gordon Bell finalists have used to some degree DSL’s.
+Devito is one of the few DSL’s to have challenged the performance of handwritten HPC codes for real engineering applications at scale.
 While Devito is fully open source, we have not been able to publish any head-to-head comparisons with commercial codes in this domain owing to commercial sensitivity.
-However, one can conclude from the fact that Devito is being adopted by companies commercially that comparisons were favourable.
+However, one can conclude from the fact that Devito is being adopted by companies commercially that comparisons were favorable.
 The research is highly relevant to SC as it demonstrates the HPC cliche that one can have high-productivity languages or high-performance, but not both, is false.
 
-> Parallelization is pretty straightforward [...]
+    Parallelization is pretty straightforward […]
 
-Whether or not parallelization is straightforward is not the issue. One can make the same comment about any finite difference, stencil code, anything involving linear algebra...
-The novelty our work is that it involves developing layers of abstraction to allow finite difference based solvers to be implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism.
-This is not "straightforward" - the only comparable software technologies that we are aware of are FEnICS and Firedrake.
-We also point out in the paper [Sections I and II-B] that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed NumPy arrays, and applies several optimizations for efficient halo exchange.
+One can make the same comment about any finite difference, stencil code, linear algebra… The novelty our work is that it involves developing layers of abstraction to allow finite difference based solvers to be implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism.
+This is not “straightforward” - the only comparable software technologies would be FEnICS/Firedrake.
+We also point out in the paper [Sections I/II-B] that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed NumPy arrays, and applies several optimizations for efficient halo exchange.
 
-> high arithmetic intensity [...] challenging for optimization
+    high arithmetic intensity […]
 
-High arithmetic intensity does not imply "not as challenging for optimization".
-Take high-order finite elements for instance -- extremely high arithmetic intensity can be easily obtained, but without tensor-product basis functions and optimizations such as sum-factorization results (GFlops/s) are inflated since a huge fraction of flops are redundant (e.g., common sub-expressions).
-TTI, a relatively sophisticated propagation model used in industry, is similar.
-As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtain very high GFlops/s, but terrible GPoints/s (and therefore poor time to solution).
-Optimizing away such redundancies, one of the compiler transformations performed by Devito,  is challenging to do by hand since there can be thousands of algebraic terms.
+High arithmetic intensity does not imply “not as challenging for optimization”.
+Take high-order finite elements for instance – extremely high arithmetic intensity can be easily obtained, but without tensor-product basis functions and optimizations such as sum-factorization results (GFlops/s) are inflated since a huge fraction of flops are redundant (e.g., common sub-expressions).
+TTI, a sophisticated propagation model used in industry, is similar.
+As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives) among thousands of algebraic terms, we would obtain very high GFlops/s, but terrible GPoints/s (and therefore poor time to solution).
 
-> [...] MPI which is a bit behind times these days.
+    […] MPI which is a bit behind
 
 The authors are curious to know if the reviewer also believes all MPI focused papers are unsuitable for SC.
-Devito uses a range of parallel programming models, usually nested, depending on the context.
-For example, SIMD registers, OpenMP for multi-core shared memory, OpenMP 5 offloading and OpenACC for accelerators, and frameworks such as Dask for parallelizing task-graphs.
+Devito uses a range of parallel programming models, usually nested, depending on the context such as SIMD registers, OpenMP, OpenMP 5 offloading, OpenACC, and task-based parallelism (Dask/julia).
 For domain-decomposition parallelism on distributed memory computers, MPI is still the most widely accepted solution.
 
-## R3
+R3
 
-> little confusing but [...] style parallelism.
+    little confusing but […] style parallelism.
 
-The complexity arises from the fact that there are several layers of parallelism and parallel programming patterns.
-While we here focus on MPI for domain decomposition, a single application will also use threading, accelerators, SIMD vectorization, irregular computation and distributed numpy arrays.
+The complexity arises from the several layers of parallelisms. See R2 response.
 
-> some mention of threads, [...]. Only MPI is detailed.
+    Only MPI is detailed.
 
@@
 
-> [...] FMA?
+    […] FMA?
 
@@
 
-> Scalability performance
+    Scalability performance
 
@@
 
-> much of the performance [...]
+    much of the performance […]
 
-We show performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations.
+We show performance results of three different codes – TTI, elastic, and isotropic acoustic – thus exploring different physics and discretizations.
 As explained in R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system.
-To achieve this, we are using state-of-the-art edge technologies -- an established DSL-based system, Azure cloud, Docker, GitHub-Actions for CI, etc.
+To achieve this, we us state-of-the-art edge technologies – an established DSL-based system, Azure cloud, Docker, GitHub-Actions for CI, etc.
 The use and orchestration of these technologies is thoroughly described in the paper and is of keen interest to SC (strong industry interest for production scale).
 
-> For actual seismic [...] it helps a little.
+    For actual seismic […] it helps a little.
 
@@
 
-> Dropping [...] 3D elastic [...] might demonstrate [...].
+    Dropping […] 3D elastic […]
 
-The new capabilities of Devito -- tensor language, MPI from high-level abstractions -- for multiple physics and discretizations is as essential as a strong scaling experiment.
-Strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:
+The new capabilities of Devito – tensor language, MPI from high-level abstractions – for multiple physics and discretizations are as essential as a strong scaling experiment. See R1 response.
+R4
 
-* even with high-frequency FWI, a typical shot-level computation requires no more than a bunch of nodes (often order of units);
-* there is an outer-level of parallelism for inversion, essentially a task farm where the individual workers are internally MPI-parallel.
+    […] how much is new
 
-## R4
+We have improved the manuscript to clarify new contributions.
 
-> [...] how much of this work is new
-
-We have improved the manuscript to clarify new contributions. These are also reiterated throughout the whole paper.
-
-> [...] vector and tensors [...] unclear [...] involved.
+    […] vector and tensors […]
 
@@
 
-> for SC [...] scalability and performance.
+    for SC […] scalability and performance.
 
@@
 
-> section III describing new work?
+    section III describing new work?
 
@@
 
-> parallelism just via MPI?
+    parallelism just via MPI?
 
-The inversion (outer-level of parallelism) requires what is essentially task parallelism; each task is internally MPI-parallel (domain decomposition).
-In this context, task-parallelism means parallelization over sources, accomplished using `batch-shipyard` in a serverless setting [Section II].
+The inversion (outer-level of parallelism) requires what is essentially task-parallelism; each task is internally MPI-parallel.
+In this context, task-parallelism means parallelization over sources, accomplished using batch-shipyard in a serverless setting [Section II].
 
-> What are the possibilities [...]
+    What are the possibilities […]
 
-GPU support is in development and currently (Devito-v4.2) covers some applications. GPU capabilities will be covered in forthcoming work.
+GPU support is in development and currently (Devito-v4.2) covers some applications and will be covered in forthcoming work.
 
-> Minor issues [...]
+    Minor issues.
 
-Fixed, thanks.
+Fixed.

### 2020-06-05 paper.tex
100 words

@@
 
-\author{
-\IEEEauthorblockN{1\textsuperscript{st} Mathias Louboutin}
-\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
-Atlanta, GA \\
-mlouboutin3@gatech.edu}
+% \author{
+% \IEEEauthorblockN{1\textsuperscript{st} Mathias Louboutin}
+% \IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
+% Atlanta, GA \\
+% mlouboutin3@gatech.edu}
+% \and
+% \IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
+% \IEEEauthorblockA{\textit{Devito Codes}\\
+% London, UK \\
+% fabio@devitocodes.com}
+% \and
+% \IEEEauthorblockN{3\textsuperscript{rd} Rhodri Nelson}
+% \IEEEauthorblockA{\textit{Imperial College London}\\
+% London, UK \\
+% rhodri.nelson@imperial.ac.uk}
+% \and
+% \IEEEauthorblockN{4\textsuperscript{th} Philipp Witte}
+% \IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
+% Atlanta, GA \\
+% pwitte3@gatech.edu}
+% \and
+% \IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
+% \IEEEauthorblockA{\textit{Imperial College London}\\
+% London, UK \\
+% g.bisbas18@imperial.ac.uk}
+% \and
+% \IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
+% \IEEEauthorblockA{\textit{TU-Delft}\\
+% Delft, NL \\
+% J.W.Thorbecke@tudelft.nl}
+% \and
+% \IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
+% \IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
+% Atlanta, GA\\
+% felix.herrmann@gatech.edu}
+% \and
+% \IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
+% \IEEEauthorblockA{\textit{Imperial College London}\\
+% London, UK \\
+% g.gorman@imperial.ac.uk}
+% }
+
+\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
 \and
-\IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
-\IEEEauthorblockA{\textit{Devito Codes}\\
-London, UK \\
-fabio@devitocodes.com}
+\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
 \and
-\IEEEauthorblockN{3\textsuperscript{rd} Rhodri Nelson}
-\IEEEauthorblockA{\textit{Imperial College London}\\
-London, UK \\
-rhodri.nelson@imperial.ac.uk}
+\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
 \and
-\IEEEauthorblockN{4\textsuperscript{th} Philipp Witte}
-\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
-Atlanta, GA \\
-pwitte3@gatech.edu}
+\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
 \and
-\IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
-\IEEEauthorblockA{\textit{Imperial College London}\\
-London, UK \\
-g.bisbas18@imperial.ac.uk}
+\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
 \and
-\IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
-\IEEEauthorblockA{\textit{TU-Delft}\\
-Delft, NL \\
-J.W.Thorbecke@tudelft.nl}
-\and
-\IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
-\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
-Atlanta, GA\\
-felix.herrmann@gatech.edu}
-\and
-\IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
-\IEEEauthorblockA{\textit{Imperial College London}\\
-London, UK \\
-g.gorman@imperial.ac.uk}
+\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
 }
-
-% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID}
-% \and
-% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID}
-% \and
-% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID}
-% \and
-% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID}
-% \and
-% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID}
-% \and
-% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID}
-% }
 
